---
title: "03_TF_activity_kidney"
author: "Jordan Whitlock"
date: '2023-06-07'
output: html_document
---
# Setup
```{r load-libraries-functions}
library(ComplexHeatmap)
library(circlize)
library(here)
library(readr)
library(purrr)
library(styler)
library(dplyr)
library(gprofiler2)
library(Seurat)
library(tidyr)
library(tibble)
library(patchwork)
library(cowplot)
source(here("src/functions/functions.R"))
ptm <- proc.time()
```

```{r load-geneset}
# load in gene set for Setbp1
setbp1_genes <- read.csv(here("results/seurat/setbp1_targets.csv"))
setbp1_genes <- setbp1_genes[, -1] %>% as.vector()
setbp1_genes <- append(setbp1_genes, "Setbp1")
```

```{r}
load(here("data/kidney_integrated_celltypes_postSoup.Rdata"))
```

# kidney: 
```{r}
files <- list.files(here("results/decoupleR/kidney"), full.names = TRUE)

name_acts_list <- list()

# calculating the average TF activity score per gene within eac cell type across all cells
for (i in files) {
  load(i) # object name: acts
  name <- basename(i)
  name <- gsub("_acts.RData", "", name) # grab only portion of name before '_acts.RData'
  name_acts <- acts
  name_acts_list[[name]] <- name_acts
  print(paste0("loaded acts for ", name))
}

# combining and fixing cell type barcode convention issue
kidney_acts <- bind_rows(name_acts_list)
kidney_acts$condition <- gsub("\\.1$", "-1", kidney_acts$condition)
kidney_acts[1:6, 1:6]
```

```{r}
# Extract mlm and store it in tfmlm in kidney_int_celltypes
kidney_int_celltypes[["tfmlm"]] <- kidney_acts %>%
  pivot_wider(
    id_cols = "source", names_from = "condition",
    values_from = "score"
  ) %>%
  column_to_rownames("source") %>%
  Seurat::CreateAssayObject(.)
```

```{r}
# Change assay
DefaultAssay(object = kidney_int_celltypes) <- "tfmlm"

# Scale the data
kidney_int_celltypes <- ScaleData(kidney_int_celltypes)
kidney_int_celltypes@assays$tfmlm@data <- kidney_int_celltypes@assays$tfmlm@scale.data
```

```{r split-object}
split <- SplitObject(kidney_int_celltypes, split.by = "type")
```

```{r}
het_df <- t(as.matrix(split$heterozygous@assays$tfmlm@data)) %>%
  as.data.frame() %>%
  mutate(cluster = Idents(split$heterozygous)) %>%
  pivot_longer(cols = -cluster, names_to = "source", values_to = "score") %>%
  group_by(cluster, source) %>%
  summarise(mean = mean(score))

hist(het_df$mean)

ctrl_df <- t(as.matrix(split$control@assays$tfmlm@data)) %>%
  as.data.frame() %>%
  mutate(cluster = Idents(split$control)) %>%
  pivot_longer(cols = -cluster, names_to = "source", values_to = "score") %>%
  group_by(cluster, source) %>%
  summarise(mean = mean(score))

hist(ctrl_df$mean)
```
Since data is on the same scale, in order to compare between conditions we wanted to calculate the percentage change of the Setbp1 and its gene set between condtions (and for all TFs)

Calculating the absolute percentage change in order to see which TFs exhibit the largest difference in activities are most different between conditions. The more positive percent change indicates that the magnitude of the change in TF activity between conditions was higher in the S858R than WT and the inverse for WT. All percent changes were then quartile filtered to only include TFs that had a percent change > Q3 or < Q1 for S858R and WT respectively. Original TF Activity values were then plotted for these TFs across cell types in each respective tissue, filtering the TFs to only focus on those that are in the Setbp1 gene set or Setbp1.
```{r}
# merge control and het dataframes
merged_df <- merge(ctrl_df, het_df, by = c("cluster", "source"))
colnames(merged_df) <- c("cluster", "source", "ctrl_mean", "het_mean")

# calculate percent change between mean TF activity between condition; large positive percent change indicates the magnitude of average TF acitvity for that TF was larger in the S858R
merged_df$percent_change <- ((abs(merged_df$het_mean) - abs(merged_df$ctrl_mean)) / abs(merged_df$ctrl_mean)) * 100

# next want to filter the data at a specific threshold; using the tail quartiles
summary(merged_df$percent_change) # Q3 is at 35.1, Q1 is at -35.3
merged_df_Q3 <- merged_df[merged_df$percent_change > 35.1, ]
merged_df_Q1 <- merged_df[merged_df$percent_change < -35.3, ]

# combining into single data frame
merged_df <- rbind(merged_df_Q3, merged_df_Q1)

# grabbing pos and neg for plotting
merged_df_pos <- subset(merged_df, merged_df$percent_change > 0)

geneset_acts_mat_pos <- merged_df_pos %>%
  filter(source %in% setbp1_genes) %>%
  pivot_wider(
    id_cols = "cluster", names_from = "source",
    values_from = "het_mean"
  ) %>%
  column_to_rownames("cluster") %>%
  mutate_all(~ replace(., is.na(.), 0))


merged_df_neg <- subset(merged_df, merged_df$percent_change < 0)

geneset_acts_mat_neg <- merged_df_neg %>%
  filter(source %in% setbp1_genes) %>%
  pivot_wider(
    id_cols = "cluster", names_from = "source",
    values_from = "het_mean"
  ) %>%
  column_to_rownames("cluster") %>%
  mutate_all(~ replace(., is.na(.), 0))


# order TFs same in both

## id common columns and sort
common <- intersect(colnames(geneset_acts_mat_neg), colnames(geneset_acts_mat_pos))
sorted <- sort(common)

## id non overlapping columns
non_pos <- setdiff(colnames(geneset_acts_mat_pos), common)
non_neg <- setdiff(colnames(geneset_acts_mat_neg), common)

geneset_acts_mat_pos <- cbind(geneset_acts_mat_pos[, sorted], geneset_acts_mat_pos[, non_pos])

geneset_acts_mat_neg <- cbind(geneset_acts_mat_neg[, sorted], geneset_acts_mat_neg[, non_neg])

# Choose color palette
palette_length <- 100
my_color <- colorRampPalette(c("Darkblue", "white", "red"))(palette_length)

my_breaks <- c(
  seq(-3, 0, length.out = ceiling(palette_length / 2) + 1),
  seq(0.05, 3, length.out = floor(palette_length / 2))
)

## plot positive
png(
  filename = here("results/TF_activity/kidney_setbp1_geneset_pos.png"),
  width = 900,
  height = 1000,
  res = 250
)
pheatmap(t(geneset_acts_mat_pos), border_color = NA, color = my_color, breaks = my_breaks, cluster_rows = FALSE, cluster_cols = FALSE)
dev.off()

## plot negative
png(
  filename = here("results/TF_activity/kidney_setbp1_geneset_neg.png"),
  width = 900,
  height = 1000,
  res = 250
)
pheatmap(t(geneset_acts_mat_neg), border_color = NA, color = my_color, breaks = my_breaks, cluster_rows = FALSE, cluster_cols = FALSE)
dev.off()
```

```{r processing-time}
fptm <- proc.time() - ptm
fptm[3] / 60
```
 elapsed 
2.731617 

```{r style-file}
# run style
style_file(here("src/network_scripts/decoupleR/03_TF_activity_kidney.Rmd")) # commented out after being run once
# lintr was run as well
```

```{r session-info}
sessionInfo()
```
R version 4.1.3 (2022-03-10)
Platform: x86_64-pc-linux-gnu (64-bit)
Running under: Ubuntu 20.04.6 LTS

Matrix products: default
BLAS/LAPACK: /usr/lib/x86_64-linux-gnu/openblas-pthread/libopenblasp-r0.3.8.so

locale:
[1] C

attached base packages:
[1] grid      stats     graphics  grDevices utils     datasets  methods   base     

other attached packages:
 [1] ggplot2_3.4.1         cowplot_1.1.1         patchwork_1.1.2       tibble_3.1.8          tidyr_1.3.0           SeuratObject_4.1.3    Seurat_4.3.0         
 [8] gprofiler2_0.2.1      dplyr_1.1.0           styler_1.9.0          purrr_1.0.1           readr_2.1.4           here_1.0.1            circlize_0.4.15      
[15] ComplexHeatmap_2.10.0

loaded via a namespace (and not attached):
  [1] Rtsne_0.16             colorspace_2.1-0       deldir_1.0-6           rjson_0.2.21           ellipsis_0.3.2         ggridges_0.5.4        
  [7] rprojroot_2.0.3        GlobalOptions_0.1.2    spatstat.data_3.0-0    clue_0.3-64            rstudioapi_0.14        leiden_0.4.3          
 [13] listenv_0.9.0          ggrepel_0.9.3          fansi_1.0.4            codetools_0.2-18       splines_4.1.3          R.methodsS3_1.8.2     
 [19] doParallel_1.0.17      knitr_1.42             polyclip_1.10-4        jsonlite_1.8.4         ica_1.0-3              cluster_2.1.2         
 [25] png_0.1-8              R.oo_1.25.0            uwot_0.1.14            spatstat.sparse_3.0-0  shiny_1.7.4            sctransform_0.3.5     
 [31] compiler_4.1.3         httr_1.4.5             Matrix_1.5-3           fastmap_1.1.1          lazyeval_0.2.2         cli_3.6.0             
 [37] later_1.3.0            htmltools_0.5.4        tools_4.1.3            igraph_1.4.1           gtable_0.3.1           glue_1.6.2            
 [43] RANN_2.6.1             reshape2_1.4.4         Rcpp_1.0.10            scattermore_0.8        vctrs_0.5.2            nlme_3.1-155          
 [49] spatstat.explore_3.0-6 progressr_0.13.0       iterators_1.0.14       lmtest_0.9-40          spatstat.random_3.1-3  stringr_1.5.0         
 [55] xfun_0.37              globals_0.16.2         mime_0.12              miniUI_0.1.1.1         lifecycle_1.0.3        irlba_2.3.5.1         
 [61] goftest_1.2-3          future_1.31.0          MASS_7.3-55            zoo_1.8-11             scales_1.2.1           spatstat.utils_3.0-1  
 [67] hms_1.1.2              promises_1.2.0.1       parallel_4.1.3         RColorBrewer_1.1-3     gridExtra_2.3          reticulate_1.28       
 [73] pbapply_1.7-0          stringi_1.7.12         S4Vectors_0.32.4       foreach_1.5.2          BiocGenerics_0.40.0    shape_1.4.6           
 [79] rlang_1.0.6            pkgconfig_2.0.3        matrixStats_0.63.0     lattice_0.20-45        tensor_1.5             ROCR_1.0-11           
 [85] htmlwidgets_1.6.1      tidyselect_1.2.0       parallelly_1.34.0      RcppAnnoy_0.0.20       plyr_1.8.8             magrittr_2.0.3        
 [91] R6_2.5.1               IRanges_2.28.0         generics_0.1.3         withr_2.5.0            pillar_1.8.1           fitdistrplus_1.1-8    
 [97] abind_1.4-5            survival_3.3-1         sp_1.6-0               future.apply_1.10.0    crayon_1.5.2           KernSmooth_2.23-20    
[103] utf8_1.2.3             spatstat.geom_3.0-6    plotly_4.10.1          tzdb_0.3.0             GetoptLong_1.0.5       data.table_1.14.8     
[109] digest_0.6.31          xtable_1.8-4           R.cache_0.16.0         httpuv_1.6.9           R.utils_2.12.2         stats4_4.1.3          
[115] munsell_0.5.0          viridisLite_0.4.1     
