---
title: "02_MouseSetbp1_decoupleR_expression"
author: "Jordan Whitlock"
date: '2023-05-12'
output: html_document
---

```{r}
set.seed(2178)
library(lintr)
library(styler)
library(here)
ptm <- proc.time()
```

```{r}
load(here("data/kidney_integrated_celltypes_postSoup.Rdata"))
load(here("data/setbp1_cerebralintcelltypes.Rdata"))
```

```{r}
cerebral_decoupleR_expression <- list()

for (i in levels(cerebral_int_celltypes)) {
  data <- subset(cerebral_int_celltypes, subset = (type == "heterozygous" & cell_type == i))
  data <- as.data.frame(data@assays$RNA@data)
  cerebral_decoupleR_expression[[paste0("heterozygous_", i)]] <- data

  data <- subset(cerebral_int_celltypes, subset = (type == "control" & cell_type == i))
  data <- as.data.frame(data@assays$RNA@data)
  cerebral_decoupleR_expression[[paste0("control_", i)]] <- data
}

save(cerebral_decoupleR_expression,
  file = here("data/processed/decoupleR_expression_inputs/cerebral_expression.Rdata")
)
```

```{r}
kidney_decoupleR_expression <- list()

for (i in levels(kidney_int_celltypes)) {
  data <- subset(kidney_int_celltypes,
    subset = (type == "heterozygous" & cell_type == i)
  )
  data <- as.data.frame(data@assays$RNA@data)
  kidney_decoupleR_expression[[paste0("heterozygous_", i)]] <- data

  data <- subset(kidney_int_celltypes,
    subset = (type == "control" & cell_type == i)
  )
  data <- as.data.frame(data@assays$RNA@data)
  kidney_decoupleR_expression[[paste0("control_", i)]] <- data
}

save(kidney_decoupleR_expression,
  file = here("data/processed/decoupleR_expression_inputs/kidney_expression.Rdata")
)
```

```{r}
style_file(here("src/network_scripts/decoupleR_input_construction/02_MouseSetbp1_decoupleR_expression.Rmd"))

fptm <- proc.time() - ptm
(fptm[3] / 60)

sessionInfo()
```
elapsed 
 8.7971 
 
 R version 4.1.3 (2022-03-10)
Platform: x86_64-pc-linux-gnu (64-bit)
Running under: Ubuntu 20.04.5 LTS

Matrix products: default
BLAS/LAPACK: /usr/lib/x86_64-linux-gnu/openblas-pthread/libopenblasp-r0.3.8.so

locale:
[1] C

attached base packages:
[1] stats     graphics  grDevices utils     datasets  methods   base     

other attached packages:
[1] here_1.0.1   styler_1.9.0 lintr_3.0.2 

loaded via a namespace (and not attached):
  [1] Seurat_4.3.0           Rtsne_0.16             colorspace_2.1-0       deldir_1.0-6           ellipsis_0.3.2        
  [6] ggridges_0.5.4         rprojroot_2.0.3        spatstat.data_3.0-0    rstudioapi_0.14        leiden_0.4.3          
 [11] listenv_0.9.0          remotes_2.4.2          ggrepel_0.9.3          fansi_1.0.4            xml2_1.3.3            
 [16] codetools_0.2-18       splines_4.1.3          R.methodsS3_1.8.2      knitr_1.42             polyclip_1.10-4       
 [21] jsonlite_1.8.4         ica_1.0-3              cluster_2.1.2          png_0.1-8              R.oo_1.25.0           
 [26] uwot_0.1.14            spatstat.sparse_3.0-0  shiny_1.7.4            sctransform_0.3.5      compiler_4.1.3        
 [31] httr_1.4.5             SeuratObject_4.1.3     Matrix_1.5-3           fastmap_1.1.1          lazyeval_0.2.2        
 [36] cli_3.6.0              later_1.3.0            htmltools_0.5.4        tools_4.1.3            igraph_1.4.1          
 [41] gtable_0.3.1           glue_1.6.2             RANN_2.6.1             reshape2_1.4.4         dplyr_1.1.0           
 [46] Rcpp_1.0.10            scattermore_0.8        vctrs_0.5.2            nlme_3.1-155           spatstat.explore_3.0-6
 [51] progressr_0.13.0       lmtest_0.9-40          spatstat.random_3.1-3  stringr_1.5.0          xfun_0.37             
 [56] globals_0.16.2         ps_1.7.2               mime_0.12              miniUI_0.1.1.1         lifecycle_1.0.3       
 [61] irlba_2.3.5.1          goftest_1.2-3          future_1.31.0          MASS_7.3-55            zoo_1.8-11            
 [66] scales_1.2.1           spatstat.utils_3.0-1   promises_1.2.0.1       rex_1.2.1              parallel_4.1.3        
 [71] RColorBrewer_1.1-3     gridExtra_2.3          reticulate_1.28        pbapply_1.7-0          ggplot2_3.4.1         
 [76] stringi_1.7.12         desc_1.4.2             cyclocomp_1.1.0        rlang_1.0.6            pkgconfig_2.0.3       
 [81] matrixStats_0.63.0     lattice_0.20-45        tensor_1.5             ROCR_1.0-11            purrr_1.0.1           
 [86] patchwork_1.1.2        htmlwidgets_1.6.1      cowplot_1.1.1          processx_3.8.0         tidyselect_1.2.0      
 [91] parallelly_1.34.0      RcppAnnoy_0.0.20       plyr_1.8.8             magrittr_2.0.3         R6_2.5.1              
 [96] generics_0.1.3         pillar_1.8.1           withr_2.5.0            fitdistrplus_1.1-8     abind_1.4-5           
[101] survival_3.3-1         sp_1.6-0               tibble_3.1.8           future.apply_1.10.0    crayon_1.5.2          
[106] KernSmooth_2.23-20     utf8_1.2.3             spatstat.geom_3.0-6    plotly_4.10.1          grid_4.1.3            
[111] data.table_1.14.8      callr_3.7.3            digest_0.6.31          xtable_1.8-4           R.cache_0.16.0        
[116] tidyr_1.3.0            httpuv_1.6.9           R.utils_2.12.2         munsell_0.5.0          viridisLite_0.4.1  
