---
title: "01_decoupleR_formatting_prior"
author: "Jordan Whitlock"
date: '2023-05-15'
output: html_document
---

Note for Code Reviewer, this was run in 1.0.5 docker

```{r}
set.seed(2178)
library(here)
library(reshape)
source(here("src/functions/functions.R"))
ptm <- proc.time()
```


```{r}
#make list of expression files needed to loop through
files <- list.files(here("results/decoupleR_PANDA"), full.names = TRUE)
files

#run for loop
for(i in files){
  #load in the expression data files needed and rename:
  PANDA <- loadRData(i)
  #extract regNet and melt into `source`, `target` and `mor`, required format for decoupleR.
  #These equate to `TF`, `gene` and `score` respectively. 
  regNet <- PANDA@regNet
  regNet <- melt(regNet)
  colnames(regNet) <- c("source", "target", "mor")
  # #save melted output
  name <- sub("_PANDA.Rdata", "", basename(i))
  save(regNet, file = paste0(here("results/decoupleR_PANDA_melted/"), name, "_melted.Rdata"))
  rm(PANDA)
  rm(regNet)
  print(paste0(name, " network melted and saved."))
}
```
```{r}
fptm[3]/60 <- proc.time() - ptm
```


```{r}
sessionInfo()
```

R version 4.1.3 (2022-03-10)
Platform: x86_64-pc-linux-gnu (64-bit)
Running under: Ubuntu 20.04.5 LTS

Matrix products: default
BLAS/LAPACK: /usr/lib/x86_64-linux-gnu/openblas-pthread/libopenblasp-r0.3.8.so

locale:
[1] C

attached base packages:
[1] stats     graphics  grDevices utils     datasets  methods   base     

other attached packages:
[1] ggplot2_3.4.1       reshape_0.8.9       here_1.0.1          pandaR_1.26.0       Biobase_2.54.0     
[6] BiocGenerics_0.40.0

loaded via a namespace (and not attached):
 [1] Rcpp_1.0.10        pillar_1.8.1       compiler_4.1.3     plyr_1.8.8         tools_4.1.3        digest_0.6.31     
 [7] evaluate_0.20      lifecycle_1.0.3    tibble_3.1.8       gtable_0.3.1       lattice_0.20-45    pkgconfig_2.0.3   
[13] rlang_1.0.6        igraph_1.4.1       cli_3.6.0          rstudioapi_0.14    yaml_2.3.7         hexbin_1.28.2     
[19] xfun_0.37          fastmap_1.1.1      withr_2.5.0        dplyr_1.1.0        knitr_1.42         generics_0.1.3    
[25] vctrs_0.5.2        rprojroot_2.0.3    grid_4.1.3         tidyselect_1.2.0   glue_1.6.2         R6_2.5.1          
[31] fansi_1.0.4        rmarkdown_2.20     magrittr_2.0.3     htmltools_0.5.4    scales_1.2.1       matrixStats_0.63.0
[37] RUnit_0.4.32       colorspace_2.1-0   utf8_1.2.3         munsell_0.5.0 