---
_title: "230315_ELISA_BCA"
author: "Jordan Whitlock"
date: '2023-03-14'
output: html_document
---

```{r library-seed-setup}
set.seed(2178)
library(here)
library(readr)
library(ggplot2)
library(ggpubr)
library(styler)
library(lintr)
library(tidyr)
```

```{r load-exp-results}
ELISA <- read_csv(here("results/ELISA/230315_jw_setbp1ELISA.csv"))
BCA <- read_csv(here("results/ELISA/230315_jw_setbp1BCA.csv"))
```


```{r analyzing-ELISA}
# average entire dataset except blanks:
measurements <- t(colMeans(ELISA[, -c(7:8)]))

# average blanks
blanks <- ELISA[, c(7:8)]
blanks$sum <- rowSums(blanks)
blank <- colSums(blanks) / 6
blank <- blank[3]


# subtract blank from measurement
measurements <- measurements - blank

# add concentration
concentrations <- c(3.12, 6.25, 12.5, 25, 50, 100)

standards <- rbind(measurements[, 1:6], concentrations)
row.names(standards) <- c("OD", "concentration")

# fit curve and get formula for standards
standards <- as.data.frame(t(standards))

ggplot(data = standards) +
  geom_point(aes(x = concentration, y = OD)) +
  theme_bw() +
  xlab("Concentration") +
  ylab("Optical Density (OD) at 450 nm") +
  geom_smooth(aes(x = concentration, y = OD),
    method = lm, se = FALSE,
    formula = y ~ x
  )

# extract intercept and slope from line
line <- lm(standards$OD ~ standards$concentration)
int <- line$coefficients[1]
int # 0.1075113 0.06374742
slope <- line$coefficients[2]
slope # 0.0129213

# using formula for line, determine unknown concentration from kidney_rep and cortex_rep (formula: OD  = slope*concentration + int, concentration = (OD - int/slope)
sample_concentrations <- (measurements[, 7:10] - int) / slope
sample_concentrations <- unlist(as.vector(c(concentrations, sample_concentrations)))
sample_concentrations

# wrangle and format concentrations
setbp1ELISA_data <- rbind(measurements, sample_concentrations)

setbp1ELISA_data <- as.data.frame(t(setbp1ELISA_data))
setbp1ELISA_data$sample_ID <- c(
  "standard", "standard", "standard", "standard",
  "standard", "standard", "cerebral cortex het",
  "cerebral cortex WT", "kidney het", "kidney WT"
)
colnames(setbp1ELISA_data) <- c("OD", "concentration", "sample_ID")

# plot data
ggplot(data = setbp1ELISA_data) +
  geom_point(
    aes(
      x = concentration,
      y = OD,
      color = sample_ID
    ),
    position = "jitter"
  ) +
  theme_bw() +
  xlab("Concentration") +
  ylab("Optical Density (OD) at 450 nm")

# plot data with eq
png(here("results/ELISA/setbp1ELISA_quantification.png"),
  width = 1000, height = 1000
)
ggscatter(
  data = setbp1ELISA_data,
  x = "concentration",
  y = "OD",
  color = "sample_ID",
  size = 3, shape = "sample_ID",
  palette = get_palette(c("#644DD4", "#5F9EFD", "#FE6100", "#FFB000", "black"), 5),
  add = "reg.line", add.params = list(color = "black"),
  ggtheme = theme_bw(),
  xlab = "Concentration",
  ylab = "Optical Density (OD) at 450 nm",
  title = "Setbp1 ELISA Quantification"
) +
  stat_regline_equation(label.x = 3, label.y = 1.2)
dev.off()
```

```{r plotting-individual-ELISA-reps}
measurements <- ELISA[, c(1:6, 9:12)] - blank

# add concentration
concentrations <- c(3.12, 6.25, 12.5, 25, 50, 100)

standards <- rbind(measurements[, 1:6], concentrations)
row.names(standards) <- c("rep1", "rep2", "rep3", "concentration")

# fit curve and get formula for standards
ELISA_repdata <- as.data.frame(t(standards))
ELISA_repdata <- pivot_longer(ELISA_repdata, values_to = "OD", cols = 1:3)

ggplot(data = ELISA_repdata) +
  geom_point(aes(x = concentration, y = OD)) +
  theme_bw() +
  xlab("Concentration") +
  ylab("Optical Density (OD) at 450 nm") +
  geom_smooth(aes(x = concentration, y = OD),
    method = lm,
    se = FALSE,
    formula = y ~ x
  )

# extract intercept and slope from line
line <- lm(ELISA_repdata$OD ~ ELISA_repdata$concentration)
int <- line$coefficients[1]
int # 0.1075113
slope <- line$coefficients[2]
slope # 0.0129213

# using formula for line, determine unknown concentration from kidney_rep and cortex_rep (formula: OD  = slope*concentration + int, concentration = (OD - int/slope)
sample_concentrations <- (measurements[, 7:10] - int) / slope
sample_concentrations <- unlist(as.vector(sample_concentrations))
sample_concentrations

# wrangle and format concentrations
sample_ID <- c(
  "cortex_het", "cortex_het", "cortex_het", "cortex_wt",
  "cortex_wt", "cortex_wt", "kidney_het", "kidney_het",
  "kidney_het", "kidney_wt", "kidney_wt", "kidney_wt"
)
rep_conc_ELISA <- as.data.frame(cbind(sample_ID, as.numeric(sample_concentrations)))

# plot data
png(here("results/ELISA/setbp1ELISA_quantification_boxplot.png"),
  width = 1000, height = 1000
)
ggplot(data = rep_conc_ELISA, aes(
  x = sample_ID,
  y = sample_concentrations,
  fill = sample_ID
)) +
  geom_boxplot() +
  geom_point() +
  theme_bw() +
  xlab("Group") +
  ylab("Concentration") +
  stat_boxplot(
    geom = "errorbar", # Boxplot with error bars
    width = 0.2
  ) +
  ggtitle("Setbp1 Protein Concentration") +
  scale_fill_manual(values = c(
    "kidney_wt" = "#FE6100",
    "kidney_het" = "#FFB000",
    "cortex_het" = "#5F9EFD",
    "cortex_wt" = "#644DD4"
  ))
dev.off()
```

```{r analyzing-BCA}
# average entire dataset except blank:
measurements <- t(colMeans(BCA[, 2:13]))

# grabbing blank
blanks <- BCA[, 1]
blank <- colSums(blanks) / 3

# subtract blank from measurement
measurements <- measurements - blank

# add concentration
concentrations <- c(25, 125, 250, 500, 750, 1000, 1500, 2000)

standards <- rbind(measurements[, 1:8], concentrations)
row.names(standards) <- c("OD", "concentration")

# fit curve and get formula for standards
standards <- as.data.frame(t(standards))
standards <- standards[1:5, ] # note: we dropped 1000 (standard 7), 1500 (standard 8), and 2000 (standard 9)

ggplot(data = standards) +
  geom_point(aes(x = concentration, y = OD)) +
  theme_bw() +
  xlab("Concentration") +
  ylab("Optical Density (OD) at 450 nm") +
  geom_smooth(aes(x = concentration, y = OD),
    method = lm, se = FALSE,
    formula = y ~ x
  )

# extract intercept and slope from line
line <- lm(standards$OD ~ standards$concentration)
int <- line$coefficients[1]
int # 0.2909356
slope <- line$coefficients[2]
slope # 0.0006567205

# using formula for line, determine unknown concentration from kidney_rep and cortex_rep (formula: OD  = slope*concentration + int, concentration = (OD - int/slope)
sample_concentrations <- (measurements[, 9:12] - int) / slope
sample_concentrations <- unlist(as.vector(c(concentrations[1:5], sample_concentrations)))
sample_concentrations

# wrangle and format concentrations
setbp1BCA_data <- rbind(measurements[, c(1:5, 9:12)], sample_concentrations)

setbp1BCA_data <- as.data.frame(t(setbp1BCA_data))
setbp1BCA_data$sample_ID <- c(
  "standard", "standard", "standard", "standard",
  "standard", "cerebral cortex het",
  "cerebral cortex WT", "kidney het", "kidney WT"
)
colnames(setbp1BCA_data) <- c("OD", "concentration", "sample_ID")

# plot data
ggplot(data = setbp1BCA_data) +
  geom_point(
    aes(
      x = concentration,
      y = OD,
      color = sample_ID
    ),
    position = "jitter"
  ) +
  theme_bw() +
  xlab("Concentration") +
  ylab("Optical Density (OD) at 562 nm")

# plot data with eq
png(here("results/ELISA/setbp1BCA_quantification.png"),
  width = 1000, height = 1000
)
ggscatter(
  data = setbp1BCA_data,
  x = "concentration",
  y = "OD",
  color = "sample_ID", size = 3,
  shape = "sample_ID",
  add = "reg.line", add.params = list(color = "black"),
  palette = get_palette(c("#644DD4", "#5F9EFD", "#FE6100", "#FFB000", "black"), 5),
  ggtheme = theme_bw(),
  xlab = "Concentration",
  ylab = "Optical Density (OD) at 562 nm",
  title = "Setbp1 BCA Quantification"
) +
  stat_regline_equation(label.x = 3, label.y = 1.2)
dev.off()
```

```{r plotting-individual-BCAreps}
measurements <- BCA[, c(2:6, 10:13)] - blank

# add concentration
concentrations <- c(25, 125, 250, 500, 750)

standards <- rbind(measurements[, 1:5], concentrations)
row.names(standards) <- c("rep1", "rep2", "rep3", "concentration")

# fit curve and get formula for standards
BCA_repdata <- as.data.frame(t(standards))
BCA_repdata <- pivot_longer(BCA_repdata, values_to = "OD", cols = 1:3)

ggplot(data = BCA_repdata) +
  geom_point(aes(x = concentration, y = OD)) +
  theme_bw() +
  xlab("Concentration") +
  ylab("Optical Density (OD) at 562 nm") +
  geom_smooth(aes(x = concentration, y = OD),
    method = lm,
    se = FALSE,
    formula = y ~ x
  )

# extract intercept and slope from line
line <- lm(BCA_repdata$OD ~ BCA_repdata$concentration)
int <- line$coefficients[1]
int # 0.3657495 0.2909356
slope <- line$coefficients[2]
slope # 0.0006567205

# using formula for line, determine unknown concentration from kidney_rep and cortex_rep (formula: OD  = slope*concentration + int, concentration = (OD - int/slope)
sample_concentrations <- (measurements[, 6:9] - int) / slope
sample_concentrations <- unlist(as.vector(sample_concentrations))
sample_concentrations

# wrangle and format concentrations
sample_ID <- c("cortex_het", "cortex_het", "cortex_het", "cortex_wt", "cortex_wt", "cortex_wt", "kidney_het", "kidney_het", "kidney_het", "kidney_wt", "kidney_wt", "kidney_wt")
rep_conc_BCA <- as.data.frame(cbind(sample_ID, as.numeric(sample_concentrations)))

# plot data
png(here("results/ELISA/setbp1BCA_quantification_boxplot.png"),
  width = 1000, height = 1000
)
ggplot(data = rep_conc_BCA, aes(
  x = sample_ID,
  y = sample_concentrations,
  fill = sample_ID
)) +
  scale_fill_manual(values = c(
    "kidney_wt" = "#FE6100",
    "kidney_het" = "#FFB000",
    "cortex_het" = "#5F9EFD",
    "cortex_wt" = "#644DD4"
  )) +
  geom_boxplot() +
  geom_point() +
  theme_bw() +
  xlab("Group") +
  ylab("Concentration") +
  stat_boxplot(
    geom = "errorbar", # Boxplot with error bars
    width = 0.2
  ) +
  ggtitle("Setbp1 Protein Concentration BCA")
dev.off()
```

Determining total proportion of total protein that Setbp1 is:
*ELISA quantifies Setbp1 protein concentration in ng/mL
*BCA quantifies total protein in ug/mL

Step 1: Convert concentrations to the same units. To convert ng/mL to ug/mL, divide the ELISA concentrations for ng/mL by 1000
```{r convert-units}
ng_concentrations <- setbp1ELISA_data[7:10, 2:3] # subset Setbp1 samples
ng_concentrations$ELISA_ug_mL <- ng_concentrations$concentration / 1000 # convert units
ELISA_concentrations <- ng_concentrations[, 2:3] # save in new df for merging
```

Step 2: Compare ELISA in ug/mL to BCA in ug/mL to calculate a proportion
```{r calc-prop}
BCA_concentrations <- setbp1BCA_data[6:9, 2:3] # subset Setbp1 samples
colnames(BCA_concentrations) <- c("BCA_ug_mL", "sample_ID") # rename cols to make more clear for merging

concentrations_ug_mL <- merge(BCA_concentrations, ELISA_concentrations)
concentrations_ug_mL$Setbp1_prop <- concentrations_ug_mL$ELISA_ug_mL / concentrations_ug_mL$BCA_ug_mL # dividing ELISA concentration by total protein concentration
concentrations_ug_mL$Setbp1_percent <- concentrations_ug_mL$Setbp1_prop * 100 # converting to a percent

ggplot(concentrations_ug_mL, aes(y = Setbp1_percent, x = sample_ID, fill = sample_ID)) +
  scale_fill_manual(values = c(
    "kidney_wt" = "#FE6100",
    "kidney_het" = "#FFB000",
    "cortex_het" = "#5F9EFD",
    "cortex_wt" = "#644DD4"
  )) +
  geom_bar(stat = "identity") +
  theme_bw() +
  geom_signif(
    comparisons = list(
      c("cerebral cortex WT", "cerebral cortex het"),
      c("kidney WT", "kidney het")
    ),
    map_signif_level = TRUE
  )
```

Step 1: Convert concentrations to the same units. To convert ng/mL to ug/mL, divide the ELISA concentrations for ng/mL by 1000
```{r convert-units}
ng_concentrations <- setbp1ELISA_data[7:10, 2:3] # subset Setbp1 samples
ng_concentrations$ELISA_ug_mL <- ng_concentrations$concentration / 1000 # convert units
ELISA_concentrations <- ng_concentrations[, 2:3] # save in new df for merging
```

Step 2: Compare ELISA in ug/mL to BCA in ug/mL to calculate a proportion
```{r calc-prop}
BCA_concentrations <- setbp1BCA_data[6:9, 2:3] # subset Setbp1 samples
colnames(BCA_concentrations) <- c("BCA_ug_mL", "sample_ID") # rename cols to make more clear for merging

concentrations_ug_mL <- merge(BCA_concentrations, ELISA_concentrations)
concentrations_ug_mL$Setbp1_prop <- concentrations_ug_mL$ELISA_ug_mL / concentrations_ug_mL$BCA_ug_mL # dividing ELISA concentration by total protein concentration
concentrations_ug_mL$Setbp1_percent <- concentrations_ug_mL$Setbp1_prop * 100 # converting to a percent

png(here("results/ELISA/setbp1_percent_total_protein_barplot.png"),
  width = 1000, height = 1000
)
ggplot(concentrations_ug_mL, aes(y = Setbp1_percent, x = sample_ID, fill = sample_ID)) +
  geom_bar(stat = "identity") +
  theme_bw() +
  scale_fill_manual(values = c(
    "kidney_wt" = "#FE6100",
    "kidney_het" = "#FFB000",
    "cortex_het" = "#5F9EFD",
    "cortex_wt" = "#644DD4"
  )) +
  ggtitle("Proportion of Total Protein") +
  ylab("percent of Setb1p protein") +
  xlab("Group")
dev.off()
```

# plot proportion of total protein of replicates

Step 1: Convert concentrations to the same units. To convert ng/mL to ug/mL, divide the ELISA concentrations for ng/mL by 1000
```{r convert-units}
rep_conc_ELISA$ELISA_ug_mL <- as.numeric(rep_conc_ELISA$V2) / 1000 # convert units
ELISA_concentrations_reps <- rep_conc_ELISA[, c(1, 3)] # save in new df for merging
```

Step 2: Compare ELISA in ug/mL to BCA in ug/mL to calculate a proportion
```{r calculate-prop}
colnames(rep_conc_BCA) <- c("sample_ID", "BCA_ug_mL") # rename cols to make more clear for merging

concentrations_ug_mL <- cbind(rep_conc_BCA, ELISA_concentrations_reps) %>% .[, -3]
concentrations_ug_mL$Setbp1_prop <- concentrations_ug_mL$ELISA_ug_mL / as.numeric(concentrations_ug_mL$BCA_ug_mL) # dividing ELISA concentration by total protein concentration
concentrations_ug_mL$Setbp1_percent <- concentrations_ug_mL$Setbp1_prop * 100 # converting to a percent

# caluclate a p-value and plot with singificance indicated
compare_means(Setbp1_percent ~ sample_ID, concentrations_ug_mL, method = "t.test")
my_comparisons <- list(
  c("cortex_het", "cortex_wt"),
  c("kidney_het", "kidney_wt"), c("cortex_het", "kidney_het")
)

png(here("results/ELISA/setbp1_percent_total_protein_boxplot.png"),
  width = 1000, height = 1000
)
ggplot(data = concentrations_ug_mL, aes(
  x = sample_ID,
  y = Setbp1_percent,
  fill = sample_ID,
  add = "jitter",
  legend = "none"
)) +
  scale_fill_manual(values = c(
    "kidney_wt" = "#FE6100",
    "kidney_het" = "#FFB000",
    "cortex_het" = "#5F9EFD",
    "cortex_wt" = "#644DD4"
  )) +
  geom_boxplot() +
  geom_point() +
  theme_bw() +
  xlab("Group") +
  ylab("Percent of total protein") +
  stat_boxplot(
    geom = "errorbar", # Boxplot with error bars
    width = 0.2
  ) +
  ggtitle("Setbp1 proportion of Total Protein with t.test") +
  stat_compare_means(
    comparisons = my_comparisons,
    label = "p.signif",
    method = "t.test",
  )
dev.off()
```

```{r}
proc.time()
```
   user  system elapsed 
 10.119   2.389 139.639 
 
 
```{r}
sessionInfo()
```
R version 4.1.3 (2022-03-10)
Platform: x86_64-pc-linux-gnu (64-bit)
Running under: Ubuntu 20.04.5 LTS

Matrix products: default
BLAS/LAPACK: /usr/lib/x86_64-linux-gnu/openblas-pthread/libopenblasp-r0.3.8.so

locale:
[1] C

attached base packages:
[1] stats     graphics  grDevices utils     datasets  methods   base     

other attached packages:
[1] tidyr_1.3.0   lintr_3.0.2   styler_1.9.0  ggpubr_0.6.0  ggplot2_3.4.1 readr_2.1.4   here_1.0.1   

loaded via a namespace (and not attached):
 [1] tidyselect_1.2.0  xfun_0.37         remotes_2.4.2     purrr_1.0.1       lattice_0.20-45   splines_4.1.3     carData_3.0-5     colorspace_2.1-0 
 [9] vctrs_0.5.2       generics_0.1.3    mgcv_1.8-39       utf8_1.2.3        rlang_1.0.6       R.oo_1.25.0       pillar_1.8.1      glue_1.6.2       
[17] withr_2.5.0       R.utils_2.12.2    bit64_4.0.5       lifecycle_1.0.3   R.cache_0.16.0    munsell_0.5.0     ggsignif_0.6.4    gtable_0.3.1     
[25] R.methodsS3_1.8.2 labeling_0.4.2    knitr_1.42        tzdb_0.3.0        callr_3.7.3       ps_1.7.2          rex_1.2.1         parallel_4.1.3   
[33] fansi_1.0.4       broom_1.0.3       polynom_1.4-1     backports_1.4.1   scales_1.2.1      desc_1.4.2        vroom_1.6.1       abind_1.4-5      
[41] farver_2.1.1      bit_4.0.5         hms_1.1.2         digest_0.6.31     rstatix_0.7.2     processx_3.8.0    dplyr_1.1.0       grid_4.1.3       
[49] rprojroot_2.0.3   cli_3.6.0         tools_4.1.3       magrittr_2.0.3    lazyeval_0.2.2    tibble_3.1.8      crayon_1.5.2      car_3.1-1        
[57] pkgconfig_2.0.3   Matrix_1.5-3      ellipsis_0.3.2    xml2_1.3.3        cyclocomp_1.1.0   rstudioapi_0.14   R6_2.5.1          nlme_3.1-155  

```{r lint-file-style}
# run style
style_file(here("src/ELISA_scripts/230315_ELISA_BCA.Rmd"))
# lintr was run as well
```


