---
title: "Setbp1_MarkersCortex_04"
author: "Jordan Whitlock"
date: '2022-09-07'
output:     
  html_document:
      toc: true
      toc_depth: 2
      toc_float: true
---
This analysis comes after 'Setbp1_Clustering_03.Rmd' and precedes 'Setbp1_CellTypesCortex_05.Rmd'

# *Single-nuclei analysis: FindMarkers Cerebral Cortex*
* Loading libraries
* Selecting resolution
* Finding all markers for cerebral cortex

## Code: 

Accompanying dry lab notebook for this analysis is here: https://docs.google.com/document/d/1FO_Je9oxbBFF9S_k-8tVF2t4ba88O4C77sw1v7DIFec/edit?usp=sharing

### Setting up libraries:

```{r}
# loading packages
library(Seurat)
library(here)
library(styler)
library(lintr)
set.seed(2178)

# loading in the objects
load(here("data/cerebral_leiden.Rdata"))
```

```{r}
# setting selected resolution
select_resolution <- "RNA_snn_res.1"
cerebral_int <- SetIdent(cerebral_int, value = select_resolution)
```

```{r}
# compute highly differential expression for brain
cerebral_markers_genes <- FindAllMarkers(cerebral_int,
  log2FC.threshold = 0.2, test.use = "wilcox",
  min.pct = 0.1, min.diff.pct = 0.2, only.pos = TRUE, max.cells.per.ident = 50,
  assay = "RNA"
)
save(cerebral_markers_genes, file = here("results/seurat/Markers_04/cerebral_markers.Rdata"))
```

```{r}
proc.time()
```
    user   system  elapsed 
 315.445   94.581 1270.240 

```{r}
sessionInfo()
```

R version 4.1.3 (2022-03-10)
Platform: x86_64-pc-linux-gnu (64-bit)
Running under: Ubuntu 20.04.5 LTS

Matrix products: default
BLAS/LAPACK: /usr/lib/x86_64-linux-gnu/openblas-pthread/libopenblasp-r0.3.8.so

locale:
[1] C

attached base packages:
[1] stats     graphics  grDevices utils     datasets  methods   base     

other attached packages:
[1] lintr_3.0.2        styler_1.9.0       here_1.0.1         SeuratObject_4.1.3 Seurat_4.3.0      

loaded via a namespace (and not attached):
  [1] Rtsne_0.16             colorspace_2.1-0       deldir_1.0-6           ellipsis_0.3.2         ggridges_0.5.4         rprojroot_2.0.3        rstudioapi_0.14       
  [8] spatstat.data_3.0-0    leiden_0.4.3           listenv_0.9.0          remotes_2.4.2          ggrepel_0.9.3          fansi_1.0.4            xml2_1.3.3            
 [15] codetools_0.2-18       splines_4.1.3          R.methodsS3_1.8.2      knitr_1.42             polyclip_1.10-4        jsonlite_1.8.4         ica_1.0-3             
 [22] cluster_2.1.2          png_0.1-8              R.oo_1.25.0            uwot_0.1.14            shiny_1.7.4            sctransform_0.3.5      spatstat.sparse_3.0-0 
 [29] compiler_4.1.3         httr_1.4.5             Matrix_1.5-3           fastmap_1.1.1          lazyeval_0.2.2         limma_3.50.3           cli_3.6.0             
 [36] later_1.3.0            htmltools_0.5.4        tools_4.1.3            igraph_1.4.1           gtable_0.3.1           glue_1.6.2             RANN_2.6.1            
 [43] reshape2_1.4.4         dplyr_1.1.0            Rcpp_1.0.10            scattermore_0.8        vctrs_0.5.2            spatstat.explore_3.0-6 nlme_3.1-155          
 [50] progressr_0.13.0       lmtest_0.9-40          spatstat.random_3.1-3  xfun_0.37              stringr_1.5.0          ps_1.7.2               globals_0.16.2        
 [57] mime_0.12              miniUI_0.1.1.1         lifecycle_1.0.3        irlba_2.3.5.1          goftest_1.2-3          future_1.31.0          MASS_7.3-55           
 [64] zoo_1.8-11             scales_1.2.1           promises_1.2.0.1       spatstat.utils_3.0-1   rex_1.2.1              parallel_4.1.3         RColorBrewer_1.1-3    
 [71] reticulate_1.28        pbapply_1.7-0          gridExtra_2.3          ggplot2_3.4.1          stringi_1.7.12         desc_1.4.2             cyclocomp_1.1.0       
 [78] rlang_1.0.6            pkgconfig_2.0.3        matrixStats_0.63.0     lattice_0.20-45        ROCR_1.0-11            purrr_1.0.1            tensor_1.5            
 [85] patchwork_1.1.2        htmlwidgets_1.6.1      processx_3.8.0         cowplot_1.1.1          tidyselect_1.2.0       parallelly_1.34.0      RcppAnnoy_0.0.20      
 [92] plyr_1.8.8             magrittr_2.0.3         R6_2.5.1               generics_0.1.3         withr_2.5.0            pillar_1.8.1           fitdistrplus_1.1-8    
 [99] survival_3.3-1         abind_1.4-5            sp_1.6-0               tibble_3.1.8           future.apply_1.10.0    crayon_1.5.2           KernSmooth_2.23-20    
[106] utf8_1.2.3             spatstat.geom_3.0-6    plotly_4.10.1          grid_4.1.3             data.table_1.14.8      callr_3.7.3            digest_0.6.31         
[113] xtable_1.8-4           R.cache_0.16.0         tidyr_1.3.0            httpuv_1.6.9           R.utils_2.12.2         munsell_0.5.0          viridisLite_0.4.1     


```{r}
# run style
style_file(here("src/seurat_scripts/Setbp1_MarkersCortex_04.Rmd"))
# lintr was run as well
```
