---
title: "Setbp1_CellTypesKidney_SoupX_06"
author: "Jordan Whitlock"
date: '2023-04-12'
output: html_document
---
# *Single-nuclei analysis: Identify Cell Types Kidney for SoupX*
* Loading libraries
* Selecting resolution
* Investigating top markers for Kidney Cortex
* ID cell types for SoupX

## Code: 

This analysis follows 'Setbp1_MarkersKidney_05.Rmd' and precedes "Setbp1_AmbientRNA_SoupX_02.Rmd"

#### load in libraries:
```{r}
set.seed(2178)
library(Seurat)
library(ggplot2)
library(styler)
library(here)
library(lintr)
library(magrittr)
library(dplyr)
```

#### load in markers:
```{r}
load(here("results/seurat/Markers_04/kidney_markers.Rdata"))
```

*also be sure that the previously generated clustered objects at resolution 1.7 are here if not loaded already, run below chunk:
```{r}
load(here("data/kidney_leiden.Rdata"))
select_resolution <- "RNA_snn_res.1.7"
kidney_int <- SetIdent(kidney_int, value = select_resolution)
```

#### identify the top25 markers for kidney :
```{r}
# top 25 kidney
kidney_top25 <- kidney_markers_genes %>%
  group_by(cluster) %>%
  top_n(-25, p_val_adj)

save(kidney_top25, file = here("results/seurat/CellTypes_06/kidney_top25.Rdata"))
```

#### investigate top markers and assign for kidney:
see notes on resources used to do this here as well as this paper: https://docs.google.com/document/d/1FO_Je9oxbBFF9S_k-8tVF2t4ba88O4C77sw1v7DIFec/edit?usp=sharing, https://www.science.org/doi/pdf/10.1126/science.aar2131)


##### Pericytes
```{r pericyte-features}
png(
  filename = here("results/figures/pericyte_markers.png"),
  width = 1000,
  height = 1000
)
FeaturePlot(kidney_int,
  features = c("Pdgfrb"),
  reduction = "umap_harmony",
  label = TRUE
) 
dev.off()
```

##### Endothelial
```{r endo-features}
png(
  filename = here("results/figures/endothelial_markers.png"),
  width = 1000,
  height = 1000
)
FeaturePlot(kidney_int,
  features = c("Kdr", "Ptprb"),
  reduction = "umap_harmony",
  label = TRUE
)
dev.off()
```

##### Proximal Tubule
```{r PT-features}
png(
  filename = here("results/figures/proximaltubule_markers.png"),
  width = 1000,
  height = 1000
)
FeaturePlot(kidney_int,
  features = c("Slc34a1", "Slc13a3"),
  reduction = "umap_harmony",
  label = TRUE
)
dev.off()
```

##### Proximal Straight Tubule (PST)
```{r PST-features}
png(
  filename = here("results/figures/PST_markers.png"),
  width = 1000,
  height = 1000
)
FeaturePlot(kidney_int,
  features = c("Slc22a7", "Atp11a"),
  reduction = "umap_harmony",
  label = TRUE
)
dev.off()
```

##### PCT-S1
```{r PCTS1-features}
png(
  filename = here("results/figures/PCTS1_markers.png"),
  width = 1000,
  height = 1000
)
FeaturePlot(kidney_int,
  features = c("Slc5a2", "Slc5a12"),
  reduction = "umap_harmony",
  label = TRUE
)
dev.off()
```

##### PCT-S2
```{r PCTS2-features}
png(
  filename = here("results/figures/PCTS2_markers.png"),
  width = 1000,
  height = 1000
)
FeaturePlot(kidney_int,
  features = c("Fxyd2"),
  reduction = "umap_harmony",
  label = TRUE
)
dev.off()
```

##### LOH
```{r LOH-features}
png(
  filename = here("results/figures/LOH_markers.png"),
  width = 1000,
  height = 1000
)
FeaturePlot(kidney_int,
  features = c("Slc12a1"),
  reduction = "umap_harmony",
  label = TRUE
)
dev.off()
```

##### CD-PC
```{r CDPC-features}
png(
  filename = here("results/figures/CDPC_markers.png"),
  width = 1000,
  height = 1000
)
FeaturePlot(kidney_int,
  features = c("Aqp2", "Hsd11b2"),
  reduction = "umap_harmony",
  label = TRUE
)
dev.off()
```

##### DCT
```{r DCT-features}
png(
  filename = here("results/figures/DCT_markers.png"),
  width = 1000,
  height = 1000
)
FeaturePlot(kidney_int,
  features = c("Slc12a3"),
  reduction = "umap_harmony",
  label = TRUE
)
dev.off()
```

##### Macrophages
```{r macro-features}
png(
  filename = here("results/figures/macrophage_markers.png"),
  width = 1000,
  height = 1000
)
FeaturePlot(kidney_int,
  features = c("Ptprc", "Runx1"),
  reduction = "umap_harmony",
  label = TRUE
)
dev.off()
```

##### DLH
https://esbl.nhlbi.nih.gov/Databases/AllTranscripts/
```{r DLH-features}
png(
  filename = here("results/figures/DLH_markers.png"),
  width = 1000,
  height = 1000
)
FeaturePlot(kidney_int,
  features = c("Bst1", "Akr1b3"),
  reduction = "umap_harmony",
  label = TRUE
)
dev.off()
```

##### CD-IC-A
```{r}
png(
  filename = here("results/figures/CDICA_markers.png"),
  width = 1000,
  height = 1000
)
FeaturePlot(kidney_int,
  features = c("Aqp6"),
  reduction = "umap_harmony",
  label = TRUE
)
dev.off()
```

##### CD-IC-B
```{r CDICB-features}
png(
  filename = here("results/figures/CDICB_markers.png"),
  width = 1000,
  height = 1000
)
FeaturePlot(kidney_int,
  features = c("Hmx2"),
  reduction = "umap_harmony",
  label = TRUE
)
dev.off()
```

##### CD-IC
```{r CDIC-features}
png(
  filename = here("results/figures/CDIC_markers.png"),
  width = 1000,
  height = 1000
)
FeaturePlot(kidney_int,
  features = c("Atp6v1g3", "Atp6v0d2"),
  reduction = "umap_harmony",
  label = TRUE
)
dev.off()
```

##### B cell
```{r bcell-features}
png(
  filename = here("results/figures/bcell_markers.png"),
  width = 1000,
  height = 1000
)
FeaturePlot(kidney_int,
  features = c("Cd79b", "Bank1"),
  reduction = "umap_harmony",
  label = TRUE
)
dev.off()
```

##### podocytes
https://pubmed.ncbi.nlm.nih.gov/11912180/, https://www.sciencedirect.com/science/article/pii/S0002944013005853
```{r podo-features}
png(
  filename = here("results/figures/podocyte_markers.png"),
  width = 1000,
  height = 1000
)
FeaturePlot(kidney_int,
  features = c("Nphs1", "Nphs2", "Wt1"),
  reduction = "umap_harmony",
  label = TRUE
)
dev.off()
```


##### Fibroblasts
```{r fibro-features}
png(
  filename = here("results/figures/fibroblast_markers.png"),
  width = 1000,
  height = 1000
)
FeaturePlot(kidney_int,
  features = c("Pdgfra"),
  reduction = "umap_harmony",
  label = TRUE
)
dev.off()
```

##### Smooth Muscle Cells
```{r cluster-smc-features}
png(
  file = here("results/figures/smooth_muscle_cells.png"),
  width = 1000,
  height = 1000
)
FeaturePlot(kidney_int, features = c("Atp1a2"), reduction = "umap_harmony", label = TRUE)
dev.off()
```

##### Plotting all markers

For some reason the levels were not in numerical order, so I re-ordered them. 
```{r reordering-levels}
levels(kidney_int)

levels(kidney_int) <- c("1", "2", "3", "4", "5", "6", "7", "8", "9", "10",
                        "11", "12", "13", "14", "15", "16", "17", "18", "19",
                        "20", "21", "22", "23", "24", "25", "26", "27", "28",
                        "29", "30", "31", "32", "33", "34", "35", "36", "37",
                        "38", "39", "40", "41", "42", "43", "44", "45", "46",
                        "47", "48", "49", "50", "51", "52", "53")

levels(kidney_int)
```


```{r plotting-final-features}
kidney_features <- c("Atp1a2", "Pdgfra", "Nphs1", "Nphs2", "Wt1", "Cd79b",
                     "Bank1", "Hmx2", "Aqp6", "Atp6v1g3", "Atp6v0d2", "Bst1",
                     "Akr1b3", "Runx1", "Ptprc", "Slc12a3", "Aqp2", "Hsd11b2",
                     "Slc12a1", "Fxyd2", "Slc5a2", "Slc5a12", "Slc22a7",
                     "Atp11a", "Slc34a1", "Slc13a3", "Kdr", "Ptprb", "Pdgfrb")

png(
  filename = here("results/figures/setbp1_kidney_markers_dotplot.png"),
  width = 1000,
  height = 1000
)
DotPlot(kidney_int, kidney_features, assay = "RNA") +
  theme_bw() + theme(legend.position = "left") +
  theme(axis.title.x = element_blank()) +
  theme(panel.grid = element_blank()) +
  theme(axis.text.x = element_text(angle = 90)) +
  coord_flip()
dev.off()
```

## assigning kidney cell markers
```{r assigning-markers}
kidney_int_celltypes <- SetIdent(kidney_int, value = select_resolution)
kidney_int_celltypes <- RenameIdents(kidney_int,
  "1" = "PCTS2",
  "2" = "PCTS1",
  "3" = "PST",
  "4" = "fibroblasts",
  "5" = "LOH",
  "6" = "LOH",
  "7" = "endothelial",
  "8" = "proximal_tubule",
  "9" = "PCTS1",
  "10" = "endothelial",
  "11" = "endothelial",
  "12" = "DCT",
  "13" = "PCTS1",
  "14" = "PCTS2",
  "15" = "proximal_tubule",
  "16" = "PCTS1",
  "17" = "PST",
  "18" = "PCTS2",
  "19" = "DCT",
  "20" = "fibroblasts",
  "21" = "PCTS1",
  "22" = "PCTS1",
  "23" = "CDIC_typeA",
  "24" = "CDPC",
  "25" = "PST",
  "26" = "DLH",
  "27" = "CDIC_typeB",
  "28" = "CDPC",
  "29" = "macrophages",
  "30" = "DLH",
  "31" = "PCTS2",
  "32" = "proximal_tubule",
  "33" = "CDPC",
  "34" = "CDPC",
  "35" = "proximal_tubule",
  "36" = "PST",
  "37" = "podocytes",
  "38" = "fibroblasts",
  "39" = "fibroblasts",
  "40" = "pericytes",
  "41" = "PCTS1",
  "42" = "endothelial",
  "43" = "DLH",
  "44" = "macrophages",
  "45" = "podocytes",
  "46" = "B_cells",
  "47" = "LOH",
  "48" = "PCTS2",
  "49" = "DCT",
  "50" = "macrophages",
  "51" = "LOH",
  "52" = "PST",
  "53" = "smooth_muscle_cells"
)
kidney_int_celltypes <- AddMetaData(object = kidney_int_celltypes, as.vector(kidney_int_celltypes@active.ident), col.name = "cell_type")

save(kidney_int_celltypes, file = here("data/kidney_integrated_celltypes.Rdata"))
```

```{r}
png(
  filename = here("results/figures/kidney_setbp1_UMAP_preSoup.png"),
  width = 1200,
  height = 600
)
DimPlot(kidney_int_celltypes, reduction = "umap_harmony",
        group.by = "cell_type",
        cols = c("B_cells" = "#FFBE1D",
                 "CDIC_typeA" = "#4E3801",
                 "CDIC_typeB" = "#EEDF37",
                 "CDPC" = "#F03F00",
                 "DCT" = "#6D0404",
                 "DLH" = "#AA6320",
                 "endothelial" = "#CF9400",
                 "fibroblasts" = "#FFC8C4",
                 "LOH" = "#AA937E",
                 "macrophages" = "#C70F0F",
                 "PCTS1" = "#A85A5A",
                 "podocytes" = "#BD085E",
                 "PST" = "#948802",
                 "pericytes" = "#FFE196",
                 "PCTS2" = "#FF7600",
                 "proximal_tubule" = "#F8766D",
                 "smooth_muscle_cells" = "#FFE196"))
dev.off()
```

```{r}
proc.time()
```
    user   system  elapsed 
  95.904   25.687 3545.222 
  
```{r}
sessionInfo()
```
R version 4.1.3 (2022-03-10)
Platform: x86_64-pc-linux-gnu (64-bit)
Running under: Ubuntu 20.04.5 LTS

Matrix products: default
BLAS/LAPACK: /usr/lib/x86_64-linux-gnu/openblas-pthread/libopenblasp-r0.3.8.so

locale:
[1] C

attached base packages:
[1] stats     graphics  grDevices utils     datasets  methods   base     

other attached packages:
[1] dplyr_1.1.0        magrittr_2.0.3     lintr_3.0.2        here_1.0.1         styler_1.9.0       ggplot2_3.4.1      SeuratObject_4.1.3 Seurat_4.3.0      

loaded via a namespace (and not attached):
  [1] Rtsne_0.16             colorspace_2.1-0       deldir_1.0-6           ellipsis_0.3.2         ggridges_0.5.4         rprojroot_2.0.3       
  [7] rstudioapi_0.14        spatstat.data_3.0-0    farver_2.1.1           leiden_0.4.3           listenv_0.9.0          remotes_2.4.2         
 [13] ggrepel_0.9.3          fansi_1.0.4            xml2_1.3.3             codetools_0.2-18       splines_4.1.3          R.methodsS3_1.8.2     
 [19] knitr_1.42             polyclip_1.10-4        jsonlite_1.8.4         ica_1.0-3              cluster_2.1.2          png_0.1-8             
 [25] R.oo_1.25.0            uwot_0.1.14            shiny_1.7.4            sctransform_0.3.5      spatstat.sparse_3.0-0  compiler_4.1.3        
 [31] httr_1.4.5             Matrix_1.5-3           fastmap_1.1.1          lazyeval_0.2.2         cli_3.6.0              later_1.3.0           
 [37] htmltools_0.5.4        tools_4.1.3            igraph_1.4.1           gtable_0.3.1           glue_1.6.2             RANN_2.6.1            
 [43] reshape2_1.4.4         Rcpp_1.0.10            scattermore_0.8        vctrs_0.5.2            spatstat.explore_3.0-6 nlme_3.1-155          
 [49] progressr_0.13.0       lmtest_0.9-40          spatstat.random_3.1-3  xfun_0.37              stringr_1.5.0          ps_1.7.2              
 [55] globals_0.16.2         mime_0.12              miniUI_0.1.1.1         lifecycle_1.0.3        irlba_2.3.5.1          goftest_1.2-3         
 [61] future_1.31.0          MASS_7.3-55            zoo_1.8-11             scales_1.2.1           promises_1.2.0.1       spatstat.utils_3.0-1  
 [67] rex_1.2.1              parallel_4.1.3         RColorBrewer_1.1-3     reticulate_1.28        pbapply_1.7-0          gridExtra_2.3         
 [73] stringi_1.7.12         desc_1.4.2             cyclocomp_1.1.0        rlang_1.0.6            pkgconfig_2.0.3        matrixStats_0.63.0    
 [79] lattice_0.20-45        ROCR_1.0-11            purrr_1.0.1            tensor_1.5             labeling_0.4.2         patchwork_1.1.2       
 [85] htmlwidgets_1.6.1      processx_3.8.0         cowplot_1.1.1          tidyselect_1.2.0       parallelly_1.34.0      RcppAnnoy_0.0.20      
 [91] plyr_1.8.8             R6_2.5.1               generics_0.1.3         pillar_1.8.1           withr_2.5.0            fitdistrplus_1.1-8    
 [97] survival_3.3-1         abind_1.4-5            sp_1.6-0               tibble_3.1.8           future.apply_1.10.0    crayon_1.5.2          
[103] KernSmooth_2.23-20     utf8_1.2.3             spatstat.geom_3.0-6    plotly_4.10.1          grid_4.1.3             data.table_1.14.8     
[109] callr_3.7.3            digest_0.6.31          xtable_1.8-4           R.cache_0.16.0         tidyr_1.3.0            httpuv_1.6.9          
[115] R.utils_2.12.2         munsell_0.5.0          viridisLite_0.4.1     

