---
title: "Setbp1_DGEcortex_06"
author: "Jordan Whitlock"
date: '2022-09-08'
output:     
  html_document:
      toc: true
      toc_depth: 2
      toc_float: true
---

This analysis comes after 'Setbp1_CellTypesCortex_05.Rmd' and precedes 'Setbp1_PathwayAnalysisCortexKidney_07.Rmd'

# *Single-nuclei analysis: DGE cortex*

## Code: 

#### Loading in libraries:
```{r}
library(Seurat)
library(ggplot2)
library(tidyr)
library(dplyr)
library(ComplexHeatmap)
library(circlize)
library(here)
library(styler)
library(lintr)
source(here("src/functions/functions.R"))
set.seed(2178)
```

#### Load in object:
```{r}
load(here("data/setbp1_cerebralintcelltypes.Rdata"))
```

#### checking to see if Setbp1 is even in the seurat object:

It is present in the data, but appears very similar across conditions.
```{r}
# visualize with FeaturePlot
png(
  file = here("results/seurat/DGE_06/Setbp1_expression_cortex.png"),
  width = 1000,
  height = 1000
)
plot <- FeaturePlot(cerebral_int_celltypes, feature = "Setbp1", reduction = "umap_harmony", split.by = "type", label = TRUE)
plot
dev.off()
```

#### split violin plotting of Setbp1 across cell types:
```{r}
# grabbing the Setbp1 expression data and moving from S4 to a dataframe
setbp1_exp <- FetchData(cerebral_int_celltypes, vars = c("Setbp1", "cell_type", "type"))
colnames(setbp1_exp) <- c("Expression", "cell_type", "type")

png(
  filename = here("results/figures/cortex_setbp1_expression.png"),
  width = 1500,
  height = 1000
)
ggplot(setbp1_exp, aes(cell_type, Expression, fill = type)) +
  geom_split_violin() +
  theme_bw() +
  scale_fill_manual(values = c("control" = "#644DD4", "heterozygous" = "#5F9EFD")) +
  ggtitle("Setbp1 Expression Across Cerebral Cortex")
dev.off()
```


#### Calculating DGE by cell type: log2FC of 0.1, and plotting etc

---excitatory neurons---
```{r}
# subset seurat object for each cell type
exneuro <- subset(x = cerebral_int_celltypes, idents = "excitatory_neurons")
Idents(exneuro) <- exneuro@meta.data$type
excitatory_neurons_data <- FindAllMarkers(exneuro, logfc.threshold = 0.1, test.use = "wilcox", only.pos = FALSE, assay = "RNA")
head(excitatory_neurons_data, n = 5)
```

```{r}
excitatory_neurons_het <- excitatory_neurons_data[excitatory_neurons_data$p_val_adj < 0.05 & excitatory_neurons_data$avg_log2FC > 0.1 | excitatory_neurons_data$avg_log2FC < -0.1, ]
excitatory_neurons_het <- excitatory_neurons_het[excitatory_neurons_het$cluster == "heterozygous", ]

# add column to data to indicate if DE or not:
excitatory_neurons_het$diffexpressed <- "NO"
# if log2Foldchange > 0.1 and pvalue < 0.05, set as "UP"
excitatory_neurons_het$diffexpressed[excitatory_neurons_het$avg_log2FC > 0.1 & excitatory_neurons_het$p_val_adj < 0.05] <- "UP"
# if log2Foldchange < -0.1 and pvalue < 0.05, set as "DOWN"
excitatory_neurons_het$diffexpressed[excitatory_neurons_het$avg_log2FC < -0.1 & excitatory_neurons_het$p_val_adj < 0.05] <- "DOWN"
```


---inhibitory neurons ---

```{r}
# subset seurat object for each cell type
inneuro <- subset(x = cerebral_int_celltypes, idents = "inhibitory_neurons")
Idents(inneuro) <- inneuro@meta.data$type
inhibitory_neurons_data <- FindAllMarkers(inneuro, logfc.threshold = 0.1, test.use = "wilcox", only.pos = FALSE, assay = "RNA")
head(inhibitory_neurons_data, n = 5)
```

```{r}
inhibitory_neurons_het <- inhibitory_neurons_data[inhibitory_neurons_data$p_val_adj < 0.05 & inhibitory_neurons_data$avg_log2FC > 0.1 | inhibitory_neurons_data$avg_log2FC < -0.1, ]
inhibitory_neurons_het <- inhibitory_neurons_het[inhibitory_neurons_het$cluster == "heterozygous", ]

# add column to data to indicate if DE or not:
inhibitory_neurons_het$diffexpressed <- "NO"
# if log2Foldchange > 0.1 and pvalue < 0.05, set as "UP"
inhibitory_neurons_het$diffexpressed[inhibitory_neurons_het$avg_log2FC > 0.1 & inhibitory_neurons_het$p_val_adj < 0.05] <- "UP"
# if log2Foldchange < -0.1 and pvalue < 0.05, set as "DOWN"
inhibitory_neurons_het$diffexpressed[inhibitory_neurons_het$avg_log2FC < -0.1 & inhibitory_neurons_het$p_val_adj < 0.05] <- "DOWN"
```


---astrocytes---
```{r}
# subset seurat object for each cell type
astro <- subset(x = cerebral_int_celltypes, idents = "astrocytes")
Idents(astro) <- astro@meta.data$type
astrocytes_data <- FindAllMarkers(astro, logfc.threshold = 0.1, test.use = "wilcox", only.pos = FALSE, assay = "RNA")
head(astrocytes_data, n = 5)
```

```{r}
astrocytes_het <- astrocytes_data[astrocytes_data$p_val_adj < 0.05 & astrocytes_data$avg_log2FC > 0.1 | astrocytes_data$avg_log2FC < -0.1, ]
astrocytes_het <- astrocytes_het[astrocytes_het$cluster == "heterozygous", ]

# add column to data to indicate if DE or not:
astrocytes_het$diffexpressed <- "NO"
# if log2Foldchange > 0.1 and pvalue < 0.05, set as "UP"
astrocytes_het$diffexpressed[astrocytes_het$avg_log2FC > 0.1 & astrocytes_het$p_val_adj < 0.05] <- "UP"
# if log2Foldchange < -0.1 and pvalue < 0.05, set as "DOWN"
astrocytes_het$diffexpressed[astrocytes_het$avg_log2FC < -0.1 & astrocytes_het$p_val_adj < 0.05] <- "DOWN"
```


---microglia---
```{r}
# subset seurat object for each cell type
micro <- subset(x = cerebral_int_celltypes, idents = "microglia")
Idents(micro) <- micro@meta.data$type
microglia_data <- FindAllMarkers(micro, logfc.threshold = 0.1, test.use = "wilcox", only.pos = FALSE, assay = "RNA")
head(microglia_data, n = 5)
```

```{r}
microglia_het <- microglia_data[microglia_data$p_val_adj < 0.05 & microglia_data$avg_log2FC > 0.1 | microglia_data$avg_log2FC < -0.1, ]
microglia_het <- microglia_het[microglia_het$cluster == "heterozygous", ]

# add column to data to indicate if DE or not:
microglia_het$diffexpressed <- "NO"
# if log2Foldchange > 0.1 and pvalue < 0.05, set as "UP"
microglia_het$diffexpressed[microglia_het$avg_log2FC > 0.1 & microglia_het$p_val_adj < 0.05] <- "UP"
# if log2Foldchange < -0.1 and pvalue < 0.05, set as "DOWN"
microglia_het$diffexpressed[microglia_het$avg_log2FC < -0.1 & microglia_het$p_val_adj < 0.05] <- "DOWN"
```


---pericytes---
```{r}
# subset seurat object for each cell type
pericytes <- subset(x = cerebral_int_celltypes, idents = "pericytes")
Idents(pericytes) <- pericytes@meta.data$type
pericytes_data <- FindAllMarkers(pericytes, logfc.threshold = 0.1, test.use = "wilcox", only.pos = FALSE, assay = "RNA")
head(pericytes_data, n = 5)
```

```{r}
pericytes_het <- pericytes_data[pericytes_data$p_val_adj < 0.05 & pericytes_data$avg_log2FC > 0.1 | pericytes_data$avg_log2FC < -0.1, ]
pericytes_het <- pericytes_het[pericytes_het$cluster == "heterozygous", ]

# add column to data to indicate if DE or not:
pericytes_het$diffexpressed <- "NO"
# if log2Foldchange > 0.1 and pvalue < 0.05, set as "UP"
pericytes_het$diffexpressed[pericytes_het$avg_log2FC > 0.1 & pericytes_het$p_val_adj < 0.05] <- "UP"
# if log2Foldchange < -0.1 and pvalue < 0.05, set as "DOWN"
pericytes_het$diffexpressed[pericytes_het$avg_log2FC < -0.1 & pericytes_het$p_val_adj < 0.05] <- "DOWN"
```

---opcs---
```{r}
# subset seurat object for each cell type
opcs <- subset(x = cerebral_int_celltypes, idents = "opcs")
Idents(opcs) <- opcs@meta.data$type
opcs_data <- FindAllMarkers(opcs, logfc.threshold = 0.1, test.use = "wilcox", only.pos = FALSE, assay = "RNA")
head(opcs_data, n = 5)
```

```{r}
opcs_het <- opcs_data[opcs_data$p_val_adj < 0.05 & opcs_data$avg_log2FC > 0.1 | opcs_data$avg_log2FC < -0.1, ]
opcs_het <- opcs_het[opcs_het$cluster == "heterozygous", ]

# add column to data to indicate if DE or not:
opcs_het$diffexpressed <- "NO"
# if log2Foldchange > 0.1 and pvalue < 0.05, set as "UP"
opcs_het$diffexpressed[opcs_het$avg_log2FC > 0.1 & opcs_het$p_val_adj < 0.05] <- "UP"
# if log2Foldchange < -0.1 and pvalue < 0.05, set as "DOWN"
opcs_het$diffexpressed[opcs_het$avg_log2FC < -0.1 & opcs_het$p_val_adj < 0.05] <- "DOWN"
```


---oligodendrocytes---
```{r}
# subset seurat object for each cell type
oligo <- subset(x = cerebral_int_celltypes, idents = "oligodendrocytes")
Idents(oligo) <- oligo@meta.data$type
oligodendrocytes_data <- FindAllMarkers(oligo, logfc.threshold = 0.1, test.use = "wilcox", only.pos = FALSE, assay = "RNA")
head(oligodendrocytes_data, n = 5)
```

```{r}
oligodendrocytes_het <- oligodendrocytes_data[oligodendrocytes_data$p_val_adj < 0.05 & oligodendrocytes_data$avg_log2FC > 0.1 | oligodendrocytes_data$avg_log2FC < -0.1, ]
oligodendrocytes_het <- oligodendrocytes_het[oligodendrocytes_het$cluster == "heterozygous", ]

# add column to data to indicate if DE or not:
oligodendrocytes_het$diffexpressed <- "NO"
# if log2Foldchange > 0.1 and pvalue < 0.05, set as "UP"
oligodendrocytes_het$diffexpressed[oligodendrocytes_het$avg_log2FC > 0.1 & oligodendrocytes_het$p_val_adj < 0.05] <- "UP"
# if log2Foldchange < -0.1 and pvalue < 0.05, set as "DOWN"
oligodendrocytes_het$diffexpressed[oligodendrocytes_het$avg_log2FC < -0.1 & oligodendrocytes_het$p_val_adj < 0.05] <- "DOWN"
```

---fibroblasts---
```{r}
# subset seurat object for each cell type
fibro <- subset(x = cerebral_int_celltypes, idents = "fibroblasts")
Idents(fibro) <- fibro@meta.data$type
fibroblasts_data <- FindAllMarkers(fibro, logfc.threshold = 0.1, test.use = "wilcox", only.pos = FALSE, assay = "RNA")
head(fibroblasts_data, n = 5)
```


```{r}
fibroblasts_het <- fibroblasts_data[fibroblasts_data$p_val_adj < 0.05 & fibroblasts_data$avg_log2FC > 0.1 | fibroblasts_data$avg_log2FC < -0.1, ]
fibroblasts_het <- fibroblasts_het[fibroblasts_het$cluster == "heterozygous", ]

# add column to data to indicate if DE or not:
fibroblasts_het$diffexpressed <- "NO"
# if log2Foldchange > 0.1 and pvalue < 0.05, set as "UP"
fibroblasts_het$diffexpressed[fibroblasts_het$avg_log2FC > 0.1 & fibroblasts_het$p_val_adj < 0.05] <- "UP"
# if log2Foldchange < -0.1 and pvalue < 0.05, set as "DOWN"
fibroblasts_het$diffexpressed[fibroblasts_het$avg_log2FC < -0.1 & fibroblasts_het$p_val_adj < 0.05] <- "DOWN"
```



#### saving all DEG in heterozygous as .csv for supplementary methods:
```{r}
write.csv(astrocytes_het, here("results/seurat/DGE_06/DEG_astrocytes.csv"), row.names = FALSE)
write.csv(excitatory_neurons_het, here("results/seurat/DGE_06/DEG_excitatory_neurons.csv"), row.names = FALSE)
write.csv(inhibitory_neurons_het, here("results/seurat/DGE_06/DEG_inhibitory_neurons.csv"), row.names = FALSE)
write.csv(microglia_het, here("results/seurat/DGE_06/DEG_microglia.csv"), row.names = FALSE)
write.csv(pericytes_het, here("results/seurat/DGE_06/DEG_pericytes_cortex.csv"), row.names = FALSE)
write.csv(oligodendrocytes_het, here("results/seurat/DGE_06/DEG_oligodendrocytes.csv"), row.names = FALSE)
write.csv(opcs_het, here("results/seurat/DGE_06/DEG_opcs.csv"), row.names = FALSE)
write.csv(fibroblasts_het, here("results/seurat/DGE_06/DEG_fibroblasts.csv"), row.names = FALSE)
```

Outputs at log2FC of 0.1, for downstream analysis 
```{r}
save(astrocytes_het, file = here("results/DEG/astrocytes_het_DEG01.Rdata"))
save(excitatory_neurons_het, file = here("results/DEG/excitatory_het_DEG01.Rdata"))
save(inhibitory_neurons_het, file = here("results/DEG/inhibitory_het_DEG01.Rdata"))
save(oligodendrocytes_het, file = here("results/DEG/oligo_het_DEG01.Rdata"))
save(opcs_het, file = here("results/DEG/opcs_het_DEG01.Rdata"))
save(pericytes_het, file = here("results/DEG/pericytes_het_DEG01.Rdata"))
save(microglia_het, file = here("results/DEG/micro_het_DEG01.Rdata"))
save(fibroblasts_het, file = here("results/DEG/fibro_het_DEG01.Rdata"))
```

### For loop to generate combined DEG output at log2FC threshold of 0.1:
```{r}
# specify the cell types in the data to loop through
celltypes <- levels(cerebral_int_celltypes)

# create empty dataframe to store the loop outputs within
DEGs_cerebralcells <- data.frame()

# for loop, loops through each of the 'celltypes' in cerebral and calculates the DEG between conditions (heterozygous and control)
for (i in celltypes) {
  # subset seurat data for each cerebral cell type ----------
  celltype <- subset(cerebral_int_celltypes, idents = i)
  # change the subsetted object identity to be type (heterozygous or control) so that DEG are calculated between the two conditions ------------
  Idents(celltype) <- celltype@meta.data$type
  # calculate DEGs between conditions -----------
  DEGs <- FindAllMarkers(celltype,
    logfc.threshold = 0.1,
    test.use = "wilcox",
    only.pos = FALSE,
    assay = "RNA"
  )
  # add an additional column to the df with cell type annotation (for downstream analysis purposes) -----------
  DEGs$cell_type <- i
  # combine the dataframe generated in the for loop with the entire list of all DEGs across cell types in cerebral ----------
  DEGs_cerebralcells <- rbind(DEGs_cerebralcells, DEGs)
}

write.csv(DEGs_cerebralcells, file = here("results/seurat/DGE_06/DEGs_cerebralcells.csv")) #written out once then commented out

#DEGs_cerebralcells <- read.csv(here("results/seurat/DGE_06/DEGs_cerebralcells.csv"))
```

#### Heatmap of expression of regulators and targets of Setbp1 across cell types:

Setbp1 target gene list was constructed in 'Setbp1_target_list_construction.Rmd' and is stored in ./results/seurat/
```{r}
setbp1_genes <- read.csv(here("results/seurat/setbp1_targets.csv"))
setbp1_genes <- setbp1_genes[, -1]
setbp1_genes <- append(setbp1_genes, "Setbp1")
```


# generating seurat input for complex heatmap from DGE data
resource: 
* NormalizeData, ScaleData  was not used because of https://www.embopress.org/doi/full/10.15252/msb.20188746

```{r}
# filter DEG list for just setbp1 targets
cerebral_subset <- DEGs_cerebralcells[DEGs_cerebralcells$gene %in% setbp1_genes, ]

# grab just log2fc values from DEG matrix, keep rownames and then order (below) by rowname
genes_mat <- cerebral_subset[, c("avg_log2FC", "cell_type", "gene", "cluster", "p_val_adj"), drop = FALSE]

# filtering for significant values
genes_mat <- genes_mat[genes_mat$p_val_adj < 0.05, ]
genes_mat <- genes_mat[, -5]

# grabbbing just S858R
genes_mat <- genes_mat[genes_mat$cluster == "heterozygous",]
genes_mat <- subset(genes_mat, select = -cluster)

genes_pivot <- pivot_wider(genes_mat, names_from = c("cell_type"), values_from = "avg_log2FC", names_repair = "minimal")
genes_pivot[is.na(genes_pivot)] <- 0
rownames <- genes_pivot$gene # add gene names as rownames
genes_pivot <- subset(genes_pivot, select = -gene) # drop gene column
rownames(genes_pivot) <- rownames

# grab meta data for plot:
meta <- as.data.frame(colnames(genes_pivot))
rownames(meta) <- meta$`colnames(genes_pivot)`
colnames(meta) <- c("cell_type")


annotation_colors <- list("cell_type" = c("excitatory_neurons" = "#DC71FA", "inhibitory_neurons" = "#00BBDB", "oligodendrocytes" = "#7997FF", "astrocytes" = "#6968B4", "microglia" = "#C216B7", "opcs" = "#00C1AA", "pericytes" = "#027461", "fibroblasts" = "#0a2f6f"))

# set heatmap annotations:
heat.anno <- HeatmapAnnotation(df = meta, show_annotation_name = TRUE, col = annotation_colors)


# ensure the column order of the expression data matches row order of annotation table
genes_pivot <- genes_pivot[, rownames(meta), drop = FALSE]
rownames(genes_pivot) <- rownames

# convert dataframe to matrix
mat <- as.matrix(genes_pivot)

# plot heatmap
# Heatmap
col_fun <- colorRamp2(c(-0.5, 0, 0.5), c("#8C5209", "white", "#35978F"))
png(
  filename = here("results/figures/cortex_DEGsetbp1_heatmap.png"),
  width = 1000,
  height = 1000
)
Heatmap(mat,
  col = col_fun, #brewer.pal(9, "BrBG"),
  heatmap_legend_param = list(title = "avg_log2FC"),
  cluster_rows = TRUE,
  cluster_columns = FALSE,
  column_order = NULL,
  show_row_dend = FALSE,
  show_column_dend = FALSE,
  show_row_names = TRUE,
  show_column_names = FALSE,
  use_raster = TRUE,
  raster_device = c("png"),
  bottom_annotation = NULL,
  top_annotation = heat.anno,
  column_title = "Differential Expression of Setbp1 Targets Across Cell-types in Cerebral Cortex", row_title = "Setbp1 Targets", row_title_side = "right"
)
dev.off()
```

```{r}
proc.time()
```
    user   system  elapsed 
 904.210   47.793 5169.713 

```{r}
sessionInfo()
```
R version 4.1.3 (2022-03-10)
Platform: x86_64-pc-linux-gnu (64-bit)
Running under: Ubuntu 20.04.5 LTS

Matrix products: default
BLAS/LAPACK: /usr/lib/x86_64-linux-gnu/openblas-pthread/libopenblasp-r0.3.8.so

locale:
[1] C

attached base packages:
[1] grid      stats     graphics  grDevices utils     datasets  methods   base     

other attached packages:
 [1] lintr_3.0.2           styler_1.9.0          here_1.0.1            circlize_0.4.15       ComplexHeatmap_2.10.0 dplyr_1.1.0           tidyr_1.3.0          
 [8] ggplot2_3.4.1         SeuratObject_4.1.3    Seurat_4.3.0         

loaded via a namespace (and not attached):
  [1] plyr_1.8.8             igraph_1.4.1           lazyeval_0.2.2         sp_1.6-0               splines_4.1.3          listenv_0.9.0          scattermore_0.8       
  [8] digest_0.6.31          foreach_1.5.2          htmltools_0.5.4        fansi_1.0.4            magrittr_2.0.3         tensor_1.5             cluster_2.1.2         
 [15] doParallel_1.0.17      ROCR_1.0-11            remotes_2.4.2          globals_0.16.2         matrixStats_0.63.0     R.utils_2.12.2         spatstat.sparse_3.0-0 
 [22] colorspace_2.1-0       ggrepel_0.9.3          xfun_0.37              callr_3.7.3            crayon_1.5.2           jsonlite_1.8.4         progressr_0.13.0      
 [29] spatstat.data_3.0-0    survival_3.3-1         zoo_1.8-11             iterators_1.0.14       glue_1.6.2             polyclip_1.10-4        gtable_0.3.1          
 [36] leiden_0.4.3           GetoptLong_1.0.5       R.cache_0.16.0         future.apply_1.10.0    shape_1.4.6            BiocGenerics_0.40.0    abind_1.4-5           
 [43] scales_1.2.1           spatstat.random_3.1-3  miniUI_0.1.1.1         Rcpp_1.0.10            viridisLite_0.4.1      xtable_1.8-4           clue_0.3-64           
 [50] reticulate_1.28        stats4_4.1.3           htmlwidgets_1.6.1      rex_1.2.1              httr_1.4.5             RColorBrewer_1.1-3     ellipsis_0.3.2        
 [57] ica_1.0-3              pkgconfig_2.0.3        R.methodsS3_1.8.2      uwot_0.1.14            deldir_1.0-6           utf8_1.2.3             tidyselect_1.2.0      
 [64] rlang_1.0.6            reshape2_1.4.4         later_1.3.0            munsell_0.5.0          tools_4.1.3            cli_3.6.0              generics_0.1.3        
 [71] ggridges_0.5.4         stringr_1.5.0          fastmap_1.1.1          goftest_1.2-3          processx_3.8.0         knitr_1.42             fitdistrplus_1.1-8    
 [78] purrr_1.0.1            RANN_2.6.1             pbapply_1.7-0          future_1.31.0          nlme_3.1-155           mime_0.12              R.oo_1.25.0           
 [85] xml2_1.3.3             compiler_4.1.3         rstudioapi_0.14        plotly_4.10.1          png_0.1-8              spatstat.utils_3.0-1   tibble_3.1.8          
 [92] stringi_1.7.12         cyclocomp_1.1.0        ps_1.7.2               desc_1.4.2             lattice_0.20-45        Matrix_1.5-3           vctrs_0.5.2           
 [99] pillar_1.8.1           lifecycle_1.0.3        spatstat.geom_3.0-6    lmtest_0.9-40          GlobalOptions_0.1.2    RcppAnnoy_0.0.20       data.table_1.14.8     
[106] cowplot_1.1.1          irlba_2.3.5.1          httpuv_1.6.9           patchwork_1.1.2        R6_2.5.1               promises_1.2.0.1       KernSmooth_2.23-20    
[113] gridExtra_2.3          IRanges_2.28.0         parallelly_1.34.0      codetools_0.2-18       MASS_7.3-55            rprojroot_2.0.3        rjson_0.2.21          
[120] withr_2.5.0            sctransform_0.3.5      S4Vectors_0.32.4       parallel_4.1.3         Rtsne_0.16             spatstat.explore_3.0-6 shiny_1.7.4          


```{r}
# run style
style_file(here("src/seurat_scripts/Setbp1_DGEcortex_06.Rmd"))
# lintr was run as well
```
