---
title: "story_time"
author: "Jordan Whitlock"
date: '2023-06-13'
output: html_document
---

Goal #1: Investigate expression of SET in kidney and cortex S858R samples, in order to confirm it's expression aligns with what is known in the literature or that it is not different across cell types and conditions. 

Goal #2: Identify if there are any NPCs in the the data. This was done by looking at Sox2 expression, a [known](https://www.sanbio.nl/resources/blogs/blog-about-npcs-nscs) marker of proliferative NPCs in the brain. 

```{r}
set.seed(2178)
library(Seurat)
library(here)
library(clustree)
ptm <- proc.time()
source(here("src/functions/functions.R"))
```


```{r}
load(here("data/setbp1_cerebralintcelltypes.Rdata"))
load(here("data/kidney_integrated_celltypes_postSoup.Rdata"))
```

```{r}
# cortex
setbp1_exp <- FetchData(cerebral_int_celltypes, vars = c("Set", "cell_type", "type"))
colnames(setbp1_exp) <- c("expression", "cell type", "condition")

cs <- ggplot(setbp1_exp, aes(`cell type`, expression, fill = condition)) +
  geom_split_violin() +
  theme_bw() +
  scale_fill_manual(values = c("control" = "#644DD4", "heterozygous" = "#5F9EFD")) + 
  theme(text = element_text(family = "Helvetica")) +
  theme(axis.text.x = element_text(angle = 45, color = "black", hjust = 1)) + 
  theme(axis.title.x = element_blank()) + 
  theme(axis.text = element_text(face = "bold")) +
  theme(legend.title = element_text(face = "bold")) +
  theme(legend.text = element_text(face = "bold")) +
  labs(y = "Expression") +
  theme(axis.title.y = element_text(face = "bold"))
  #ggtitle("Setbp1 Expression Across Cerebral Cortex")

#kidney
setbp1_exp <- FetchData(kidney_int_celltypes,
  vars = c("Set", "cell_type", "type")
)
colnames(setbp1_exp) <- c("expression", "cell type", "condition")

ks <- ggplot(setbp1_exp, aes(`cell type`, expression, fill = condition)) +
  geom_split_violin() +
  theme_bw() +
  scale_fill_manual(values = c("control" = "#FE6100", "heterozygous" = "#FFB000")) + 
  theme(text = element_text(family = "Helvetica")) +
  theme(axis.text.x = element_text(angle = 45, color = "black", hjust = 1)) + 
  theme(axis.title.x = element_blank()) +
  theme(axis.text = element_text(face = "bold")) + 
  theme(legend.title = element_text(face = "bold"))+
  theme(legend.text = element_text(face = "bold")) +
  labs(y = "Expression") +
  theme(axis.title.y = element_text(face = "bold"))

panel_2AB <- cs / ks
panel_2AB

png()
ggsave(here("results/figures/SUPP_set.png"), dpi = 250)
panel_2AB
dev.off()
```

```{r}
# cortex
setbp1_exp <- FetchData(cerebral_int_celltypes, vars = c("Sox2", "cell_type", "type"))
colnames(setbp1_exp) <- c("expression", "cell type", "condition")

cs <- VlnPlot(cerebral_int_celltypes, features = "Sox2", split.by = "type", pt.size = 0) +
  theme_bw() +
  scale_fill_manual(values = c("control" = "#644DD4", "heterozygous" = "#5F9EFD")) + 
  theme(text = element_text(family = "Helvetica")) +
  theme(axis.text.x = element_text(angle = 45, color = "black", hjust = 1)) + 
  theme(axis.title.x = element_blank()) + 
  theme(axis.text = element_text(face = "bold")) +
  theme(legend.title = element_text(face = "bold")) +
  theme(legend.text = element_text(face = "bold")) +
  labs(y = "Expression") +
  theme(axis.title.y = element_text(face = "bold"))
  #ggtitle("Setbp1 Expression Across Cerebral Cortex")

#kidney
setbp1_exp <- FetchData(kidney_int_celltypes,
  vars = c("Sox2", "cell_type", "type")
)
colnames(setbp1_exp) <- c("expression", "cell type", "condition")

ks <-VlnPlot(kidney_int_celltypes, features = "Sox2", split.by = "type", pt.size = 0) +
  theme_bw() +
  scale_fill_manual(values = c("control" = "#FE6100", "heterozygous" = "#FFB000")) + 
  theme(text = element_text(family = "Helvetica")) +
  theme(axis.text.x = element_text(angle = 45, color = "black", hjust = 1)) + 
  theme(axis.title.x = element_blank()) +
  theme(axis.text = element_text(face = "bold")) + 
  theme(legend.title = element_text(face = "bold"))+
  theme(legend.text = element_text(face = "bold")) +
  labs(y = "Expression") +
  theme(axis.title.y = element_text(face = "bold"))

panel_2AB <- cs / ks
panel_2AB

png()
ggsave(here("results/figures/SUPP_Sox2.png"), dpi = 250)
panel_2AB
dev.off()
```

```{r}
fptm <- proc.time() - ptm
fptm <- (fptm[3] / 60) / 60
print(paste0("Run time: ", fptm, " hours"))
```
[1] "Run time: 0.149947777777778 hours"

```{r}
style_file(here("src/module_scores/15_Set_Sox2_profiling.Rmd")) # Note: lintr was run as well
```

```{r}
sessionInfo()
```
R version 4.1.3 (2022-03-10)
Platform: x86_64-pc-linux-gnu (64-bit)
Running under: Ubuntu 20.04.6 LTS

Matrix products: default
BLAS/LAPACK: /usr/lib/x86_64-linux-gnu/openblas-pthread/libopenblasp-r0.3.8.so

locale:
[1] C

attached base packages:
[1] stats     graphics  grDevices utils     datasets  methods   base     

other attached packages:
[1] clustree_0.5.0     ggraph_2.1.0       ggplot2_3.4.1      here_1.0.1         SeuratObject_4.1.3 Seurat_4.3.0      

loaded via a namespace (and not attached):
  [1] Rtsne_0.16             colorspace_2.1-0       deldir_1.0-6           ellipsis_0.3.2         ggridges_0.5.4         rprojroot_2.0.3       
  [7] rstudioapi_0.14        spatstat.data_3.0-0    farver_2.1.1           leiden_0.4.3           listenv_0.9.0          graphlayouts_0.8.4    
 [13] ggrepel_0.9.3          fansi_1.0.4            codetools_0.2-18       splines_4.1.3          knitr_1.42             polyclip_1.10-4       
 [19] jsonlite_1.8.4         ica_1.0-3              cluster_2.1.2          png_0.1-8              uwot_0.1.14            ggforce_0.4.1         
 [25] shiny_1.7.4            sctransform_0.3.5      spatstat.sparse_3.0-0  compiler_4.1.3         httr_1.4.5             backports_1.4.1       
 [31] Matrix_1.5-3           fastmap_1.1.1          lazyeval_0.2.2         limma_3.50.3           cli_3.6.1              later_1.3.1           
 [37] tweenr_2.0.2           htmltools_0.5.4        tools_4.1.3            igraph_1.4.3           gtable_0.3.1           glue_1.6.2            
 [43] RANN_2.6.1             reshape2_1.4.4         dplyr_1.1.2            rappdirs_0.3.3         Rcpp_1.0.10            scattermore_0.8       
 [49] vctrs_0.6.2            spatstat.explore_3.0-6 nlme_3.1-155           progressr_0.13.0       lmtest_0.9-40          spatstat.random_3.1-3 
 [55] xfun_0.37              stringr_1.5.0          globals_0.16.2         mime_0.12              miniUI_0.1.1.1         lifecycle_1.0.3       
 [61] irlba_2.3.5.1          goftest_1.2-3          future_1.31.0          MASS_7.3-55            zoo_1.8-11             scales_1.2.1          
 [67] tidygraph_1.2.3        promises_1.2.0.1       spatstat.utils_3.0-1   parallel_4.1.3         RColorBrewer_1.1-3     yaml_2.3.7            
 [73] reticulate_1.28        pbapply_1.7-0          gridExtra_2.3          stringi_1.7.12         checkmate_2.1.0        rlang_1.1.1           
 [79] pkgconfig_2.0.3        matrixStats_0.63.0     evaluate_0.20          lattice_0.20-45        ROCR_1.0-11            purrr_1.0.1           
 [85] tensor_1.5             labeling_0.4.2         patchwork_1.1.2        htmlwidgets_1.6.1      cowplot_1.1.1          tidyselect_1.2.0      
 [91] parallelly_1.34.0      RcppAnnoy_0.0.20       plyr_1.8.8             magrittr_2.0.3         R6_2.5.1               generics_0.1.3        
 [97] pillar_1.9.0           withr_2.5.0            fitdistrplus_1.1-8     survival_3.3-1         abind_1.4-5            sp_1.6-0              
[103] tibble_3.2.1           future.apply_1.10.0    KernSmooth_2.23-20     utf8_1.2.3             spatstat.geom_3.0-6    plotly_4.10.1         
[109] rmarkdown_2.20         viridis_0.6.2          grid_4.1.3             data.table_1.14.8      digest_0.6.31          xtable_1.8-4          
[115] tidyr_1.3.0            httpuv_1.6.11          munsell_0.5.0          viridisLite_0.4.1     