---
title: "profiling failed repair in kidney"
author: "Jordan Whitlock"
date: '2023-06-13'
output: html_document
---

Goal #1: Investigate expression of failed repair markers in kindey injury in snRNA-seq celltypes of S858R and WT. 

```{r}
set.seed(2178)
library(Seurat)
library(here)
library(styler)
library(lintr)
library(clustree)
ptm <- proc.time()
source(here("src/functions/functions.R"))
```

```{r}
load(here("data/kidney_integrated_celltypes_postSoup.Rdata"))
```

```{r failed-repair-genes}
kidney_repair <- c("Slc7a13", "Slc5a12", "Slc22a30", "Sema5a", "Havcr1", "Dcdc2a", "Myc", "Krt20", "Sox9", "Vcam1")

for (i in kidney_repair) {
  png(
    filename = here(paste0("results/figures/SUPP_", i, ".png")),
    width = 1500,
    height = 1000,
    res = 250
  )
  print(VlnPlot(kidney_int_celltypes, features = i, split.by = "type", pt.size = 0) +
    theme_bw() +
    scale_fill_manual(values = c("control" = "#FE6100", "heterozygous" = "#FFB000")) +
    theme(text = element_text(family = "Helvetica")) +
    theme(axis.text.x = element_text(angle = 45, color = "black", hjust = 1)) +
    theme(axis.title.x = element_blank()) +
    theme(axis.text = element_text(face = "bold")) +
    theme(legend.title = element_text(face = "bold")) +
    theme(legend.text = element_text(face = "bold")) +
    labs(y = "Expression") +
    theme(axis.title.y = element_text(face = "bold")))
  dev.off()
}
```

```{r}
fptm <- proc.time() - ptm
fptm <- (fptm[3] / 60) / 60
print(paste0("Run time: ", fptm, " hours"))
```
[1] "Run time: 0.0116697222222223 hours"

```{r}
style_file(here("src/seurat_scripts/16_failed_repair_kidney_profiling.Rmd")) # Note: lintr was run as well
```

```{r}
sessionInfo()
```
R version 4.1.3 (2022-03-10)
Platform: x86_64-pc-linux-gnu (64-bit)
Running under: Ubuntu 20.04.6 LTS

Matrix products: default
BLAS/LAPACK: /usr/lib/x86_64-linux-gnu/openblas-pthread/libopenblasp-r0.3.8.so

locale:
[1] C

attached base packages:
[1] stats     graphics  grDevices utils     datasets  methods   base     

other attached packages:
[1] lintr_3.0.2        styler_1.9.0       clustree_0.5.0     ggraph_2.1.0       ggplot2_3.4.1      here_1.0.1         SeuratObject_4.1.3
[8] Seurat_4.3.0      

loaded via a namespace (and not attached):
  [1] Rtsne_0.16             colorspace_2.1-0       deldir_1.0-6           ellipsis_0.3.2         ggridges_0.5.4         rprojroot_2.0.3       
  [7] rstudioapi_0.14        spatstat.data_3.0-0    farver_2.1.1           leiden_0.4.3           listenv_0.9.0          remotes_2.4.2         
 [13] graphlayouts_0.8.4     ggrepel_0.9.3          fansi_1.0.4            xml2_1.3.3             R.methodsS3_1.8.2      codetools_0.2-18      
 [19] splines_4.1.3          knitr_1.42             polyclip_1.10-4        jsonlite_1.8.4         ica_1.0-3              cluster_2.1.2         
 [25] R.oo_1.25.0            png_0.1-8              uwot_0.1.14            ggforce_0.4.1          shiny_1.7.4            sctransform_0.3.5     
 [31] spatstat.sparse_3.0-0  compiler_4.1.3         httr_1.4.5             Matrix_1.5-3           fastmap_1.1.1          lazyeval_0.2.2        
 [37] cli_3.6.0              tweenr_2.0.2           later_1.3.0            htmltools_0.5.4        tools_4.1.3            igraph_1.4.1          
 [43] gtable_0.3.1           glue_1.6.2             RANN_2.6.1             reshape2_1.4.4         dplyr_1.1.0            Rcpp_1.0.10           
 [49] scattermore_0.8        vctrs_0.5.2            spatstat.explore_3.0-6 nlme_3.1-155           progressr_0.13.0       lmtest_0.9-40         
 [55] spatstat.random_3.1-3  xfun_0.37              stringr_1.5.0          ps_1.7.2               globals_0.16.2         mime_0.12             
 [61] miniUI_0.1.1.1         lifecycle_1.0.3        irlba_2.3.5.1          goftest_1.2-3          future_1.31.0          MASS_7.3-55           
 [67] zoo_1.8-11             scales_1.2.1           tidygraph_1.2.3        promises_1.2.0.1       spatstat.utils_3.0-1   rex_1.2.1             
 [73] parallel_4.1.3         RColorBrewer_1.1-3     yaml_2.3.7             reticulate_1.28        pbapply_1.7-0          gridExtra_2.3         
 [79] stringi_1.7.12         desc_1.4.2             cyclocomp_1.1.0        rlang_1.0.6            pkgconfig_2.0.3        matrixStats_0.63.0    
 [85] evaluate_0.20          lattice_0.20-45        ROCR_1.0-11            purrr_1.0.1            tensor_1.5             labeling_0.4.2        
 [91] patchwork_1.1.2        htmlwidgets_1.6.1      processx_3.8.0         cowplot_1.1.1          tidyselect_1.2.0       parallelly_1.34.0     
 [97] RcppAnnoy_0.0.20       plyr_1.8.8             magrittr_2.0.3         R6_2.5.1               generics_0.1.3         pillar_1.8.1          
[103] withr_2.5.0            fitdistrplus_1.1-8     survival_3.3-1         abind_1.4-5            sp_1.6-0               tibble_3.1.8          
[109] future.apply_1.10.0    crayon_1.5.2           KernSmooth_2.23-20     utf8_1.2.3             spatstat.geom_3.0-6    plotly_4.10.1         
[115] rmarkdown_2.20         viridis_0.6.2          grid_4.1.3             data.table_1.14.8      callr_3.7.3            digest_0.6.31         
[121] R.cache_0.16.0         xtable_1.8-4           tidyr_1.3.0            httpuv_1.6.9           R.utils_2.12.2         munsell_0.5.0         
[127] viridisLite_0.4.1     
