---
title: "Setbp1_QC_01"
author: "Jordan Whitlock"
date: '2022-09-06'
output:     
  html_document:
      toc: true
      toc_depth: 2
      toc_float: true
---
# *Single-nuclei analysis: Quality Control*
* load in data and create Seurat objects
* check proportion of mitochondrial and ribosomal genes
* visualize data feature distribution
* filter nFeatureRNA and mt
* check cell cycle scoring
* re-visualize data feature distribution

## Code: 

Accompanying dry lab notebook for this analysis is here: https://docs.google.com/document/d/1FO_Je9oxbBFF9S_k-8tVF2t4ba88O4C77sw1v7DIFec/edit?usp=sharing

### Checking libraries and working path settings:

```{r setup}
getwd()
.libPaths() # "/usr/local/lib/R/site-library" "/usr/local/lib/R/library" "/home/jbarham3/R/x86_64-pc-linux-gnu-library/4.1"
set.seed(2178)
```

### Loading in all required libraries
```{r, load-libraries message=FALSE}
library(devtools)
library(magrittr)
library(readr)
library(dplyr)
library(Matrix)
library(Seurat)
library(cowplot)
library(ggplot2)
library(here)
library(tidyverse)
library(gprofiler2)
library(styler)
library(lintr)
```


### Load in the Setbp1 data:
All Cell Ranger outs were moved from the introns_processed_data respective outs directory to a counts directory at:/data/project/lasseigne_lab/JordanWhitlock/jw_setbp1/counts/ to make it easier to loop between and load in all files at once.

Make sure all of your outs are placed into a single head directory called "counts" and within that directory there is a folder for each sample and within each sample specific folder a "barcodes.tsv.gx", "features.tsv.gz" and "matrix.mtx.gz" file.
```{r}
counts_list <- list.dirs("/data/project/lasseigne_lab/JordanWhitlock/jw_setbp1/counts", recursive = FALSE)
```
False makes sure the "/data/project/lasseigne_lab/JordanWhitlock/jw_setbp1/counts" is not included in the list. This command lists all the sample folders in your counts directory
# Loading in all of the files and creating seurat objects named as "sample.id_seurat_object" using the sample.ids vector above:
```{r}
for (i in counts_list) {
  counts <- Read10X(i)
  print(i)
  sample_name <- basename(i)
  sample_name <- gsub("[[:punct:]]", "", sample_name)
  seurat_object <- CreateSeuratObject(counts = counts, project = sample_name)
  assign(paste0(sample_name, "_seurat_object"), seurat_object)
}
```


### Annotating seurat objects for condition with "control" and "heterozygous" metadata tags:
```{r}
J1_seurat_object$type <- "heterozygous"
J2_seurat_object$type <- "control"
J3_seurat_object$type <- "control"
J4_seurat_object$type <- "control"
J13_seurat_object$type <- "heterozygous"
J15_seurat_object$type <- "heterozygous"
K1_seurat_object$type <- "control"
K2_seurat_object$type <- "heterozygous"
K3_seurat_object$type <- "control"
K4_seurat_object$type <- "heterozygous"
K5_seurat_object$type <- "control"
K6_seurat_object$type <- "heterozygous"
```


### Creating kidney and cerebral cortex seurat objects and adding cell.ids:
```{r}
# kidney
kidney_seurat <- merge(K1_seurat_object,
  c(K2_seurat_object, K3_seurat_object, K4_seurat_object, K5_seurat_object, K6_seurat_object),
  add.cell.ids = c("K1_ctrl", "K2_het", "K3_ctrl", "K4_het", "K5_ctrl", "K6_het")
)

# cerebral cortex
cerebral_seurat <- merge(J1_seurat_object,
  c(J2_seurat_object, J3_seurat_object, J4_seurat_object, J13_seurat_object, J15_seurat_object),
  add.cell.ids = c("J1_het", "J2_ctrl", "J3_ctrl", "J4_ctrl", "J13_het", "J15_het")
)

# checking metadata assignment:
as.data.frame(kidney_seurat@assays$RNA@counts[1:10, 1:2])
head(kidney_seurat@meta.data, 10)
as.data.frame(cerebral_seurat@assays$RNA@counts[1:10, 1:2])
head(cerebral_seurat@meta.data, 10)

# freeing up global environment to release R memory and have clean working space:
rm(list = setdiff(ls(), c("cerebral_seurat", "kidney_seurat")))

# running garbage to collect free memory:
gc()
```

### Perform Quality Control:
Having the data in a suitable format, we can start calculating some quality metrics. We can for example calculate the percentage of mitochondrial and ribosomal genes per cell and add that to the metadata. This will be helpful to visualize them across different metadata parameters. There are several ways of doing this, and below I manually calculate the proportion of mitochondrial reads and add to the metadata table.

Note from Simple Single Cell workflows (Lun, McCarthy & Marioni, 2017): High proportions are indicative of poor-quality cells (Islam et al. 2014; Ilicic et al. 2016), possibly because of loss of cytoplasmic RNA from perforated cells. The reasoning is that mitochondria are larger than individual transcript molecules and less likely to escape through tears in the cell membrane.

#### Remove mitochondrial genes

Note: PercentageFeatureSet function has grepping out of Mt prefixes built in. It is an automatic way to remove mt genes instead of manually grepping from rownames. (Source: https://nbisweden.github.io/workshop-scRNAseq/labs/compiled/seurat/seurat_01_qc.html )

In the mm10 reference (version 3.0.0) in the Cell Ranger reference files reveals that for whatever reason, the MT genes are labeled with lowercase mt instead. (source: https://www.michaelchimenti.com/2019/03/calculate-mitochondrial-for-mouse-scrna-seq/)
```{r}
cerebral_seurat <- PercentageFeatureSet(cerebral_seurat,
  "mt-",
  col.name = "percent_mito"
)
kidney_seurat <- PercentageFeatureSet(kidney_seurat,
  "mt-",
  col.name = "percent_mito"
)

# check it ran properly
head(cerebral_seurat)
head(kidney_seurat)
```


#### Calculate proportion of gene expression from ribosomal genes:
```{r}
cerebral_seurat <- PercentageFeatureSet(cerebral_seurat,
  "^Rp[sl]",
  col.name = "percent_ribo"
)
kidney_seurat <- PercentageFeatureSet(kidney_seurat,
  "^Rp[sl]",
  col.name = "percent_ribo"
)
# check it ran properly
head(cerebral_seurat)
head(kidney_seurat)
```


### Plot Quality Control metrics just calculated
```{r}
feats <- c("nFeature_RNA", "nCount_RNA", "percent_mito", "percent_ribo")
# violin:
png(
  filename = here("results/seurat/QC_01/preQC_metrics_VlnPlot_cortex.png"),
  width = 1000,
  height = 1000
)
VlnPlot(cerebral_seurat,
  group.by = "orig.ident",
  features = feats,
  pt.size = 0.1,
  ncol = 3
)
dev.off()

png(
  filename = here("results/seurat/QC_01/preQC_metrics_VlnPlot_kidney.png"),
  width = 1000,
  height = 1000
)
VlnPlot(kidney_seurat,
  group.by = "orig.ident",
  features = feats,
  pt.size = 0.1, ncol = 3
)
dev.off()

# scatter:
png(
  filename = here("results/seurat/QC_01/preQC_metrics_scatter_countsfeatures_cortex.png"),
  width = 1000,
  height = 1000
)
FeatureScatter(cerebral_seurat,
  "nCount_RNA",
  "nFeature_RNA",
  group.by = "orig.ident",
  pt.size = 0.5
)
dev.off()

png(
  filename = here("results/seurat/QC_01/preQC_metrics_scatter_mitoribo_cortex.png"),
  width = 1000,
  height = 1000
)
FeatureScatter(cerebral_seurat,
  "percent_mito",
  "percent_ribo",
  group.by = "orig.ident",
  pt.size = 0.5
)
dev.off()

png(
  filename = here("results/seurat/QC_01/preQC_metrics_scatter_countsfeatures_kidney.png"),
  width = 1000,
  height = 1000
)
FeatureScatter(kidney_seurat,
  "nCount_RNA",
  "nFeature_RNA",
  group.by = "orig.ident",
  pt.size = 0.5
)
dev.off()

png(
  filename = here("results/seurat/QC_01/preQC_metrics_scatter_mitoribo_kidney.png"),
  width = 1000,
  height = 1000
)
FeatureScatter(kidney_seurat,
  "percent_mito",
  "percent_ribo",
  group.by = "orig.ident",
  pt.size = 0.5
)
dev.off()
```


##### Visualize the number of cell counts per sample
```{r}
# Add number of genes per UMI for each cell to metadata
kidney_seurat$log10GenesPerUMI <- log10(kidney_seurat$nFeature_RNA) / log10(kidney_seurat$nCount_RNA)
cerebral_seurat$log10GenesPerUMI <- log10(cerebral_seurat$nFeature_RNA) / log10(cerebral_seurat$nCount_RNA)

png(
  filename = here("results/seurat/QC_01/preQC_metrics_nucleicount_bysample_kidney.png"),
  width = 1000,
  height = 1000
)
kidney_seurat@meta.data %>%
  ggplot(aes(x = orig.ident, fill = orig.ident)) +
  geom_bar() +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  theme(plot.title = element_text(hjust = 0.5, face = "bold")) +
  ggtitle("NCells")
dev.off()

png(
  filename = here("results/seurat/QC_01/preQC_metrics_nucleicount_bysample_cortex.png"),
  width = 1000,
  height = 1000
)
cerebral_seurat@meta.data %>%
  ggplot(aes(x = orig.ident, fill = orig.ident)) +
  geom_bar() +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  theme(plot.title = element_text(hjust = 0.5, face = "bold")) +
  ggtitle("NCells")
dev.off()

png(
  filename = here("results/seurat/QC_01/preQC_metrics_nucleicount_bytype_kidney.png"),
  width = 1000,
  height = 1000
)
kidney_seurat@meta.data %>%
  ggplot(aes(x = orig.ident, fill = type)) +
  geom_bar() +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  theme(plot.title = element_text(hjust = 0.5, face = "bold")) +
  ggtitle("NCells")
dev.off()

png(
  filename = here("results/seurat/QC_01/preQC_metrics_nucleicount_bytype_cortex.png"),
  width = 1000,
  height = 1000
)
cerebral_seurat@meta.data %>%
  ggplot(aes(x = orig.ident, fill = type)) +
  geom_bar() +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  theme(plot.title = element_text(hjust = 0.5, face = "bold")) +
  ggtitle("NCells")
dev.off()
```


#### Visualize the number of UMIs/transcripts per cell:
nUMI = nCount_RNA
nGene = nFeature_RNA

500 is the minimum threshold for number of UMI counts per cell.
```{r}
png(
  filename = here("results/seurat/QC_01/preQC_metrics_nCount_bysample_kidney.png"),
  width = 1000,
  height = 1000
)
kidney_seurat@meta.data %>%
  ggplot(aes(color = orig.ident, x = nCount_RNA, fill = orig.ident)) +
  geom_density(alpha = 0.2) +
  scale_x_log10() +
  theme_classic() +
  ylab("Cell density") +
  geom_vline(xintercept = 500)
dev.off()

png(
  filename = here("results/seurat/QC_01/preQC_metrics_nCount_bysample_cortex.png"),
  width = 1000,
  height = 1000
)
cerebral_seurat@meta.data %>%
  ggplot(aes(color = orig.ident, x = nCount_RNA, fill = orig.ident)) +
  geom_density(alpha = 0.2) +
  scale_x_log10() +
  theme_classic() +
  ylab("Cell density") +
  geom_vline(xintercept = 500)
dev.off()

png(
  filename = here("results/seurat/QC_01/preQC_metrics_nCount_bytype_kidney.png"),
  width = 1000,
  height = 1000
)
kidney_seurat@meta.data %>%
  ggplot(aes(color = type, x = nCount_RNA, fill = type)) +
  geom_density(alpha = 0.2) +
  scale_x_log10() +
  theme_classic() +
  ylab("Cell density") +
  geom_vline(xintercept = 500)
dev.off()

png(
  filename = here("results/seurat/QC_01/preQC_metrics_nCount_bytype_cortex.png"),
  width = 1000,
  height = 1000
)
cerebral_seurat@meta.data %>%
  ggplot(aes(color = type, x = nCount_RNA, fill = type)) +
  geom_density(alpha = 0.2) +
  scale_x_log10() +
  theme_classic() +
  ylab("Cell density") +
  geom_vline(xintercept = 500)
dev.off()
```


##### Visualize the distribution of nFeature_RNA (genes) detected per cell via density plot
```{r}
png(
  filename = here("results/seurat/QC_01/preQC_metrics_nFeature_bysample_cortex.png"),
  width = 1000,
  height = 1000
)
cerebral_seurat@meta.data %>%
  ggplot(aes(color = orig.ident, x = nFeature_RNA, fill = orig.ident)) +
  geom_density(alpha = 0.2) +
  theme_classic() +
  scale_x_log10() +
  geom_vline(xintercept = 300)
dev.off()

png(
  filename = here("results/seurat/QC_01/preQC_metrics_nFeature_bysample_kidney.png"),
  width = 1000,
  height = 1000
)
kidney_seurat@meta.data %>%
  ggplot(aes(color = orig.ident, x = nFeature_RNA, fill = orig.ident)) +
  geom_density(alpha = 0.2) +
  theme_classic() +
  scale_x_log10() +
  geom_vline(xintercept = 300)
dev.off()

png(
  filename = here("results/seurat/QC_01/preQC_metrics_nFeature_bytype_cortex.png"),
  width = 1000,
  height = 1000
)
cerebral_seurat@meta.data %>%
  ggplot(aes(color = type, x = nFeature_RNA, fill = type)) +
  geom_density(alpha = 0.2) +
  theme_classic() +
  scale_x_log10() +
  geom_vline(xintercept = 300)
dev.off()

png(
  filename = here("results/seurat/QC_01/preQC_metrics_nFeature_bytype_kidney.png"),
  width = 1000,
  height = 1000
)
kidney_seurat@meta.data %>%
  ggplot(aes(color = type, x = nFeature_RNA, fill = type)) +
  geom_density(alpha = 0.2) +
  theme_classic() +
  scale_x_log10() +
  geom_vline(xintercept = 300)
dev.off()
```


#### Visualize the distribution of genes detected per cell via boxplot
```{r}
png(
  filename = here("results/seurat/QC_01/preQC_metrics_nCellvsnGenes_kidney.png"),
  width = 1000,
  height = 1000
)
kidney_seurat@meta.data %>%
  ggplot(aes(x = orig.ident, y = log10(nFeature_RNA), fill = orig.ident)) +
  geom_boxplot() +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  theme(plot.title = element_text(hjust = 0.5, face = "bold")) +
  ggtitle("NCells vs NGenes")
dev.off()

png(
  filename = here("results/seurat/QC_01/preQC_metrics_nCellvsnGenes_cortex.png"),
  width = 1000,
  height = 1000
)
cerebral_seurat@meta.data %>%
  ggplot(aes(x = orig.ident, y = log10(nFeature_RNA), fill = orig.ident)) +
  geom_boxplot() +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  theme(plot.title = element_text(hjust = 0.5, face = "bold")) +
  ggtitle("NCells vs NGenes")
dev.off()
```


#### Visualize the correlation between nFeature_RNA (genes) and nCount_RNA (nuclei)
```{r}
# determine whether strong presence of cells with low numbers of genes/UMIs
png(
  filename = here("results/seurat/QC_01/preQC_metrics_nFeaturenCountCorrelation_kidney.png"),
  width = 1000,
  height = 1000
)
kidney_seurat@meta.data %>%
  ggplot(aes(x = nCount_RNA, y = nFeature_RNA, color = percent_mito)) +
  geom_point() +
  scale_color_gradient(low = "gray90", high = "black") +
  stat_smooth(method = lm) +
  scale_x_log10() +
  scale_y_log10() +
  theme_classic() +
  geom_vline(xintercept = 500) +
  geom_hline(yintercept = 250) +
  facet_wrap(~orig.ident)
dev.off()

png(
  filename = here("results/seurat/QC_01/preQC_metrics_nFeaturenCountCorrelation_cortex.png"),
  width = 1000,
  height = 1000
)
cerebral_seurat@meta.data %>%
  ggplot(aes(x = nCount_RNA, y = nFeature_RNA, color = percent_mito)) +
  geom_point() +
  scale_color_gradient(low = "gray90", high = "black") +
  stat_smooth(method = lm) +
  scale_x_log10() +
  scale_y_log10() +
  theme_classic() +
  geom_vline(xintercept = 500) +
  geom_hline(yintercept = 250) +
  facet_wrap(~orig.ident)
dev.off()
```


#### Filter out nFeature_RNA and mt:
```{r}
kidney_seurat_filtered <- subset(
  kidney_seurat,
  subset = nFeature_RNA >= 1000 & nFeature_RNA <= 15000 & percent_mito < 5
)
cerebral_seurat_filtered <- subset(
  cerebral_seurat,
  subset = nFeature_RNA >= 1000 & nFeature_RNA <= 15000 & percent_mito < 5
)
# 5.0 here because percent_mito is a percent not a ratio.
```


#### Plot genes contributing most to reads:
```{r}
png(
  filename = here("results/seurat/QC_01/postfilt_feature_mito_contrib_genes_cortex.png"),
  width = 1000,
  height = 1000
)
par(mar = c(4, 8, 2, 1))
C <- cerebral_seurat_filtered@assays$RNA@counts
C <- Matrix::t(Matrix::t(C) / Matrix::colSums(C)) * 100
most_expressed <- order(apply(C, 1, median), decreasing = TRUE)[20:1]
boxplot(t(as.matrix(C[most_expressed, ])),
  cex = 0.1,
  las = 1, xlab = "% total count per cell in cerebral cortex",
  col = (scales::hue_pal())(20)[20:1], horizontal = TRUE
)
dev.off()

png(
  filename = here("results/seurat/QC_01/postfilt_feature_mito_contrib_genes_kidney.png"),
  width = 1000,
  height = 1000
)
par(mar = c(4, 8, 2, 1))
C <- kidney_seurat_filtered@assays$RNA@counts
C <- Matrix::t(Matrix::t(C) / Matrix::colSums(C)) * 100
C <- as(C, "dgTMatrix")
most_expressed <- order(apply(C, 1, median), decreasing = TRUE)[20:1]
boxplot(t(as.matrix(C[most_expressed, ])),
  cex = 0.1, las = 1,
  xlab = "% total count per cell in kidney",
  col = (scales::hue_pal())(20)[20:1], horizontal = TRUE
)
dev.off()
```


#### Removing Malat1:
```{r}
kidney_seurat_filtered <- kidney_seurat_filtered[!grepl(
  "Malat1",
  rownames(kidney_seurat_filtered)
), ]
cerebral_seurat_filtered <- cerebral_seurat_filtered[!grepl(
  "Malat1",
  rownames(cerebral_seurat_filtered)
), ]
```


#### Checking cell cycle scoring
```{r}
# Change default from human to mouse, Seurat uses a human list and we are working with mouse data. (code credit: TCH)
# reading in cell cycle gene list (downloaded from https://github.com/hbc/tinyatlas/blob/master/cell_cycle/Mus_musculus.csv) on 230322
cell_cycle_gene <- read_csv(here("results/seurat/230322_mouse_cell_cycle_genes.csv"))
# converting ENSMUSG to Gene Symbol
symbols <- gconvert(cell_cycle_gene$geneID,
  organism = "mmusculus",
  target = "ENTREZGENE",
  filter_na = FALSE
)
symbols <- symbols[, c(2, 5)]
colnames(symbols) <- c("geneID", "symbol")

mouse_cell_cycle_gene <- merge(cell_cycle_gene, symbols, by = "geneID") %>% .[, c(1, 2, 4)]
mouse_cell_cycle_gene <- mouse_cell_cycle_gene[order(mouse_cell_cycle_gene$phase), ] %>% .[, c(2, 3)]

# Before running CellCycleScoring the data need to be normalized and log transformed.
kidney_seurat_filtered <- NormalizeData(kidney_seurat_filtered)

# now check cell cycle scoring
kidney_seurat_filtered <- CellCycleScoring(
  object = kidney_seurat_filtered,
  g2m.features = mouse_cell_cycle_gene[1:54, ],
  s.features = mouse_cell_cycle_gene[55:97, ]
)

png(
  filename = here("results/seurat/QC_01/postfiltQC_cellcycle_genes_kidney.png"),
  width = 1000,
  height = 1000
)
VlnPlot(kidney_seurat_filtered,
  features = c("S.Score", "G2M.Score"), group.by = "orig.ident",
  ncol = 4, pt.size = 0.1
)
dev.off()

# Before running CellCycleScoring the data need to be normalized and log transformed.
cerebral_seurat_filtered <- NormalizeData(cerebral_seurat_filtered)

# now check cell cycle scoring
cerebral_seurat_filtered <- CellCycleScoring(
  object = cerebral_seurat_filtered,
  g2m.features = mouse_cell_cycle_gene[1:54, ],
  s.features = mouse_cell_cycle_gene[55:97, ]
)
png(
  filename = here("results/seurat/QC_01/postfiltQC_cellcycle_genes_cortex.png"),
  width = 1000,
  height = 1000
)
VlnPlot(cerebral_seurat_filtered,
  features = c("S.Score", "G2M.Score"), group.by = "orig.ident",
  ncol = 4, pt.size = 0.1
)
dev.off()
```

#### Revisualizing QC metrics after data has been filtered:

##### Plot Quality Control metrics just calculated
```{r}
feats <- c("nFeature_RNA", "nCount_RNA", "percent_mito", "percent_ribo")
# violin:
png(
  filename = here("results/seurat/QC_01/postQC_metrics_VlnPlot_cortex.png"),
  width = 1000,
  height = 1000
)
VlnPlot(
  cerebral_seurat_filtered,
  group.by = "orig.ident",
  features = feats,
  pt.size = 0.1,
  ncol = 3
)
dev.off()

png(
  filename = here("results/seurat/QC_01/postQC_metrics_VlnPlot_kidney.png"),
  width = 1000,
  height = 1000
)
VlnPlot(
  kidney_seurat_filtered,
  group.by = "orig.ident",
  features = feats,
  pt.size = 0.1,
  ncol = 3
)
dev.off()

# scatter:
png(
  filename = here("results/seurat/QC_01/postQC_metrics_scatter_countsfeatures_cortex.png"),
  width = 1000,
  height = 1000
)
FeatureScatter(
  cerebral_seurat_filtered,
  "nCount_RNA",
  "nFeature_RNA",
  group.by = "orig.ident",
  pt.size = 0.5
)
dev.off()

png(
  filename = here("results/seurat/QC_01/postQC_metrics_scatter_mitoribo_cortex.png"),
  width = 1000,
  height = 1000
)
FeatureScatter(
  cerebral_seurat_filtered,
  "percent_mito",
  "percent_ribo",
  group.by = "orig.ident",
  pt.size = 0.5
)
dev.off()

png(
  filename = here("results/seurat/QC_01/postQC_metrics_scatter_countsfeatures_kidney.png"),
  width = 1000,
  height = 1000
)
FeatureScatter(
  kidney_seurat_filtered,
  "nCount_RNA",
  "nFeature_RNA",
  group.by = "orig.ident",
  pt.size = 0.5
)
dev.off()

png(
  filename = here("results/seurat/QC_01/postQC_metrics_scatter_mitribo_kidney.png"),
  width = 1000,
  height = 1000
)
FeatureScatter(
  kidney_seurat_filtered,
  "percent_mito",
  "percent_ribo",
  group.by = "orig.ident",
  pt.size = 0.5
)
dev.off()
```

###### Visualize the number of cell counts per sample
```{r}
# Add number of genes per UMI for each cell to metadata -- tells the complexity of the data
kidney_seurat_filtered$log10GenesPerUMI <- log10(kidney_seurat_filtered$nFeature_RNA) / log10(kidney_seurat_filtered$nCount_RNA)
cerebral_seurat_filtered$log10GenesPerUMI <- log10(cerebral_seurat_filtered$nFeature_RNA) / log10(cerebral_seurat_filtered$nCount_RNA)

png(
  filename = here("results/seurat/QC_01/postQC_metrics_nucleicount_bysample_kidney.png"),
  width = 1000,
  height = 1000
)
kidney_seurat_filtered@meta.data %>%
  ggplot(aes(x = orig.ident, fill = orig.ident)) +
  geom_bar() +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  theme(plot.title = element_text(hjust = 0.5, face = "bold")) +
  ggtitle("NCells")
dev.off()

png(
  filename = here("results/seurat/QC_01/postQC_metrics_nucleicount_bysample_cortex.png"),
  width = 1000,
  height = 1000
)
cerebral_seurat_filtered@meta.data %>%
  ggplot(aes(x = orig.ident, fill = orig.ident)) +
  geom_bar() +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  theme(plot.title = element_text(hjust = 0.5, face = "bold")) +
  ggtitle("NCells")
dev.off()

png(
  filename = here("results/seurat/QC_01/postQC_metrics_nucleicount_bytype_kidney.png"),
  width = 1000,
  height = 1000
)
kidney_seurat_filtered@meta.data %>%
  ggplot(aes(x = orig.ident, fill = type)) +
  geom_bar() +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  theme(plot.title = element_text(hjust = 0.5, face = "bold")) +
  ggtitle("NCells")
dev.off()

png(
  filename = here("results/seurat/QC_01/postQC_metrics_nucleicount_bytype_cortex.png"),
  width = 1000,
  height = 1000
)
cerebral_seurat_filtered@meta.data %>%
  ggplot(aes(x = orig.ident, fill = type)) +
  geom_bar() +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  theme(plot.title = element_text(hjust = 0.5, face = "bold")) +
  ggtitle("NCells")
dev.off()
```
##### Visualize the number of UMIs/transcripts per cell:
nUMI = nCount_RNA
nGene = nFeature_RNA

500 is the minimum threshold for number of UMI counts per cell.
```{r}
png(
  filename = here("results/seurat/QC_01/postQC_metrics_nCount_bysample_kidney.png"),
  width = 1000,
  height = 1000
)
kidney_seurat_filtered@meta.data %>%
  ggplot(aes(color = orig.ident, x = nCount_RNA, fill = orig.ident)) +
  geom_density(alpha = 0.2) +
  scale_x_log10() +
  theme_classic() +
  ylab("Cell density") +
  geom_vline(xintercept = 500)
dev.off()

png(
  filename = here("results/seurat/QC_01/postQC_metrics_nCount_bysample_cortex.png"),
  width = 1000,
  height = 1000
)
cerebral_seurat_filtered@meta.data %>%
  ggplot(aes(color = orig.ident, x = nCount_RNA, fill = orig.ident)) +
  geom_density(alpha = 0.2) +
  scale_x_log10() +
  theme_classic() +
  ylab("Cell density") +
  geom_vline(xintercept = 500)
dev.off()

png(
  filename = here("results/seurat/QC_01/postQC_metrics_nCount_bytype_kidney.png"),
  width = 1000,
  height = 1000
)
kidney_seurat_filtered@meta.data %>%
  ggplot(aes(color = type, x = nCount_RNA, fill = type)) +
  geom_density(alpha = 0.2) +
  scale_x_log10() +
  theme_classic() +
  ylab("Cell density") +
  geom_vline(xintercept = 500)
dev.off()

png(
  filename = here("results/seurat/QC_01/postQC_metrics_nCount_bytype_cortex.png"),
  width = 1000,
  height = 1000
)
cerebral_seurat_filtered@meta.data %>%
  ggplot(aes(color = type, x = nCount_RNA, fill = type)) +
  geom_density(alpha = 0.2) +
  scale_x_log10() +
  theme_classic() +
  ylab("Cell density") +
  geom_vline(xintercept = 500)
dev.off()
```

###### Visualize the distribution of nFeature_RNA (genes) detected per cell via density plot
```{r}
png(
  filename = here("results/seurat/QC_01/postQC_metrics_nFeature_bysample_cortex.png"),
  width = 1000,
  height = 1000
)
cerebral_seurat_filtered@meta.data %>%
  ggplot(aes(color = orig.ident, x = nFeature_RNA, fill = orig.ident)) +
  geom_density(alpha = 0.2) +
  theme_classic() +
  scale_x_log10() +
  geom_vline(xintercept = 300)
dev.off()

png(
  filename = here("results/seurat/QC_01/postQC_metrics_nFeature_bysample_kidney.png"),
  width = 1000,
  height = 1000
)
kidney_seurat_filtered@meta.data %>%
  ggplot(aes(color = orig.ident, x = nFeature_RNA, fill = orig.ident)) +
  geom_density(alpha = 0.2) +
  theme_classic() +
  scale_x_log10() +
  geom_vline(xintercept = 300)
dev.off()

png(
  filename = here("results/seurat/QC_01/postQC_metrics_nFeature_bytype_cortex.png"),
  width = 1000,
  height = 1000
)
cerebral_seurat_filtered@meta.data %>%
  ggplot(aes(color = type, x = nFeature_RNA, fill = type)) +
  geom_density(alpha = 0.2) +
  theme_classic() +
  scale_x_log10() +
  geom_vline(xintercept = 300)
dev.off()

png(
  filename = here("results/seurat/QC_01/postQC_metrics_nFeature_bytype_kidney.png"),
  width = 1000,
  height = 1000
)
kidney_seurat_filtered@meta.data %>%
  ggplot(aes(color = type, x = nFeature_RNA, fill = type)) +
  geom_density(alpha = 0.2) +
  theme_classic() +
  scale_x_log10() +
  geom_vline(xintercept = 300)
dev.off()
```

##### Visualize the distribution of genes detected per cell via boxplot
```{r}
kidney_seurat_filtered@meta.data %>%
  ggplot(aes(x = orig.ident, y = log10(nFeature_RNA), fill = orig.ident)) +
  geom_boxplot() +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  theme(plot.title = element_text(hjust = 0.5, face = "bold")) +
  ggtitle("NCells vs NGenes")
cerebral_seurat_filtered@meta.data %>%
  ggplot(aes(x = orig.ident, y = log10(nFeature_RNA), fill = orig.ident)) +
  geom_boxplot() +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  theme(plot.title = element_text(hjust = 0.5, face = "bold")) +
  ggtitle("NCells vs NGenes")
```

##### Visualize the correlation between nFeature_RNA (genes) and nCount_RNA (nuclei)
```{r}
# determine whether strong presence of cells with low numbers of genes/UMIs
png(
  filename = here("results/seurat/QC_01/postQC_metrics_nFeaturenCountCorrelation_kidney.png"),
  width = 1000,
  height = 1000
)
kidney_seurat_filtered@meta.data %>%
  ggplot(aes(x = nCount_RNA, y = nFeature_RNA, color = percent_mito)) +
  geom_point() +
  scale_color_gradient(low = "gray90", high = "black") +
  stat_smooth(method = lm) +
  scale_x_log10() +
  scale_y_log10() +
  theme_classic() +
  geom_vline(xintercept = 500) +
  geom_hline(yintercept = 250) +
  facet_wrap(~orig.ident)
dev.off()

png(
  filename = here("results/seurat/QC_01/postQC_metrics_nFeaturenCountCorrelation_cortex.png"),
  width = 1000,
  height = 1000
)
cerebral_seurat_filtered@meta.data %>%
  ggplot(aes(x = nCount_RNA, y = nFeature_RNA, color = percent_mito)) +
  geom_point() +
  scale_color_gradient(low = "gray90", high = "black") +
  stat_smooth(method = lm) +
  scale_x_log10() +
  scale_y_log10() +
  theme_classic() +
  geom_vline(xintercept = 500) +
  geom_hline(yintercept = 250) +
  facet_wrap(~orig.ident)
dev.off()
```

#### Saving filtered data as new object:
Note: this object is just filtered for nCount_RNA, mitochondria, and Malat1. It was normalized and log transformed for cell cycle scoring.
```{r}
save(kidney_seurat_filtered, file = here("data/filtered_setbp1_kidney_seurat.Rdata"))
save(cerebral_seurat_filtered, file = here("data/filtered_setbp1_cerebral_seurat.Rdata"))
```

```{r}
sessionInfo()
```
R version 4.1.3 (2022-03-10)
Platform: x86_64-pc-linux-gnu (64-bit)
Running under: Ubuntu 20.04.5 LTS

Matrix products: default
BLAS/LAPACK: /usr/lib/x86_64-linux-gnu/openblas-pthread/libopenblasp-r0.3.8.so

locale:
[1] C

attached base packages:
[1] stats     graphics  grDevices utils     datasets  methods   base     

other attached packages:
 [1] lintr_3.0.2        styler_1.9.0       gprofiler2_0.2.1   lubridate_1.9.2    forcats_1.0.0      stringr_1.5.0      purrr_1.0.1        tidyr_1.3.0       
 [9] tibble_3.1.8       tidyverse_2.0.0    here_1.0.1         ggplot2_3.4.1      cowplot_1.1.1      SeuratObject_4.1.3 Seurat_4.3.0       Matrix_1.5-3      
[17] dplyr_1.1.0        readr_2.1.4        magrittr_2.0.3     devtools_2.4.5     usethis_2.1.6     

loaded via a namespace (and not attached):
  [1] plyr_1.8.8             igraph_1.4.1           lazyeval_0.2.2         sp_1.6-0               splines_4.1.3          listenv_0.9.0         
  [7] scattermore_0.8        digest_0.6.31          htmltools_0.5.4        fansi_1.0.4            memoise_2.0.1          tensor_1.5            
 [13] cluster_2.1.2          ROCR_1.0-11            tzdb_0.3.0             remotes_2.4.2          globals_0.16.2         matrixStats_0.63.0    
 [19] vroom_1.6.1            R.utils_2.12.2         timechange_0.2.0       spatstat.sparse_3.0-0  prettyunits_1.1.1      colorspace_2.1-0      
 [25] ggrepel_0.9.3          xfun_0.37              RCurl_1.98-1.10        callr_3.7.3            crayon_1.5.2           jsonlite_1.8.4        
 [31] progressr_0.13.0       spatstat.data_3.0-0    survival_3.3-1         zoo_1.8-11             glue_1.6.2             polyclip_1.10-4       
 [37] gtable_0.3.1           leiden_0.4.3           R.cache_0.16.0         pkgbuild_1.3.1         future.apply_1.10.0    abind_1.4-5           
 [43] scales_1.2.1           spatstat.random_3.1-3  miniUI_0.1.1.1         Rcpp_1.0.10            viridisLite_0.4.1      xtable_1.8-4          
 [49] reticulate_1.28        bit_4.0.5              profvis_0.3.7          rex_1.2.1              htmlwidgets_1.6.1      httr_1.4.5            
 [55] RColorBrewer_1.1-3     ellipsis_0.3.2         ica_1.0-3              farver_2.1.1           urlchecker_1.0.1       pkgconfig_2.0.3       
 [61] R.methodsS3_1.8.2      uwot_0.1.14            deldir_1.0-6           utf8_1.2.3             labeling_0.4.2         tidyselect_1.2.0      
 [67] rlang_1.0.6            reshape2_1.4.4         later_1.3.0            munsell_0.5.0          tools_4.1.3            cachem_1.0.7          
 [73] cli_3.6.0              generics_0.1.3         ggridges_0.5.4         fastmap_1.1.1          goftest_1.2-3          bit64_4.0.5           
 [79] processx_3.8.0         knitr_1.42             fs_1.6.1               fitdistrplus_1.1-8     RANN_2.6.1             pbapply_1.7-0         
 [85] future_1.31.0          nlme_3.1-155           mime_0.12              R.oo_1.25.0            xml2_1.3.3             compiler_4.1.3        
 [91] rstudioapi_0.14        plotly_4.10.1          png_0.1-8              spatstat.utils_3.0-1   stringi_1.7.12         cyclocomp_1.1.0       
 [97] ps_1.7.2               desc_1.4.2             lattice_0.20-45        vctrs_0.5.2            pillar_1.8.1           lifecycle_1.0.3       
[103] spatstat.geom_3.0-6    lmtest_0.9-40          RcppAnnoy_0.0.20       bitops_1.0-7           data.table_1.14.8      irlba_2.3.5.1         
[109] httpuv_1.6.9           patchwork_1.1.2        R6_2.5.1               promises_1.2.0.1       KernSmooth_2.23-20     gridExtra_2.3         
[115] parallelly_1.34.0      sessioninfo_1.2.2      codetools_0.2-18       MASS_7.3-55            pkgload_1.3.2          rprojroot_2.0.3       
[121] withr_2.5.0            sctransform_0.3.5      mgcv_1.8-39            parallel_4.1.3         hms_1.1.2              grid_4.1.3            
[127] Rtsne_0.16             spatstat.explore_3.0-6 shiny_1.7.4           

```{r}
# run style
style_file(here("src/seurat_scripts/Setbp1_QC_01.Rmd"))
# lintr was run as well
```


Next script: Setbp1_DoubletFinder_02
