---
title: "Setbp1_DGEkidney_06"
author: "Jordan Whitlock"
date: '2022-09-14'
output:     
  html_document:
      toc: true
      toc_depth: 2
      toc_float: true
---

This analysis comes after 'Setbp1_CellTypesKidney_05.Rmd' and precedes 'Setbp1_PathwayAnalysisCortexKidney_07.Rmd'

# *Single-nuclei analysis: Identify DEG in Kidney data*


## Code: 

#### Loading in libraries:
```{r}
set.seed(2178)
library(Seurat)
library(SeuratObject)
library(ggplot2)
library(tidyr)
library(dplyr)
library(ComplexHeatmap)
library(circlize)
library(here)
library(styler)
library(lintr)
source(here("src/functions/functions.R"))
```

#### Loading in the object:
```{r}
load(here("data/kidney_integrated_celltypes_postSoup.Rdata"))

Idents(kidney_int_celltypes) <- "cell_type"
```

#### checking to see if Setbp1 is even in the seurat object:

It is present in the data, but appears very similar across conditions.
```{r}
# visualize with FeaturePlot
png(
  file = here("results/seurat/DGE_06/Setbp1_expression_kidney.png"),
  width = 1000,
  height = 1000
)
plot <- FeaturePlot(kidney_int_celltypes,
  feature = "Setbp1",
  reduction = "umap_harmony",
  split.by = "type", label = TRUE
)
plot
dev.off()
```
#### split violin plotting of setbp1 across celltypes
```{r}
# grabbing the Setbp1 expression data and moving from S4 to a dataframe
setbp1_exp <- FetchData(kidney_int_celltypes,
  vars = c("Setbp1", "cell_type", "type")
)
colnames(setbp1_exp) <- c("Expression", "cell_type", "type")

png(
  filename = here("results/figures/kidney_setbp1_splitviolin.png"),
  width = 1500,
  height = 1000
)
ggplot(setbp1_exp, aes(cell_type, Expression, fill = type)) +
  geom_split_violin() +
  theme_bw() +
  ggtitle("Setbp1 Expression across Kidney") +
  scale_fill_manual(values = c("control" = "#FE6100", "heterozygous" = "#FFB000"))
dev.off()
```




#### Calculating DGE by cell type: log2FC of 0.1

---B_cells---
```{r}
# subset seurat object for each cell type
bcell <- subset(x = kidney_int_celltypes, idents = "B_cells")
Idents(bcell) <- bcell@meta.data$type
bcell_data <- FindAllMarkers(bcell,
  logfc.threshold = 0.1,
  test.use = "wilcox",
  only.pos = FALSE,
  assay = "RNA"
)
head(bcell_data, n = 5)
```

```{r}
# subsetting just the DEG from Het:
bcell_het <- bcell_data[bcell_data$p_val_adj < 0.5 &
  bcell_data$avg_log2FC > 0.1 |
  bcell_data$avg_log2FC < -0.1, ]
bcell_het <- bcell_het[bcell_het$cluster == "heterozygous", ]

# add column to data to indicate if DE or not:
bcell_het$diffexpressed <- "NO"
# if log2Foldchange > 0.1 and pvalue < 0.05, set as "UP"
bcell_het$diffexpressed[bcell_het$avg_log2FC > 0.1 &
  bcell_het$p_val_adj < 0.05] <- "UP"
# if log2Foldchange < -0.1 and pvalue < 0.05, set as "DOWN"
bcell_het$diffexpressed[bcell_het$avg_log2FC < -0.1 &
  bcell_het$p_val_adj < 0.05] <- "DOWN"
```

---Collecting Duct Intercalating Cells Type A---
```{r}
# subset seurat object for each cell type
CDIC_typeA <- subset(x = kidney_int_celltypes, idents = "CDIC_typeA")
Idents(CDIC_typeA) <- CDIC_typeA@meta.data$type
CDIC_typeA_data <- FindAllMarkers(CDIC_typeA,
  logfc.threshold = 0.1,
  test.use = "wilcox",
  only.pos = FALSE,
  assay = "RNA"
)
head(CDIC_typeA_data, n = 5)
```

```{r}
# subsetting just the DEG from Het:
CDIC_typeA_het <- CDIC_typeA_data[CDIC_typeA_data$p_val_adj < 0.5 &
  CDIC_typeA_data$avg_log2FC > 0.1 |
  CDIC_typeA_data$avg_log2FC < -0.1, ]
CDIC_typeA_het <- CDIC_typeA_het[CDIC_typeA_het$cluster == "heterozygous", ]

CDIC_typeA_het$diffexpressed <- "NO"
# if log2Foldchange > 0.1 and pvalue < 0.05, set as "UP"
CDIC_typeA_het$diffexpressed[CDIC_typeA_het$avg_log2FC > 0.1 &
  CDIC_typeA_het$p_val_adj < 0.05] <- "UP"
# if log2Foldchange < -0.1 and pvalue < 0.05, set as "DOWN"
CDIC_typeA_het$diffexpressed[CDIC_typeA_het$avg_log2FC < -0.1 &
  CDIC_typeA_het$p_val_adj < 0.05] <- "DOWN"
```


---Collecting Duct Intercalating Cells Type B---
```{r}
# subset seurat object for each cell type
CDIC_typeB <- subset(x = kidney_int_celltypes, idents = "CDIC_typeB")
Idents(CDIC_typeB) <- CDIC_typeB@meta.data$type
CDIC_typeB_data <- FindAllMarkers(CDIC_typeB,
  logfc.threshold = 0.1,
  test.use = "wilcox",
  only.pos = FALSE,
  assay = "RNA"
)
head(CDIC_typeB_data, n = 5)
```

```{r}
CDIC_typeB_het <- CDIC_typeB_data[CDIC_typeB_data$p_val_adj < 0.5 &
  CDIC_typeB_data$avg_log2FC > 0.1 |
  CDIC_typeB_data$avg_log2FC < -0.1, ]
CDIC_typeB_het <- CDIC_typeB_het[CDIC_typeB_het$cluster == "heterozygous", ]

# add column to data to indicate if DE or not:
CDIC_typeB_het$diffexpressed <- "NO"
# if log2Foldchange > 0.1 and pvalue < 0.05, set as "UP"
CDIC_typeB_het$diffexpressed[CDIC_typeB_het$avg_log2FC > 0.1 &
  CDIC_typeB_het$p_val_adj < 0.05] <- "UP"
# if log2Foldchange < -0.1 and pvalue < 0.05, set as "DOWN"
CDIC_typeB_het$diffexpressed[CDIC_typeB_het$avg_log2FC < -0.1 &
  CDIC_typeB_het$p_val_adj < 0.05] <- "DOWN"
```

---Collecting Duct Principal Cells---
```{r}
# subset seurat object for each cell type
CDPC <- subset(x = kidney_int_celltypes, idents = "CDPC")
Idents(CDPC) <- CDPC@meta.data$type
CDPC_data <- FindAllMarkers(CDPC,
  logfc.threshold = 0.1,
  test.use = "wilcox",
  only.pos = FALSE,
  assay = "RNA"
)
head(CDPC_data, n = 5)
```

```{r}
CDPC_het <- CDPC_data[CDPC_data$p_val_adj < 0.5 &
  CDPC_data$avg_log2FC > 0.1 |
  CDPC_data$avg_log2FC < -0.1, ]
CDPC_het <- CDPC_het[CDPC_het$cluster == "heterozygous", ]

# add column to data to indicate if DE or not:
CDPC_het$diffexpressed <- "NO"
# if log2Foldchange > 0.1 and pvalue < 0.05, set as "UP"
CDPC_het$diffexpressed[CDPC_het$avg_log2FC > 0.1 &
  CDPC_het$p_val_adj < 0.05] <- "UP"
# if log2Foldchange < -0.1 and pvalue < 0.05, set as "DOWN"
CDPC_het$diffexpressed[CDPC_het$avg_log2FC < -0.1 &
  CDPC_het$p_val_adj < 0.05] <- "DOWN"
```


---Distal Convoluted Tubule---
```{r}
# subset seurat object for each cell type
DCT <- subset(x = kidney_int_celltypes, idents = "DCT")
Idents(DCT) <- DCT@meta.data$type
DCT_data <- FindAllMarkers(DCT,
  logfc.threshold = 0.1,
  test.use = "wilcox",
  only.pos = FALSE,
  assay = "RNA"
)
head(DCT_data, n = 5)
```

```{r}
# subsetting just the DEG from Het:
DCT_het <- DCT_data[DCT_data$p_val_adj < 0.5 &
  DCT_data$avg_log2FC > 0.1 | DCT_data$avg_log2FC < -0.1, ]
DCT_het <- DCT_het[DCT_het$cluster == "heterozygous", ]

# add column to data to indicate if DE or not:
DCT_het$diffexpressed <- "NO"
# if log2Foldchange > 0.1 and pvalue < 0.05, set as "UP"
DCT_het$diffexpressed[DCT_het$avg_log2FC > 0.1 & DCT_het$p_val_adj < 0.05] <- "UP"
# if log2Foldchange < -0.1 and pvalue < 0.05, set as "DOWN"
DCT_het$diffexpressed[DCT_het$avg_log2FC < -0.1 & DCT_het$p_val_adj < 0.05] <- "DOWN"
```

---Distal Loop of Henle---
```{r}
# subset seurat object for each cell type
DLH <- subset(x = kidney_int_celltypes, idents = "DLH")
Idents(DLH) <- DLH@meta.data$type
DLH_data <- FindAllMarkers(DLH,
  logfc.threshold = 0.1,
  test.use = "wilcox",
  only.pos = FALSE,
  assay = "RNA"
)
head(DLH_data, n = 5)
```


```{r}
# subsetting just the DEG from Het:
DLH_het <- DLH_data[DLH_data$p_val_adj < 0.5 &
  DLH_data$avg_log2FC > 0.1 | DLH_data$avg_log2FC < -0.1, ]
DLH_het <- DLH_het[DLH_het$cluster == "heterozygous", ]


# add column to data to indicate if DE or not:
DLH_het$diffexpressed <- "NO"
# if log2Foldchange > 0.1 and pvalue < 0.05, set as "UP"
DLH_het$diffexpressed[DLH_het$avg_log2FC > 0.1 & DLH_het$p_val_adj < 0.05] <- "UP"
# if log2Foldchange < -0.1 and pvalue < 0.05, set as "DOWN"
DLH_het$diffexpressed[DLH_het$avg_log2FC < -0.1 & DLH_het$p_val_adj < 0.05] <- "DOWN"
```

---Endothelial---
```{r}
# subset seurat object for each cell type
endo <- subset(x = kidney_int_celltypes, idents = "endothelial")
Idents(endo) <- endo@meta.data$type
endo_data <- FindAllMarkers(endo,
  logfc.threshold = 0.1,
  test.use = "wilcox",
  only.pos = FALSE,
  assay = "RNA"
)
head(endo_data, n = 5)
```

```{r}
# subsetting just the DEG from Het:
endo_het <- endo_data[endo_data$p_val_adj < 0.5 &
  endo_data$avg_log2FC > 0.1 | endo_data$avg_log2FC < -0.1, ]
endo_het <- endo_het[endo_het$cluster == "heterozygous", ]
# add column to data to indicate if DE or not:
endo_het$diffexpressed <- "NO"
# if log2Foldchange > 0.1 and pvalue < 0.05, set as "UP"
endo_het$diffexpressed[endo_het$avg_log2FC > 0.1 &
  endo_het$p_val_adj < 0.05] <- "UP"
# if log2Foldchange < -0.1 and pvalue < 0.05, set as "DOWN"
endo_het$diffexpressed[endo_het$avg_log2FC < -0.1 &
  endo_het$p_val_adj < 0.05] <- "DOWN"
```

---Fibroblasts---
```{r}
# subset seurat object for each cell type
fibro <- subset(x = kidney_int_celltypes, idents = "fibroblasts")
Idents(fibro) <- fibro@meta.data$type
fibro_data <- FindAllMarkers(fibro,
  logfc.threshold = 0.1,
  test.use = "wilcox",
  only.pos = FALSE,
  assay = "RNA"
)
head(fibro_data, n = 5)
```

```{r}
# subsetting just the DEG from Het:
fibro_het <- fibro_data[fibro_data$p_val_adj < 0.5 &
  fibro_data$avg_log2FC > 0.1 |
  fibro_data$avg_log2FC < -0.1, ]
fibro_het <- fibro_het[fibro_het$cluster == "heterozygous", ]

# add column to data to indicate if DE or not:
fibro_het$diffexpressed <- "NO"
# if log2Foldchange > 0.1 and pvalue < 0.05, set as "UP"
fibro_het$diffexpressed[fibro_het$avg_log2FC > 0.1 &
  fibro_het$p_val_adj < 0.05] <- "UP"
# if log2Foldchange < -0.1 and pvalue < 0.05, set as "DOWN"
fibro_het$diffexpressed[fibro_het$avg_log2FC < -0.1 &
  fibro_het$p_val_adj < 0.05] <- "DOWN"
```


---LOH---
```{r}
# subset seurat object for each cell type
LOH <- subset(x = kidney_int_celltypes, idents = "LOH")
Idents(LOH) <- LOH@meta.data$type
LOH_data <- FindAllMarkers(LOH,
  logfc.threshold = 0.1,
  test.use = "wilcox",
  only.pos = FALSE,
  assay = "RNA"
)
head(LOH_data, n = 5)
```

```{r}
# subsetting just the DEG from Het:
LOH_het <- LOH_data[LOH_data$p_val_adj < 0.5 &
  LOH_data$avg_log2FC > 0.1 | LOH_data$avg_log2FC < -0.1, ]
LOH_het <- LOH_het[LOH_het$cluster == "heterozygous", ]

# add column to data to indicate if DE or not:
LOH_het$diffexpressed <- "NO"
# if log2Foldchange > 0.1 and pvalue < 0.05, set as "UP"
LOH_het$diffexpressed[LOH_het$avg_log2FC > 0.1 & LOH_het$p_val_adj < 0.05] <- "UP"
# if log2Foldchange < -0.1 and pvalue < 0.05, set as "DOWN"
LOH_het$diffexpressed[LOH_het$avg_log2FC < -0.1 & LOH_het$p_val_adj < 0.05] <- "DOWN"
```


---macrophages---
```{r}
# subset seurat object for each cell type
macro <- subset(x = kidney_int_celltypes, idents = "macrophages")
Idents(macro) <- macro@meta.data$type
macro_data <- FindAllMarkers(macro,
  logfc.threshold = 0.1,
  test.use = "wilcox",
  only.pos = FALSE,
  assay = "RNA"
)
head(macro_data, n = 5)
```

```{r}
# subsetting just the DEG from Het:
macro_het <- macro_data[macro_data$p_val_adj < 0.5 &
  macro_data$avg_log2FC > 0.1 |
  macro_data$avg_log2FC < -0.1, ]
macro_het <- macro_het[macro_het$cluster == "heterozygous", ]

# add column to data to indicate if DE or not:
macro_het$diffexpressed <- "NO"
# if log2Foldchange > 0.1 and pvalue < 0.05, set as "UP"
macro_het$diffexpressed[macro_het$avg_log2FC > 0.1 &
  macro_het$p_val_adj < 0.05] <- "UP"
# if log2Foldchange < -0.1 and pvalue < 0.05, set as "DOWN"
macro_het$diffexpressed[macro_het$avg_log2FC < -0.1 &
  macro_het$p_val_adj < 0.05] <- "DOWN"
```

---Proximal Convoluted Tubule Segment 1---
```{r}
# subset seurat object for each cell type
PCTS1 <- subset(x = kidney_int_celltypes, idents = "PCTS1")
Idents(PCTS1) <- PCTS1@meta.data$type
PCTS1_data <- FindAllMarkers(PCTS1,
  logfc.threshold = 0.1,
  test.use = "wilcox",
  only.pos = FALSE,
  assay = "RNA"
)
head(PCTS1_data, n = 5)
```

```{r}
# subsetting just the DEG from Het:
PCTS1_het <- PCTS1_data[PCTS1_data$p_val_adj < 0.5 &
  PCTS1_data$avg_log2FC > 0.1 |
  PCTS1_data$avg_log2FC < -0.1, ]
PCTS1_het <- PCTS1_het[PCTS1_het$cluster == "heterozygous", ]

# add column to data to indicate if DE or not:
PCTS1_het$diffexpressed <- "NO"
# if log2Foldchange > 0.1 and pvalue < 0.05, set as "UP"
PCTS1_het$diffexpressed[PCTS1_het$avg_log2FC > 0.1 &
  PCTS1_het$p_val_adj < 0.05] <- "UP"
# if log2Foldchange < -0.1 and pvalue < 0.05, set as "DOWN"
PCTS1_het$diffexpressed[PCTS1_het$avg_log2FC < -0.1 &
  PCTS1_het$p_val_adj < 0.05] <- "DOWN"
```


---podocytes---
```{r}
# subset seurat object for each cell type
podo <- subset(x = kidney_int_celltypes, idents = "podocytes")
Idents(podo) <- podo@meta.data$type
podo_data <- FindAllMarkers(podo,
  logfc.threshold = 0.1,
  test.use = "wilcox",
  only.pos = FALSE,
  assay = "RNA"
)
head(podo_data, n = 5)
```



```{r}
# subsetting just the DEG from Het:
podo_het <- podo_data[podo_data$p_val_adj < 0.5 &
  podo_data$avg_log2FC > 0.1 |
  podo_data$avg_log2FC < -0.1, ]
podo_het <- podo_het[podo_het$cluster == "heterozygous", ]

# add column to data to indicate if DE or not:
podo_het$diffexpressed <- "NO"
# if log2Foldchange > 0.1 and pvalue < 0.05, set as "UP"
podo_het$diffexpressed[podo_het$avg_log2FC > 0.1 &
  podo_het$p_val_adj < 0.05] <- "UP"
# if log2Foldchange < -0.1 and pvalue < 0.05, set as "DOWN"
podo_het$diffexpressed[podo_het$avg_log2FC < -0.1 &
  podo_het$p_val_adj < 0.05] <- "DOWN"
```


---Proximal Straight Tubule---
```{r}
# subset seurat object for each cell type
PST <- subset(x = kidney_int_celltypes, idents = "PST")
Idents(PST) <- PST@meta.data$type
PST_data <- FindAllMarkers(PST,
  logfc.threshold = 0.1,
  test.use = "wilcox",
  only.pos = FALSE,
  assay = "RNA"
)
head(PST_data, n = 5)
```



```{r}
# subsetting just the DEG from Het:
PST_het <- PST_data[PST_data$p_val_adj < 0.5 & PST_data$avg_log2FC > 0.1 |
  PST_data$avg_log2FC < -0.1, ]
PST_het <- PST_het[PST_het$cluster == "heterozygous", ]

# add column to data to indicate if DE or not:
PST_het$diffexpressed <- "NO"
# if log2Foldchange > 0.1 and pvalue < 0.05, set as "UP"
PST_het$diffexpressed[PST_het$avg_log2FC > 0.1 & PST_het$p_val_adj < 0.05] <- "UP"
# if log2Foldchange < -0.1 and pvalue < 0.05, set as "DOWN"
PST_het$diffexpressed[PST_het$avg_log2FC < -0.1 & PST_het$p_val_adj < 0.05] <- "DOWN"
```



---pericytes---
```{r}
# subset seurat object for each cell type
pericytes <- subset(x = kidney_int_celltypes, idents = "pericytes")
Idents(pericytes) <- pericytes@meta.data$type
pericytes_data <- FindAllMarkers(pericytes,
  logfc.threshold = 0.1,
  test.use = "wilcox",
  only.pos = FALSE,
  assay = "RNA"
)
head(pericytes_data, n = 5)
```



```{r}
# subsetting just the DEG from Het:
pericytes_het <- pericytes_data[pericytes_data$p_val_adj < 0.5 &
  pericytes_data$avg_log2FC > 0.1 |
  pericytes_data$avg_log2FC < -0.1, ]
pericytes_het <- pericytes_het[pericytes_het$cluster == "heterozygous", ]

# add column to data to indicate if DE or not:
pericytes_het$diffexpressed <- "NO"
# if log2Foldchange > 0.1 and pvalue < 0.05, set as "UP"
pericytes_het$diffexpressed[pericytes_het$avg_log2FC > 0.1 &
  pericytes_het$p_val_adj < 0.05] <- "UP"
# if log2Foldchange < -0.1 and pvalue < 0.05, set as "DOWN"
pericytes_het$diffexpressed[pericytes_het$avg_log2FC < -0.1 &
  pericytes_het$p_val_adj < 0.05] <- "DOWN"
```

---PCTS2---
```{r}
# subset seurat object for each cell type
PCTS2 <- subset(x = kidney_int_celltypes, idents = "PCTS2")
Idents(PCTS2) <- PCTS2@meta.data$type
PCTS2_data <- FindAllMarkers(PCTS2,
  logfc.threshold = 0.1,
  test.use = "wilcox",
  only.pos = FALSE,
  assay = "RNA"
)
head(PCTS2_data, n = 5)
```

```{r}
# subsetting just the DEG from Het:
PCTS2_het <- PCTS2_data[PCTS2_data$p_val_adj < 0.5 &
  PCTS2_data$avg_log2FC > 0.1 |
  PCTS2_data$avg_log2FC < -0.1, ]
PCTS2_het <- PCTS2_het[PCTS2_het$cluster == "heterozygous", ]

# add column to data to indicate if DE or not:
PCTS2_het$diffexpressed <- "NO"
# if log2Foldchange > 0.1 and pvalue < 0.05, set as "UP"
PCTS2_het$diffexpressed[PCTS2_het$avg_log2FC > 0.1 &
  PCTS2_het$p_val_adj < 0.05] <- "UP"
# if log2Foldchange < -0.1 and pvalue < 0.05, set as "DOWN"
PCTS2_het$diffexpressed[PCTS2_het$avg_log2FC < -0.1 &
  PCTS2_het$p_val_adj < 0.05] <- "DOWN"
```

---smcs---
```{r}
# subset seurat object for each cell type
smcs <- subset(x = kidney_int_celltypes, idents = "smooth_muscle_cells")
Idents(smcs) <- smcs@meta.data$type
smcs_data <- FindAllMarkers(smcs,
  logfc.threshold = 0.1,
  test.use = "wilcox",
  only.pos = FALSE,
  assay = "RNA"
)
head(smcs_data, n = 5)
```

```{r}
# subsetting just the DEG from Het:
smcs_het <- smcs_data[smcs_data$p_val_adj < 0.5 &
  smcs_data$avg_log2FC > 0.1 |
  smcs_data$avg_log2FC < -0.1, ]
smcs_het <- smcs_het[smcs_het$cluster == "heterozygous", ]

# add column to data to indicate if DE or not:
smcs_het$diffexpressed <- "NO"
# if log2Foldchange > 0.1 and pvalue < 0.05, set as "UP"
smcs_het$diffexpressed[smcs_het$avg_log2FC > 0.1 &
  smcs_het$p_val_adj < 0.05] <- "UP"
# if log2Foldchange < -0.1 and pvalue < 0.05, set as "DOWN"
smcs_het$diffexpressed[smcs_het$avg_log2FC < -0.1 &
  smcs_het$p_val_adj < 0.05] <- "DOWN"
```

---prox_tubule---
```{r}
# subset seurat object for each cell type
prox_tubule <- subset(x = kidney_int_celltypes, idents = "proximal_tubule")
Idents(prox_tubule) <- prox_tubule@meta.data$type
prox_tubule_data <- FindAllMarkers(prox_tubule,
  logfc.threshold = 0.1,
  test.use = "wilcox",
  only.pos = FALSE,
  assay = "RNA"
)
head(prox_tubule_data, n = 5)
```

```{r}
# subsetting just the DEG from Het:
prox_tubule_het <- prox_tubule_data[prox_tubule_data$p_val_adj < 0.5 &
  prox_tubule_data$avg_log2FC > 0.1 |
  prox_tubule_data$avg_log2FC < -0.1, ]
prox_tubule_het <- prox_tubule_het[prox_tubule_het$cluster == "heterozygous", ]

# add column to data to indicate if DE or not:
prox_tubule_het$diffexpressed <- "NO"
# if log2Foldchange > 0.1 and pvalue < 0.05, set as "UP"
prox_tubule_het$diffexpressed[prox_tubule_het$avg_log2FC > 0.1 &
  prox_tubule_het$p_val_adj < 0.05] <- "UP"
# if log2Foldchange < -0.1 and pvalue < 0.05, set as "DOWN"
prox_tubule_het$diffexpressed[prox_tubule_het$avg_log2FC < -0.1 &
  prox_tubule_het$p_val_adj < 0.05] <- "DOWN"
```

Outputs at log2FC of 0.1:
```{r}
save(bcell_het, file = here("results/DEG/bcell_het_kid_DEG01.Rdata"))
save(CDPC_het, file = here("results/DEG/CDPC_het_kid_DEG01.Rdata"))
save(CDIC_typeA_het, file = here("results/DEG/CDICA_het_kid_DEG01.Rdata"))
save(CDIC_typeB_het, file = here("results/DEG/CDICB_het_kid_DEG01.Rdata"))
save(DCT_het, file = here("results/DEG/DCT_het_kid_DEG01.Rdata"))
save(DLH_het, file = here("results/DEG/DLH_het_kid_DEG01.Rdata"))
save(endo_het, file = here("results/DEG/endo_het_kid_DEG01.Rdata"))
save(fibro_het, file = here("results/DEG/fibro_het_kid_DEG01.Rdata"))
save(LOH_het, file = here("results/DEG/LOH_het_kid_DEG01.Rdata"))
save(macro_het, file = here("results/DEG/macro_het_kid_DEG01.Rdata"))
save(PCTS1_het, file = here("results/DEG/PCTS1_het_kid_DEG01.Rdata"))
save(podo_het, file = here("results/DEG/podo_het_kid_DEG01.Rdata"))
save(PST_het, file = here("results/DEG/PST_het_kid_DEG01.Rdata"))
save(pericytes_het, file = here("results/DEG/pericytes_het_kid_DEG01.Rdata"))
save(PCTS2_het, file = here("results/DEG/PCTS2_het_kid_DEG01.Rdata"))
save(smcs_het, file = here("results/DEG/smcs_het_kid_DEG01.Rdata"))
save(prox_tubule_het, file = here("results/DEG/prox_tubule_het_kid_DEG01.Rdata"))
```

#### saving all DEG in heterozygous as .csv for supplementary methods:
```{r}
write.csv(bcell_het, here("results/seurat/DGE_06/DEG_bcell.csv"), row.names = FALSE)
write.csv(CDPC_het, here("results/seurat/DGE_06/DEG_CDPC.csv"), row.names = FALSE)
write.csv(CDIC_typeA_het, here("results/seurat/DGE_06/DEG_CDIC_typeA.csv"), row.names = FALSE)
write.csv(CDIC_typeB_het, here("results/seurat/DGE_06/DEG_CDIC_typeB.csv"), row.names = FALSE)
write.csv(DCT_het, here("results/seurat/DGE_06/DEG_DCT.csv"), row.names = FALSE)
write.csv(DLH_het, here("results/seurat/DGE_06/DEG_DLH.csv"), row.names = FALSE)
write.csv(endo_het, here("results/seurat/DGE_06/DEG_endo.csv"), row.names = FALSE)
write.csv(fibro_het, here("results/seurat/DGE_06/DEG_fibro_kid.csv"), row.names = FALSE)
write.csv(LOH_het, here("results/seurat/DGE_06/DEG_LOH.csv"), row.names = FALSE)
write.csv(macro_het, here("results/seurat/DGE_06/DEG_macro.csv"), row.names = FALSE)
write.csv(PCTS1_het, here("results/seurat/DGE_06/DEG_PCTS1.csv"), row.names = FALSE)
write.csv(podo_het, here("results/seurat/DGE_06/DEG_podo.csv"), row.names = FALSE)
write.csv(PST_het, here("results/seurat/DGE_06/DEG_PST.csv"), row.names = FALSE)
write.csv(pericytes_het, here("results/seurat/DGE_06/DEG_pericytes.csv"), row.names = FALSE)
write.csv(PCTS2_het, here("results/seurat/DGE_06/DEG_PCTS2.csv"), row.names = FALSE)
write.csv(smcs_het, here("results/seurat/DGE_06/DEG_smcs.csv"), row.names = FALSE)
write.csv(prox_tubule_het, here("results/seurat/DGE_06/DEG_prox_tubule.csv"), row.names = FALSE)
```

### For loop to generate DEG output at log2FC threshold of 0.1 :
```{r}
# specify the cell types in the data to loop through
celltypes <- levels(kidney_int_celltypes)

# create empty dataframe to store the loop outputs within
DEGs_kidneycells <- data.frame()

# for loop, loops through each of the 'celltypes' in kidney and calculates the DEG between conditions (heterozygous and control)
for (i in celltypes) {
  # subset seurat data for each kidney cell type ----------
  celltype <- subset(kidney_int_celltypes, idents = i)
  # change the subsetted object identity to be type (heterozygous or control) so that DEG are calculated between the two conditions ------------
  Idents(celltype) <- celltype@meta.data$type
  # calculate DEGs between conditions -----------
  DEGs <- FindAllMarkers(celltype,
    logfc.threshold = 0.1,
    test.use = "wilcox",
    only.pos = FALSE,
    assay = "RNA"
  )
  # add an additional column to the df with celltype annotation (for downstream analysis purposes) -----------
  DEGs$cell_type <- i
  # combine the dataframe generated in the for loop with the entire list of all DEGs across cell types in kidney ----------
  DEGs_kidneycells <- rbind(DEGs_kidneycells, DEGs)
}

write.csv(DEGs_kidneycells, file = here("results/seurat/DGE_06/DEGs_kidneycells.csv")) # written out once then commented out
# DEGs_kidneycells <- read.csv(here("results/seurat/DGE_06/DEGs_kidneycells.csv"))
```


#### Heatmap of expression of regulators and targets of Setbp1 across cell types:
Setbp1 target gene list was constructed in 'Setbp1_target_list_construction.Rmd' 
```{r}
setbp1_genes <- read.csv(here("results/seurat/setbp1_targets.csv"))
setbp1_genes <- setbp1_genes[, -1]
setbp1_genes <- append(setbp1_genes, "Setbp1")
```

# generating seurat input for complex heatmap from DGE data
```{r}
# filter DEG list for just setbp1 targets
kidney_subset <- DEGs_kidneycells[DEGs_kidneycells$gene %in% setbp1_genes, ]

# grab just log2fc values from DEG matrix, keep rownames and then order (below) by rowname
genes_mat <- kidney_subset[, c("avg_log2FC", "cell_type", "gene", "cluster", "p_val_adj"),
  drop = FALSE
]

# filtering for significant values
genes_mat <- genes_mat[genes_mat$p_val_adj < 0.5, ]
genes_mat <- genes_mat[, -5]

genes_pivot <- pivot_wider(genes_mat,
  names_from = c("cell_type", "cluster"),
  values_from = "avg_log2FC",
  names_repair = "minimal"
)
genes_pivot[is.na(genes_pivot)] <- 0
rownames <- genes_pivot$gene # add gene names as rownames
genes_pivot <- genes_pivot[, -1, drop = FALSE] # drop gene column
rownames(genes_pivot) <- rownames

# grab meta data for plot:
meta <- as.data.frame(colnames(genes_pivot))
meta.data <- t(as.data.frame(strsplit(sub("(_)(?=[^_]+$)", " ", meta$`colnames(genes_pivot)`, perl = TRUE), " ")))
rownames(meta.data) <- meta$`colnames(genes_pivot)`
colnames(meta.data) <- c("cell_type", "condition")


# define order of annotations and colors:
ordered.meta.data <- meta.data[order(rownames(meta.data)), ] # order by cell type


annotation_colors <- list(
  "condition" = c(
    "control" = "#FE6100",
    "heterozygous" = "#FFB000"
  ),
  "cell_type" = c(
    "B_cells" = "#FFBE1D",
    "CDIC_typeA" = "#4E3801",
    "CDIC_typeB" = "#EEDF37",
    "CDPC" = "#F03F00",
    "DCT" = "#6D0404",
    "DLH" = "#AA6320",
    "endothelial" = "#CF9400",
    "fibroblasts" = "#FFC8C4",
    "LOH" = "#AA937E",
    "macrophages" = "#C70F0F",
    "PCTS1" = "#A85A5A",
    "podocytes" = "#BD085E",
    "PST" = "#948802",
    "pericytes" = "#FFE196",
    "PCTS2" = "#FF7600",
    "proximal_tubule" = "#F8766D",
    "smooth_muscle_cells" = "#FFE196"
  )
)

# set heatmap annotations:
heat.anno <- HeatmapAnnotation(
  df = ordered.meta.data,
  show_annotation_name = TRUE,
  col = annotation_colors
)


# ensure the column order of the expression data matches row order of annotation table
genes_pivot <- genes_pivot[, rownames(ordered.meta.data), drop = FALSE]
rownames(genes_pivot) <- rownames

# convert dataframe to matrix
mat <- as.matrix(genes_pivot)

# plot heatmap
# Heatmap
col_fun <- colorRamp2(c(-1, 0, 1), c("purple", "black", "yellow"))
png(
  filename = here("results/figures/kidney_DEGsetbp1_heatmap.png"),
  width = 2000,
  height = 2000
)
Heatmap(mat,
  col = col_fun,
  heatmap_legend_param = list(title = "avg_log2FC"),
  cluster_rows = TRUE,
  cluster_columns = FALSE,
  column_order = NULL,
  show_row_dend = TRUE,
  show_column_dend = TRUE,
  show_row_names = TRUE,
  show_column_names = FALSE,
  use_raster = TRUE,
  raster_device = c("png"),
  bottom_annotation = NULL,
  top_annotation = heat.anno,
  column_title = "Differential Expression of Setbp1 Targets Across Cell-types in Kidney", row_title = "Setbp1 Targets", row_title_side = "right"
)
dev.off()
```

```{r}
proc.time()
```
     user    system   elapsed 
 5027.117   112.462 12964.077

```{r}
sessionInfo()
```

R version 4.1.3 (2022-03-10)
Platform: x86_64-pc-linux-gnu (64-bit)
Running under: Ubuntu 20.04.5 LTS

Matrix products: default
BLAS/LAPACK: /usr/lib/x86_64-linux-gnu/openblas-pthread/libopenblasp-r0.3.8.so

locale:
[1] C

attached base packages:
[1] grid      stats     graphics  grDevices utils     datasets  methods   base     

other attached packages:
 [1] lintr_3.0.2           styler_1.9.0          here_1.0.1            circlize_0.4.15       ComplexHeatmap_2.10.0 dplyr_1.1.0           tidyr_1.3.0          
 [8] ggplot2_3.4.1         SeuratObject_4.1.3    Seurat_4.3.0         

loaded via a namespace (and not attached):
  [1] plyr_1.8.8             igraph_1.4.1           lazyeval_0.2.2         sp_1.6-0               splines_4.1.3          listenv_0.9.0          scattermore_0.8       
  [8] digest_0.6.31          foreach_1.5.2          htmltools_0.5.4        fansi_1.0.4            magrittr_2.0.3         tensor_1.5             cluster_2.1.2         
 [15] doParallel_1.0.17      ROCR_1.0-11            remotes_2.4.2          globals_0.16.2         matrixStats_0.63.0     R.utils_2.12.2         spatstat.sparse_3.0-0 
 [22] colorspace_2.1-0       ggrepel_0.9.3          xfun_0.37              callr_3.7.3            crayon_1.5.2           jsonlite_1.8.4         progressr_0.13.0      
 [29] spatstat.data_3.0-0    survival_3.3-1         zoo_1.8-11             iterators_1.0.14       glue_1.6.2             polyclip_1.10-4        gtable_0.3.1          
 [36] leiden_0.4.3           GetoptLong_1.0.5       R.cache_0.16.0         future.apply_1.10.0    shape_1.4.6            BiocGenerics_0.40.0    abind_1.4-5           
 [43] scales_1.2.1           spatstat.random_3.1-3  miniUI_0.1.1.1         Rcpp_1.0.10            viridisLite_0.4.1      xtable_1.8-4           clue_0.3-64           
 [50] reticulate_1.28        stats4_4.1.3           htmlwidgets_1.6.1      rex_1.2.1              httr_1.4.5             RColorBrewer_1.1-3     ellipsis_0.3.2        
 [57] ica_1.0-3              pkgconfig_2.0.3        R.methodsS3_1.8.2      uwot_0.1.14            deldir_1.0-6           utf8_1.2.3             tidyselect_1.2.0      
 [64] rlang_1.0.6            reshape2_1.4.4         later_1.3.0            munsell_0.5.0          tools_4.1.3            cli_3.6.0              generics_0.1.3        
 [71] ggridges_0.5.4         stringr_1.5.0          fastmap_1.1.1          goftest_1.2-3          processx_3.8.0         knitr_1.42             fitdistrplus_1.1-8    
 [78] purrr_1.0.1            RANN_2.6.1             pbapply_1.7-0          future_1.31.0          nlme_3.1-155           mime_0.12              R.oo_1.25.0           
 [85] xml2_1.3.3             compiler_4.1.3         rstudioapi_0.14        plotly_4.10.1          png_0.1-8              spatstat.utils_3.0-1   tibble_3.1.8          
 [92] stringi_1.7.12         cyclocomp_1.1.0        ps_1.7.2               desc_1.4.2             lattice_0.20-45        Matrix_1.5-3           vctrs_0.5.2           
 [99] pillar_1.8.1           lifecycle_1.0.3        spatstat.geom_3.0-6    lmtest_0.9-40          GlobalOptions_0.1.2    RcppAnnoy_0.0.20       data.table_1.14.8     
[106] cowplot_1.1.1          irlba_2.3.5.1          httpuv_1.6.9           patchwork_1.1.2        R6_2.5.1               promises_1.2.0.1       KernSmooth_2.23-20    
[113] gridExtra_2.3          IRanges_2.28.0         parallelly_1.34.0      codetools_0.2-18       MASS_7.3-55            rprojroot_2.0.3        rjson_0.2.21          
[120] withr_2.5.0            sctransform_0.3.5      S4Vectors_0.32.4       parallel_4.1.3         Rtsne_0.16             spatstat.explore_3.0-6 shiny_1.7.4          


```{r}
# run style
style_file(here("src/seurat_scripts/Setbp1_DGEkidney_06.Rmd"))
# lintr was run as well
```
