---
title: "Setbp1_Integration_postSoup_03"
author: "Jordan Whitlock"
date: '2023-04-14'
output: html_document
---
This analysis follows 'Setbp1_QC_postSoup_01.Rmd'. 

### Loading in all required libraries
```{r}
set.seed(2178)
library(devtools)
library(magrittr)
library(SingleCellExperiment)
library(readr)
library(dplyr)
library(Matrix)
library(Seurat)
library(cowplot)
library(ggplot2)
library(tidyverse)
library(kBET)
library(DoubletFinder)
library(harmony)
library(clustree)
library(purrr)
library(edgeR)
library(here)
library(styler)
library(lintr)
```

# Load in the objects after QC
```{r}
load(here("data/filtered_setbp1_kidney_seurat_postSoup.Rdata"))
```

### Normalize FindVariableFeatures and Integrate -- Harmony

#### normalize data:
```{r normalize-kidney}
kidney_seurat_filtered <- NormalizeData(kidney_seurat_filtered)
```

#### feature selection:

```{r variable-features-kidney}
suppressWarnings(suppressMessages(kidney_seurat_filtered <- FindVariableFeatures(
  kidney_seurat_filtered,
  selection.method = "vst",
  nfeatures = 4000,
  verbose = FALSE,
  assay = "RNA"
))) # changed from std 2000 for nFeatures based on data distribution


top20_kidney <- head(VariableFeatures(kidney_seurat_filtered), 20)

png(
  filename = here("results/seurat/Integration_02/variable_features_kidney_postSoup.png"),
  width = 1000,
  height = 1000
)
LabelPoints(
  plot = VariableFeaturePlot(kidney_seurat_filtered),
  points = top20_kidney,
  repel = TRUE
)
dev.off()

save(top20_kidney, file = here("results/seurat/Integration_02/top20_kidney_postSoup.Rdata"))
```

#### scale data: z-score transformation

```{r scale-kidney}
kidney_seurat_filtered <- ScaleData(
  kidney_seurat_filtered,
  vars.to.regress = c("percent_mito", "nFeature_RNA"),
  assay = "RNA"
)
```

#### check dimensions with PCA
```{r PCA-elbow-kidney}
kidney_seurat_filtered <- RunPCA(
  kidney_seurat_filtered,
  npcs = 50,
  verbose = FALSE,
  approx = FALSE
) # has a seed set already in source code

png(
  filename = here("results/seurat/Integration_02/elbow_kidney_postSoup.png"),
  width = 1000,
  height = 1000
)
ElbowPlot(kidney_seurat_filtered, reduction = "pca", ndims = 50)
dev.off()
```

#### visualize PCs:
plotting the first few
```{r visualize-PCs-kidney}
png(
  filename = here("results/seurat/Integration_02/investigate_PCs_kidney_postSoup.png"),
  width = 1000,
  height = 1000
)
plot_grid(
  ncol = 3,
  DimPlot(
    kidney_seurat_filtered,
    reduction = "pca",
    group.by = "orig.ident",
    dims = 1:2
  ),
  DimPlot(
    kidney_seurat_filtered,
    reduction = "pca",
    group.by = "orig.ident",
    dims = 3:4
  ),
  DimPlot(
    kidney_seurat_filtered,
    reduction = "pca",
    group.by = "type",
    dims = 1:2
  ),
  DimPlot(
    kidney_seurat_filtered,
    reduction = "pca",
    group.by = "type",
    dims = 3:4
  )
)
dev.off()
```

#### visualize tSNE:
```{r tSNE-kidney}
kidney_seurat_filtered <- RunTSNE(
  kidney_seurat_filtered,
  reduction = "pca",
  dims = 1:30,
  perplexity = 30,
  max_iter = 1000,
  theta = 0.5,
  eta = 200,
  num_threads = 0,
  seed.use = 1
)

png(
  filename = here("results/seurat/Integration_02/tSNE_kidney_postSoup.png"),
  width = 1000,
  height = 1000
)
plot_grid(
  ncol = 1,
  DimPlot(kidney_seurat_filtered, reduction = "tsne", group.by = "orig.ident")
)
dev.off()
```

#### visualize the UMAP:
```{r kidney-UMAP}
kidney_seurat_filtered <- RunUMAP(
  kidney_seurat_filtered,
  reduction = "pca",
  dims = 1:30,
  n.components = 2,
  n.neighbors = 30,
  n.epochs = 200,
  min.dist = 0.3,
  learning.rate = 1,
  spread = 1 # note: has seed.use = 42 set as default
)

png(
  filename = here("results/seurat/Integration_02/UMAP_kidney_postSoup.png"),
  width = 1000,
  height = 1000
)
plot_grid(
  ncol = 1,
  DimPlot(kidney_seurat_filtered, reduction = "umap", group.by = "orig.ident") +
    ggplot2::ggtitle(label = "Kidney_UMAP_on_PCA")
)
dev.off()
```

### Integrate - using harmony

```{r kidney-harmony}
kidney_harmony <- RunHarmony(kidney_seurat_filtered,
  group.by.vars = "orig.ident", reduction = "pca",
  dims.use = 1:30, assay.use = "RNA"
)
```

#### UMAP in harmony:
Here we use all PCs computed from Harmony for UMAP calculation
```{r visualize-harmony-UMAP-kidney}
kidney_int <- kidney_seurat_filtered
kidney_int[["harmony"]] <- kidney_harmony[["harmony"]]
kidney_int <- RunUMAP(
  kidney_int,
  dims = 1:30,
  reduction = "harmony",
  reduction.name = "umap_harmony"
)
```

#### UMAP before and after harmony:

Visualizing the post-harmony integration UMAP compared to raw:

```{r compare-pre-postint-kidney}
png(
  filename = here("results/seurat/Integration_02/UMAP_prepost_integration_kidney_postSoup.png"),
  width = 1000,
  height = 1000
)
# kidney
p3 <- DimPlot(
  kidney_seurat_filtered,
  reduction = "umap",
  group.by = "orig.ident"
) +
  ggtitle("Kidney UMAP raw data")
p4 <- DimPlot(
  kidney_int,
  reduction = "umap_harmony",
  group.by = "orig.ident"
) +
  ggtitle("Kidney UMAP Harmony")
leg <- get_legend(p3)
gridExtra::grid.arrange(
  gridExtra::arrangeGrob(p3 + NoLegend() + NoAxes(), p4 + NoLegend() +
    NoAxes(), nrow = 2),
  leg,
  ncol = 2, widths = c(8, 2)
)
dev.off()
```

#### save integrated objects, can start from here:
```{r save-kidney-object}
save(kidney_int, file = here("data/integrated_setbp1_kidney_postSoup.Rdata"))
```

```{r}
proc.time()
```
     user    system   elapsed 
 3666.735  1046.945 43001.592 
 
```{r}
sessionInfo()
```
R version 4.1.3 (2022-03-10)
Platform: x86_64-pc-linux-gnu (64-bit)
Running under: Ubuntu 20.04.5 LTS

Matrix products: default
BLAS/LAPACK: /usr/lib/x86_64-linux-gnu/openblas-pthread/libopenblasp-r0.3.8.so

locale:
[1] C

attached base packages:
[1] stats4    stats     graphics  grDevices utils     datasets  methods   base     

other attached packages:
 [1] lintr_3.0.2                 styler_1.9.0                here_1.0.1                  edgeR_3.36.0                limma_3.50.3               
 [6] clustree_0.5.0              ggraph_2.1.0                harmony_0.1.1               Rcpp_1.0.10                 DoubletFinder_2.0.3        
[11] kBET_0.99.6                 lubridate_1.9.2             forcats_1.0.0               stringr_1.5.0               purrr_1.0.1                
[16] tidyr_1.3.0                 tibble_3.1.8                tidyverse_2.0.0             ggplot2_3.4.1               cowplot_1.1.1              
[21] SeuratObject_4.1.3          Seurat_4.3.0                Matrix_1.5-3                dplyr_1.1.0                 readr_2.1.4                
[26] SingleCellExperiment_1.16.0 SummarizedExperiment_1.24.0 Biobase_2.54.0              GenomicRanges_1.46.1        GenomeInfoDb_1.30.1        
[31] IRanges_2.28.0              S4Vectors_0.32.4            BiocGenerics_0.40.0         MatrixGenerics_1.6.0        matrixStats_0.63.0         
[36] magrittr_2.0.3              devtools_2.4.5              usethis_2.1.6              

loaded via a namespace (and not attached):
  [1] utf8_1.2.3             R.utils_2.12.2         spatstat.explore_3.0-6 reticulate_1.28        tidyselect_1.2.0       htmlwidgets_1.6.1     
  [7] grid_4.1.3             Rtsne_0.16             munsell_0.5.0          codetools_0.2-18       ica_1.0-3              future_1.31.0         
 [13] miniUI_0.1.1.1         withr_2.5.0            spatstat.random_3.1-3  colorspace_2.1-0       progressr_0.13.0       knitr_1.42            
 [19] rstudioapi_0.14        ROCR_1.0-11            tensor_1.5             listenv_0.9.0          GenomeInfoDbData_1.2.7 polyclip_1.10-4       
 [25] farver_2.1.1           rprojroot_2.0.3        parallelly_1.34.0      vctrs_0.5.2            generics_0.1.3         xfun_0.37             
 [31] timechange_0.2.0       R6_2.5.1               graphlayouts_0.8.4     locfit_1.5-9.7         rex_1.2.1              bitops_1.0-7          
 [37] spatstat.utils_3.0-1   cachem_1.0.7           DelayedArray_0.20.0    promises_1.2.0.1       scales_1.2.1           gtable_0.3.1          
 [43] globals_0.16.2         processx_3.8.0         goftest_1.2-3          tidygraph_1.2.3        rlang_1.0.6            cyclocomp_1.1.0       
 [49] splines_4.1.3          lazyeval_0.2.2         spatstat.geom_3.0-6    reshape2_1.4.4         abind_1.4-5            httpuv_1.6.9          
 [55] tools_4.1.3            ellipsis_0.3.2         RColorBrewer_1.1-3     sessioninfo_1.2.2      ggridges_0.5.4         plyr_1.8.8            
 [61] zlibbioc_1.40.0        RCurl_1.98-1.10        ps_1.7.2               prettyunits_1.1.1      deldir_1.0-6           pbapply_1.7-0         
 [67] viridis_0.6.2          urlchecker_1.0.1       zoo_1.8-11             ggrepel_0.9.3          cluster_2.1.2          fs_1.6.1              
 [73] data.table_1.14.8      scattermore_0.8        lmtest_0.9-40          RANN_2.6.1             fitdistrplus_1.1-8     R.cache_0.16.0        
 [79] pkgload_1.3.2          hms_1.1.2              patchwork_1.1.2        mime_0.12              xtable_1.8-4           gridExtra_2.3         
 [85] compiler_4.1.3         KernSmooth_2.23-20     crayon_1.5.2           R.oo_1.25.0            htmltools_0.5.4        later_1.3.0           
 [91] tzdb_0.3.0             tweenr_2.0.2           MASS_7.3-55            cli_3.6.0              R.methodsS3_1.8.2      parallel_4.1.3        
 [97] igraph_1.4.1           pkgconfig_2.0.3        sp_1.6-0               plotly_4.10.1          spatstat.sparse_3.0-0  xml2_1.3.3            
[103] XVector_0.34.0         callr_3.7.3            digest_0.6.31          sctransform_0.3.5      RcppAnnoy_0.0.20       spatstat.data_3.0-0   
[109] leiden_0.4.3           uwot_0.1.14            shiny_1.7.4            lifecycle_1.0.3        nlme_3.1-155           jsonlite_1.8.4        
[115] desc_1.4.2             viridisLite_0.4.1      fansi_1.0.4            pillar_1.8.1           lattice_0.20-45        fastmap_1.1.1         
[121] httr_1.4.5             pkgbuild_1.3.1         survival_3.3-1         glue_1.6.2             remotes_2.4.2          FNN_1.1.3.1           
[127] png_0.1-8              ggforce_0.4.1          stringi_1.7.12         profvis_0.3.7          memoise_2.0.1          irlba_2.3.5.1         
[133] future.apply_1.10.0   

```{r}
# run style
style_file(here("src/seurat_scripts/Setbp1_Integration_postSoup_02.Rmd"))
# lintr was run as well
```


This analysis is followed by 'Setbp1_Clustering_postSoup_03.Rmd'


