---
title: "Setbp1_Clustering_03"
author: "Jordan Whitlock"
date: '2022-09-07'
output:     
  html_document:
      toc: true
      toc_depth: 2
      toc_float: true
---

This analysis comes after 'Setbp1_Integration_02.Rmd' and precedes 'Setbp1_MarkersCortex_04.Rmd' & 'Setbp1_MarkersKidney_04.Rmd'.

# *Single-nuclei analysis: Clustering*
* Finding Neighbors
* Clustering with Leiden algorithm
* Identify stable resolution with clustree
* Re-visualize bimodality
* Set identity as new resolution

## Code: 

Accompanying dry lab notebook for this analysis is here: https://docs.google.com/document/d/1FO_Je9oxbBFF9S_k-8tVF2t4ba88O4C77sw1v7DIFec/edit?usp=sharing

### Setting up libraries and working path:

```{r}
getwd()
.libPaths() # "/usr/local/lib/R/site-library" "/usr/local/lib/R/library" "/home/jbarham3/R/x86_64-pc-linux-gnu-library/4.1"
```

```{r, message=FALSE}
# loading in all required libraries
library(devtools)
library(magrittr)
library(readr)
library(dplyr)
library(Seurat)
library(cowplot)
library(ggplot2)
library(tidyverse)
library(harmony)
library(clustree)
library(here)
library(styler)
library(lintr)
set.seed(2178)
```

### Clustering on the Harmony integrated data

#### load in the objects
```{r}
load(here("data/integrated_setbp1_kidney.Rdata"))
load(here("data/integrated_setbp1_cerebral.Rdata"))
```


#### check the active assay
It should be RNA, but need to make sure the reduction used for everything downstream is set to "harmony"
```{r}
kidney_int@active.assay # RNA
cerebral_int@active.assay # RNA
```

#### Find Neighbors and Clusters

NOTE: Use LEIDEN (algorithm = 4) NOT LOUVAIN (algorithm = 1)

# find neighbors
```{r}
kidney_int <- FindNeighbors(kidney_int, dims = 1:30, reduction = "harmony")
cerebral_int <- FindNeighbors(cerebral_int, dims = 1:30, reduction = "harmony")
```

### find clusters WITH LEIDEN (algorithm = 4)
received error (described here: https://docs.google.com/document/d/1FO_Je9oxbBFF9S_k-8tVF2t4ba88O4C77sw1v7DIFec/edit?usp=sharing), need to change matrix type.

If compute power is an issue, this can be run in a job on the Cheaha supercomputer in accompanying script 'Setbp1_Clustering_leiden_04.R' and 'Setbp1_Clustering_leiden_04' bash script
```{r}
# when running the below code within the Docker HPC environment and prompted about a miniconda install respond No
for (res in c(0.5, 0.6, 0.7, 0.8, 0.9, 1, 1.1, 1.3, 1.5, 1.7, 1.9, 2.0, 2.2)) {
  kidney_int <- FindClusters(kidney_int, graph.name = "RNA_snn", resolution = res, algorithm = 4, method = "igraph")
}

png(file = here("results/seurat/Clustering_03/kidneyint_clustree.png"),
    width = 1000,
    height = 2000)
clustree(kidney_int@meta.data, prefix = "RNA_snn_res.")
dev.off()

save(kidney_int, file = here("data/kidney_leiden.Rdata"))


for (res in c(0.5, 0.6, 0.7, 0.8, 0.9, 1, 1.1, 1.3, 1.5, 1.7, 1.9)) {
  cerebral_int <- FindClusters(cerebral_int, graph.name = "RNA_snn", resolution = res, algorithm = 4, method = "igraph")
}

png(file = here("results/seurat/Clustering_03/cerebralint_clustree.png"),
    width = 1000,
    height = 2000)
clustree(cerebral_int@meta.data, prefix = "RNA_snn_res.")
dev.off()


save(cerebral_int, file = here("data/cerebral_leiden.Rdata"))
```


### load in post-leiden objects:
```{r}
load(here("data/cerebral_leiden.Rdata")) # cerebral_int
load(here("data/kidney_leiden.Rdata")) # kidney_int
```

```{r}
# harmony
png(file = here("results/seurat/Clustering_03/kidneyint_res_preSoup.png"),
    width = 1000,
    height = 1000)
DimPlot(kidney_int, reduction = "umap_harmony", group.by = "RNA_snn_res.1.7") +
  ggtitle("kidney leiden_1.7")
dev.off()


png(file = here("results/seurat/Clustering_03/cerebralint_res.png"),
    width = 1000,
    height = 1000)
DimPlot(cerebral_int, reduction = "umap_harmony", group.by = "RNA_snn_res.1") +
  ggtitle("cerebral cortex leiden_1")
dev.off()
```

#### set the identity as Leiden with resolution 0.9 for kidney and 0.9 for cerbral cortex:
```{r}
select_resolution <- "RNA_snn_res.1"
cerebral_int <- SetIdent(cerebral_int, value = select_resolution)

select_resolution <- "RNA_snn_res.1.7"
kidney_int <- SetIdent(kidney_int, value = select_resolution)
```

#### plot the clustering
```{r}
png(file = here("results/seurat/Clustering_03/cerebralint_UMAP_postharmony.png"),
    width = 2000,
    height = 1000)
plot_grid(ncol = 3, 
          DimPlot(kidney_int,
                  label = TRUE,
                  reduction = "umap_harmony") + NoAxes(), 
          DimPlot(kidney_int, 
                  group.by = "orig.ident", 
                  reduction = "umap_harmony") + NoAxes(), 
          DimPlot(kidney_int, 
                  group.by = "type", 
                  reduction = "umap_harmony") + NoAxes())
dev.off()

png(file = here("results/seurat/Clustering_03/kidneyint_UMAP_postharmony.png"),
    width = 2000,
    height = 1000)
plot_grid(ncol = 3, 
          DimPlot(cerebral_int, 
                  label = TRUE, 
                  reduction = "umap_harmony") + NoAxes(), 
          DimPlot(cerebral_int, 
                  group.by = "orig.ident", 
                  reduction = "umap_harmony") + NoAxes(), 
          DimPlot(cerebral_int, 
                  group.by = "type", 
                  reduction = "umap_harmony") + NoAxes())
```

```{r}
proc.time()
```
user    system   elapsed 
 8871.874   480.916 14890.473 


```{r}
sessionInfo()
```
R version 4.1.3 (2022-03-10)
Platform: x86_64-pc-linux-gnu (64-bit)
Running under: Ubuntu 20.04.5 LTS

Matrix products: default
BLAS/LAPACK: /usr/lib/x86_64-linux-gnu/openblas-pthread/libopenblasp-r0.3.8.so

locale:
[1] C

attached base packages:
[1] stats     graphics  grDevices utils     datasets  methods   base     

other attached packages:
 [1] lintr_3.0.2        styler_1.9.0       here_1.0.1         clustree_0.5.0     ggraph_2.1.0       harmony_0.1.1      Rcpp_1.0.10        lubridate_1.9.2    forcats_1.0.0     
[10] stringr_1.5.0      purrr_1.0.1        tidyr_1.3.0        tibble_3.1.8       tidyverse_2.0.0    ggplot2_3.4.1      cowplot_1.1.1      SeuratObject_4.1.3 Seurat_4.3.0      
[19] dplyr_1.1.0        readr_2.1.4        magrittr_2.0.3     devtools_2.4.5     usethis_2.1.6     

loaded via a namespace (and not attached):
  [1] plyr_1.8.8             igraph_1.4.1           lazyeval_0.2.2         sp_1.6-0               splines_4.1.3          listenv_0.9.0          scattermore_0.8       
  [8] digest_0.6.31          htmltools_0.5.4        viridis_0.6.2          fansi_1.0.4            memoise_2.0.1          tensor_1.5             cluster_2.1.2         
 [15] ROCR_1.0-11            tzdb_0.3.0             remotes_2.4.2          globals_0.16.2         graphlayouts_0.8.4     matrixStats_0.63.0     R.utils_2.12.2        
 [22] timechange_0.2.0       spatstat.sparse_3.0-0  prettyunits_1.1.1      colorspace_2.1-0       ggrepel_0.9.3          xfun_0.37              callr_3.7.3           
 [29] crayon_1.5.2           jsonlite_1.8.4         progressr_0.13.0       spatstat.data_3.0-0    survival_3.3-1         zoo_1.8-11             glue_1.6.2            
 [36] polyclip_1.10-4        gtable_0.3.1           leiden_0.4.3           R.cache_0.16.0         pkgbuild_1.3.1         future.apply_1.10.0    abind_1.4-5           
 [43] scales_1.2.1           spatstat.random_3.1-3  miniUI_0.1.1.1         viridisLite_0.4.1      xtable_1.8-4           reticulate_1.28        profvis_0.3.7         
 [50] rex_1.2.1              htmlwidgets_1.6.1      httr_1.4.5             RColorBrewer_1.1-3     ellipsis_0.3.2         ica_1.0-3              R.methodsS3_1.8.2     
 [57] urlchecker_1.0.1       pkgconfig_2.0.3        farver_2.1.1           uwot_0.1.14            deldir_1.0-6           utf8_1.2.3             tidyselect_1.2.0      
 [64] rlang_1.0.6            reshape2_1.4.4         later_1.3.0            munsell_0.5.0          tools_4.1.3            cachem_1.0.7           cli_3.6.0             
 [71] generics_0.1.3         ggridges_0.5.4         fastmap_1.1.1          goftest_1.2-3          processx_3.8.0         knitr_1.42             fs_1.6.1              
 [78] fitdistrplus_1.1-8     tidygraph_1.2.3        RANN_2.6.1             pbapply_1.7-0          future_1.31.0          nlme_3.1-155           mime_0.12             
 [85] R.oo_1.25.0            xml2_1.3.3             compiler_4.1.3         rstudioapi_0.14        plotly_4.10.1          png_0.1-8              spatstat.utils_3.0-1  
 [92] tweenr_2.0.2           stringi_1.7.12         cyclocomp_1.1.0        ps_1.7.2               desc_1.4.2             lattice_0.20-45        Matrix_1.5-3          
 [99] vctrs_0.5.2            pillar_1.8.1           lifecycle_1.0.3        spatstat.geom_3.0-6    lmtest_0.9-40          RcppAnnoy_0.0.20       data.table_1.14.8     
[106] irlba_2.3.5.1          httpuv_1.6.9           patchwork_1.1.2        R6_2.5.1               promises_1.2.0.1       KernSmooth_2.23-20     gridExtra_2.3         
[113] parallelly_1.34.0      sessioninfo_1.2.2      codetools_0.2-18       MASS_7.3-55            pkgload_1.3.2          rprojroot_2.0.3        withr_2.5.0           
[120] sctransform_0.3.5      parallel_4.1.3         hms_1.1.2              grid_4.1.3             Rtsne_0.16             spatstat.explore_3.0-6 ggforce_0.4.1         
[127] shiny_1.7.4           

```{r}
# run style
style_file(here("src/seurat_scripts/Setbp1_Clustering_03.Rmd"))
# lintr was run as well
```
