---
title: "Setbp1_Integration_03"
author: "Jordan Whitlock"
date: '2022-09-06'
output:     
  html_document:
      toc: true
      toc_depth: 2
      toc_float: true
---

This analysis follows 'Setbp1_DoubletFinder_02.Rmd'. 

### Loading in all required libraries
```{r}
set.seed(2178)
library(devtools)
library(magrittr)
library(SingleCellExperiment)
library(readr)
library(dplyr)
library(Matrix)
library(Seurat)
library(cowplot)
library(ggplot2)
library(tidyverse)
library(kBET)
library(DoubletFinder)
library(harmony)
library(clustree)
library(purrr)
library(edgeR)
library(here)
library(styler)
library(lintr)
```


# Load in the object after doublets are removed
```{r load-cortex-object}
load(here("data/nodoublet_setbp1_cerebral_seurat.Rdata")) # cerebral_seurat_filtered
```

```{r load-kidney-object}
load(here("data/nodoublet_setbp1_kidney_seurat.Rdata")) # kidney_seurat_filtered
```

### Normalize FindVariableFeatures and Integrate -- Harmony

#### normalize data:
```{r normalize-kidney}
kidney_seurat_filtered <- NormalizeData(kidney_seurat_filtered)
```

```{r normalize-cortex}
cerebral_seurat_filtered <- NormalizeData(cerebral_seurat_filtered)
```


#### feature selection:

```{r variable-features-kidney}
suppressWarnings(suppressMessages(kidney_seurat_filtered <- FindVariableFeatures(
  kidney_seurat_filtered,
  selection.method = "vst",
  nfeatures = 4000,
  verbose = FALSE,
  assay = "RNA"
))) # changed from std 2000 for nFeatures based on data distribution


top20_kidney <- head(VariableFeatures(kidney_seurat_filtered), 20)

png(
  filename = here("results/seurat/Integration_03/variable_features_kidney.png"),
  width = 1000,
  height = 1000
)
LabelPoints(
  plot = VariableFeaturePlot(kidney_seurat_filtered),
  points = top20_kidney,
  repel = TRUE
)
dev.off()

save(top20_kidney, file = here("results/seurat/Integration_03/top20_kidney.Rdata"))
```

```{r variable-features-cortex}
suppressWarnings(suppressMessages(cerebral_seurat_filtered <- FindVariableFeatures(
  cerebral_seurat_filtered,
  selection.method = "vst",
  nfeatures = 4000,
  verbose = FALSE,
  assay = "RNA"
))) # changed from std 2000 for nFeatures based on data distribution

top20_cerebral <- head(VariableFeatures(cerebral_seurat_filtered), 20)

png(
  filename = here("results/seurat/Integration_03/variable_features_cerebral.png"),
  width = 1000,
  height = 1000
)
LabelPoints(
  plot = VariableFeaturePlot(cerebral_seurat_filtered),
  points = top20_cerebral,
  repel = TRUE
)
dev.off()

save(top20_cerebral, file = here("results/seurat/Integration_03/top20_cerebral.Rdata"))
```


#### scale data: z-score transformation

```{r scale-kidney}
kidney_seurat_filtered <- ScaleData(
  kidney_seurat_filtered,
  vars.to.regress = c("percent_mito", "nFeature_RNA"),
  assay = "RNA"
)
```


```{r scale-cortex}
cerebral_seurat_filtered <- ScaleData(
  cerebral_seurat_filtered,
  vars.to.regress = c("percent_mito", "nFeature_RNA"),
  assay = "RNA"
)
```

#### check dimensions with PCA
```{r PCA-elbow-cortex}
cerebral_seurat_filtered <- RunPCA(
  cerebral_seurat_filtered,
  npcs = 50,
  verbose = FALSE, 
  approx = FALSE
) # has a seed already set in source code

png(
  filename = here("results/seurat/Integration_03/elbow_cerebral.png"),
  width = 1000,
  height = 1000
)
ElbowPlot(cerebral_seurat_filtered, reduction = "pca", ndims = 50)
dev.off()
```

```{r PCA-elbow-kidney}
kidney_seurat_filtered <- RunPCA(
  kidney_seurat_filtered,
  npcs = 50,
  verbose = FALSE, 
  approx = FALSE
) # has a seed set already in source code

png(
  filename = here("results/seurat/Integration_03/elbow_kidney.png"),
  width = 1000,
  height = 1000
)
ElbowPlot(kidney_seurat_filtered, reduction = "pca", ndims = 50)
dev.off()
```

#### visualize PCs:
plotting the first few
```{r visualize-PCs-cortex}
png(
  filename = here("results/seurat/Integration_03/investigate_PCs_cerebral.png"),
  width = 1000,
  height = 1000
)
plot_grid(
  ncol = 3,
  DimPlot(
    cerebral_seurat_filtered,
    reduction = "pca", group.by = "orig.ident",
    dims = 1:2
  ),
  DimPlot(
    cerebral_seurat_filtered,
    reduction = "pca",
    group.by = "orig.ident",
    dims = 3:4
  ),
  DimPlot(
    cerebral_seurat_filtered,
    reduction = "pca",
    group.by = "orig.ident",
    dims = 29:30
  )
)
dev.off()
```

```{r visualize-PCs-kidney}
png(
  filename = here("results/seurat/Integration_03/investigate_PCs_kidney.png"),
  width = 1000,
  height = 1000
)
plot_grid(
  ncol = 3,
  DimPlot(
    kidney_seurat_filtered,
    reduction = "pca",
    group.by = "orig.ident",
    dims = 1:2
  ),
  DimPlot(
    kidney_seurat_filtered,
    reduction = "pca",
    group.by = "orig.ident",
    dims = 3:4
  ),
  DimPlot(
    kidney_seurat_filtered,
    reduction = "pca",
    group.by = "orig.ident",
    dims = 29:30
  )
)
dev.off()
```

#### visualize tSNE:
```{r tSNE-cortex}
cerebral_seurat_filtered <- RunTSNE(
  cerebral_seurat_filtered,
  reduction = "pca",
  dims = 1:30,
  perplexity = 30,
  max_iter = 1000,
  theta = 0.5,
  eta = 200,
  num_threads = 0
)

png(
  filename = here("results/seurat/Integration_03/tSNE_cerebral.png"),
  width = 1000,
  height = 1000
)
plot_grid(
  ncol = 1,
  DimPlot(cerebral_seurat_filtered, reduction = "tsne", group.by = "orig.ident")
)
dev.off()
```

```{r tSNE-kidney}
kidney_seurat_filtered <- RunTSNE(
  kidney_seurat_filtered,
  reduction = "pca",
  dims = 1:30,
  perplexity = 30,
  max_iter = 1000,
  theta = 0.5,
  eta = 200,
  num_threads = 0,
  seed.use = 1
)

png(
  filename = here("results/seurat/Integration_03/tSNE_kidney.png"),
  width = 1000,
  height = 1000
)
plot_grid(
  ncol = 1,
  DimPlot(kidney_seurat_filtered, reduction = "tsne", group.by = "orig.ident")
)
dev.off()
```

#### visualize the UMAP:
```{r kidney-UMAP}
kidney_seurat_filtered <- RunUMAP(
  kidney_seurat_filtered,
  reduction = "pca",
  dims = 1:30,
  n.components = 2,
  n.neighbors = 30,
  n.epochs = 200,
  min.dist = 0.3,
  learning.rate = 1,
  spread = 1 # note: has seed.use = 42 set as default
)

png(
  filename = here("results/seurat/Integration_03/UMAP_kidney.png"),
  width = 1000,
  height = 1000
)
plot_grid(
  ncol = 1,
  DimPlot(kidney_seurat_filtered, reduction = "umap", group.by = "orig.ident") +
    ggplot2::ggtitle(label = "Kidney_UMAP_on_PCA")
)
dev.off()
```

```{r cortex-UMAP}
cerebral_seurat_filtered <- RunUMAP(
  cerebral_seurat_filtered,
  reduction = "pca",
  dims = 1:30,
  n.components = 2,
  n.neighbors = 30,
  n.epochs = 200,
  min.dist = 0.3, learning.rate = 1, spread = 1
)

png(
  filename = here("results/seurat/Integration_03/UMAP_cerebral.png"),
  width = 1000,
  height = 1000
)
plot_grid(
  ncol = 1,
  DimPlot(cerebral_seurat_filtered, reduction = "umap", group.by = "orig.ident") +
    ggplot2::ggtitle(label = "CerebralCortex_UMAP_on_PCA")
)
dev.off()
```


### Integrate - using harmony

```{r kidney-harmony}
kidney_harmony <- RunHarmony(kidney_seurat_filtered,
  group.by.vars = "orig.ident", reduction = "pca",
  dims.use = 1:30, assay.use = "RNA"
)
```

```{r cortex-harmony}
cerebral_harmony <- RunHarmony(cerebral_seurat_filtered,
  group.by.vars = "orig.ident", reduction = "pca",
  dims.use = 1:30, assay.use = "RNA"
)
```

#### UMAP in harmony:
Here we use all PCs computed from Harmony for UMAP calculation
```{r visualize-harmony-UMAP-kidney}
kidney_int <- kidney_seurat_filtered
kidney_int[["harmony"]] <- kidney_harmony[["harmony"]]
kidney_int <- RunUMAP(
  kidney_int,
  dims = 1:30,
  reduction = "harmony",
  reduction.name = "umap_harmony"
)
```

```{r visualize-harmony-UMAP-cortex}
cerebral_int <- cerebral_seurat_filtered
cerebral_int[["harmony"]] <- cerebral_harmony[["harmony"]]
cerebral_int <- RunUMAP(
  cerebral_int,
  dims = 1:30,
  reduction = "harmony",
  reduction.name = "umap_harmony"
)
```


#### UMAP before and after harmony:

Visualizing the post-harmony integration UMAP compared to raw:
```{r compare-pre-postint-cortex}
png(
  filename = here("results/seurat/Integration_03/UMAP_prepost_integration_cerebral.png"),
  width = 1000,
  height = 1000
)
# cerebral
p1 <- DimPlot(
  cerebral_seurat_filtered,
  reduction = "umap",
  group.by = "orig.ident"
) + ggtitle("Cerebral UMAP raw_data")
p2 <- DimPlot(
  cerebral_int,
  reduction = "umap_harmony",
  group.by = "orig.ident"
) + ggtitle("Cerebral UMAP Harmony")
leg <- get_legend(p1)
gridExtra::grid.arrange(
  gridExtra::arrangeGrob(p1 + NoLegend() + NoAxes(), p2 + NoLegend() +
    NoAxes(), nrow = 2),
  leg,
  ncol = 2, widths = c(8, 2)
)
dev.off()
```

```{r compare-pre-postint-kidney}
png(
  filename = here("results/seurat/Integration_03/UMAP_prepost_integration_kidney.png"),
  width = 1000,
  height = 1000
)
# kidney
p3 <- DimPlot(
  kidney_seurat_filtered,
  reduction = "umap",
  group.by = "orig.ident"
) +
  ggtitle("Kidney UMAP raw data")
p4 <- DimPlot(
  kidney_int,
  reduction = "umap_harmony",
  group.by = "orig.ident"
) +
  ggtitle("Kidney UMAP Harmony")
leg <- get_legend(p3)
gridExtra::grid.arrange(
  gridExtra::arrangeGrob(p3 + NoLegend() + NoAxes(), p4 + NoLegend() +
    NoAxes(), nrow = 2),
  leg,
  ncol = 2, widths = c(8, 2)
)
dev.off()
```

#### save integrated objects, can start from here:
```{r save-kidney-object}
save(kidney_int, file = here("data/integrated_setbp1_kidney.Rdata"))
```

```{r save-cortex-object}
save(cerebral_int, file = here("data/integrated_setbp1_cerebral.Rdata"))
```

```{r}
sessionInfo()
```
R version 4.1.3 (2022-03-10)
Platform: x86_64-pc-linux-gnu (64-bit)
Running under: Ubuntu 20.04.5 LTS

Matrix products: default
BLAS/LAPACK: /usr/lib/x86_64-linux-gnu/openblas-pthread/libopenblasp-r0.3.8.so

locale:
[1] C

attached base packages:
[1] stats4    stats     graphics  grDevices utils     datasets  methods   base     

other attached packages:
 [1] lintr_3.0.2                 styler_1.9.0                here_1.0.1                  edgeR_3.36.0                limma_3.50.3                clustree_0.5.0             
 [7] ggraph_2.1.0                harmony_0.1.1               Rcpp_1.0.10                 DoubletFinder_2.0.3         kBET_0.99.6                 lubridate_1.9.2            
[13] forcats_1.0.0               stringr_1.5.0               purrr_1.0.1                 tidyr_1.3.0                 tibble_3.1.8                tidyverse_2.0.0            
[19] ggplot2_3.4.1               cowplot_1.1.1               SeuratObject_4.1.3          Seurat_4.3.0                Matrix_1.5-3                dplyr_1.1.0                
[25] readr_2.1.4                 SingleCellExperiment_1.16.0 SummarizedExperiment_1.24.0 Biobase_2.54.0              GenomicRanges_1.46.1        GenomeInfoDb_1.30.1        
[31] IRanges_2.28.0              S4Vectors_0.32.4            BiocGenerics_0.40.0         MatrixGenerics_1.6.0        matrixStats_0.63.0          magrittr_2.0.3             
[37] devtools_2.4.5              usethis_2.1.6              

loaded via a namespace (and not attached):
  [1] utf8_1.2.3             R.utils_2.12.2         spatstat.explore_3.0-6 reticulate_1.28        tidyselect_1.2.0       htmlwidgets_1.6.1      grid_4.1.3            
  [8] Rtsne_0.16             munsell_0.5.0          codetools_0.2-18       ica_1.0-3              future_1.31.0          miniUI_0.1.1.1         withr_2.5.0           
 [15] spatstat.random_3.1-3  colorspace_2.1-0       progressr_0.13.0       knitr_1.42             rstudioapi_0.14        ROCR_1.0-11            tensor_1.5            
 [22] listenv_0.9.0          GenomeInfoDbData_1.2.7 polyclip_1.10-4        farver_2.1.1           rprojroot_2.0.3        parallelly_1.34.0      vctrs_0.5.2           
 [29] generics_0.1.3         xfun_0.37              timechange_0.2.0       R6_2.5.1               graphlayouts_0.8.4     locfit_1.5-9.7         rex_1.2.1             
 [36] bitops_1.0-7           spatstat.utils_3.0-1   cachem_1.0.7           DelayedArray_0.20.0    promises_1.2.0.1       scales_1.2.1           gtable_0.3.1          
 [43] globals_0.16.2         processx_3.8.0         goftest_1.2-3          tidygraph_1.2.3        rlang_1.0.6            cyclocomp_1.1.0        splines_4.1.3         
 [50] lazyeval_0.2.2         spatstat.geom_3.0-6    yaml_2.3.7             reshape2_1.4.4         abind_1.4-5            httpuv_1.6.9           tools_4.1.3           
 [57] ellipsis_0.3.2         RColorBrewer_1.1-3     sessioninfo_1.2.2      ggridges_0.5.4         plyr_1.8.8             zlibbioc_1.40.0        RCurl_1.98-1.10       
 [64] ps_1.7.2               prettyunits_1.1.1      deldir_1.0-6           pbapply_1.7-0          viridis_0.6.2          urlchecker_1.0.1       zoo_1.8-11            
 [71] ggrepel_0.9.3          cluster_2.1.2          fs_1.6.1               data.table_1.14.8      scattermore_0.8        lmtest_0.9-40          RANN_2.6.1            
 [78] fitdistrplus_1.1-8     R.cache_0.16.0         pkgload_1.3.2          hms_1.1.2              patchwork_1.1.2        mime_0.12              evaluate_0.20         
 [85] xtable_1.8-4           gridExtra_2.3          compiler_4.1.3         KernSmooth_2.23-20     crayon_1.5.2           R.oo_1.25.0            htmltools_0.5.4       
 [92] later_1.3.0            tzdb_0.3.0             tweenr_2.0.2           MASS_7.3-55            cli_3.6.0              R.methodsS3_1.8.2      parallel_4.1.3        
 [99] igraph_1.4.1           pkgconfig_2.0.3        sp_1.6-0               plotly_4.10.1          spatstat.sparse_3.0-0  xml2_1.3.3             XVector_0.34.0        
[106] callr_3.7.3            digest_0.6.31          sctransform_0.3.5      RcppAnnoy_0.0.20       spatstat.data_3.0-0    rmarkdown_2.20         leiden_0.4.3          
[113] uwot_0.1.14            shiny_1.7.4            lifecycle_1.0.3        nlme_3.1-155           jsonlite_1.8.4         desc_1.4.2             viridisLite_0.4.1     
[120] fansi_1.0.4            pillar_1.8.1           lattice_0.20-45        fastmap_1.1.1          httr_1.4.5             pkgbuild_1.3.1         survival_3.3-1        
[127] glue_1.6.2             remotes_2.4.2          FNN_1.1.3.1            png_0.1-8              ggforce_0.4.1          stringi_1.7.12         profvis_0.3.7         
[134] memoise_2.0.1          irlba_2.3.5.1          future.apply_1.10.0   

[Show in New Window] 

```{r}
# run style
style_file(here("src/seurat_scripts/Setbp1_Integration_03.Rmd"))
# lintr was run as well
```


This analysis is followed by 'Setbp1_Clustering_04.Rmd'


