---
title: "Setbp1_DiffTargetingCortex_01"
output: html_document
date: '2023-02-13'
---

This R markdown script is a reanalysis of the gene targeting analysis performed in "Setbp1_AllCortex_PANDAComparison_01.Rmd" in order to determine the impact of edge weight on the gene/TF targeting scores from the regulatory networks. In an effort to determine this, all networks below were filtered to only include positive weights, whereas in the previous script both positives and negatives were included. 

Targeting code was originially implemented here: http://netbooks.networkmedicine.org/user/b04f15be-cbd3-44df-932f-6e9026018fd0/notebooks/netZooR/lioness_tcga_targeting_netbook.ipynb

#load in required libraries & functions
```{r loading-libraries-functions, results='hide'}
set.seed(2178)
library(netZooR)
library(data.table)
library(dplyr)
library(limma)
library(tidyr)
library(ggpubr)
library(ggplot2)
library(ComplexHeatmap)
library(RColorBrewer)
library(magrittr)
library(circlize)
library(readr)
library(gprofiler2)
library(here)
source(here("src/functions/functions.R"))
library(styler)
library(lintr)
```

## load in PANDA object containing list of all cell types from Setbp1 data
Here each item returned by the for loop represents a PANDA object for a given cell type that contains the regNet, coopNet, and coregNet for that cell type.
```{r loading-object, results='hide'}
files <- list.files(here("results/PANDA"), pattern = "cortexexpression_PANDA.Rdata", full.names = TRUE) # grabbing cortex samples
for (i in files) {
  load(i)
  regNet <- pandaResults
  regNet <- regNet@regNet
  name <- sub(here("results/PANDA/"), "", i)
  name <- sub("cortexexpression_PANDA.Rdata*", "", name)
  assign(paste0(name, ".regNet"), regNet)
  rm(pandaResults)
  rm(regNet)
}
```

```{r}
files # 16 total
```

## Visualizing edge weight distributions filtered and unfiltered
```{r visualize-distribution-of-unfiltered-regNet-scores}
# grabbing object names
merged_list <- list()
networks <- basename(files)
networks <- sub("cortexexpression_PANDA.Rdata.*", "", networks)

# extracting scores
for (i in networks) {
  if (grepl("_heterozygous", i) == TRUE) {
    melt <- as.data.frame(reshape2::melt(get(paste0(i, ".regNet")),
      varnames = c("TF", "gene"),
      value.name = "het_edge_weight"
    )) # melting dataframe
    melt <- as.data.frame(melt[, 3])
    colnames(melt) <- "edge_weight"
    melt$cell_type <- sub("_.*", "", i)
    merged_list[i] <- melt
  } else if (grepl("_control", i) == TRUE) {
    melt <- as.data.frame(reshape2::melt(get(paste0(i, ".regNet")),
      varnames = c("TF", "gene"),
      value.name = "ctrl_edge_weight"
    )) # melting dataframe
    melt <- as.data.frame(melt[, 3])
    colnames(melt) <- "edge_weight"
    melt$cell_type <- sub("_.*", "", i)
    merged_list[i] <- melt
  }
}

# wrangling data
merged_control <- data.frame(merged_list[grep("_control", names(merged_list))]) %>% pivot_longer(., names_to = "cell_type", values_to = "edge_weight", cols = 1:8) # need to adjust column number for different data

# plotting distribution
png(
  filename = here("results/diff_targeting/cortex_regNet_edgeweight_dist_ctrl.png"),
  width = 1000,
  height = 500
)
ggplot(data = merged_control, aes(
  x = edge_weight,
  group = cell_type, color = cell_type
)) +
  geom_density(adjust = 1.5, alpha = .4) +
  ggtitle("Distribution of regNet edge weights in control Cerebral Cortex") +
  theme_bw() +
  scale_color_manual(values = c("#6968B4", "#DC71FA", "#0a2f6f", "#00BBDB", "#C216B7", "#027461", "#7997FF", "#00C1AA"))
dev.off()

# wrangling data
merged_het <- data.frame(merged_list[grep("_heterozygous", names(merged_list))]) %>% pivot_longer(., names_to = "cell_type", values_to = "edge_weight", cols = 1:8) # need to adjust column number for different data

# plotting distribution
png(
  filename = here("results/diff_targeting/cortex_regNet_edgeweight_dist_het.png"),
  width = 1000,
  height = 500
)
ggplot(data = merged_het, aes(
  x = edge_weight,
  group = cell_type, color = cell_type
)) +
  geom_density(adjust = 1.5, alpha = .4) +
  ggtitle("Distribution of regNet edge weights in S858R Cerebral Cortex") +
  theme_bw() +
  scale_color_manual(values = c("#6968B4", "#DC71FA", "#0a2f6f", "#00BBDB", "#C216B7", "#027461", "#7997FF", "#00C1AA"))
dev.off()
```

Note: all negative edge weights were changed to 0
```{r visualize-distribution-of-filtered-het-regNet-scores}
# grabbing only positive edge weights
merged_het_pos <- merged_het[merged_het$edge_weight > 0, ]
merged_control_pos <- merged_control[merged_control$edge_weight > 0, ]

# plotting only positive het edge weights
png(
  filename = here("results/diff_targeting/cortex_regNet_edgeweight_filtered_dist_het.png"),
  width = 1000,
  height = 500
)
ggplot(data = merged_het_pos, aes(x = edge_weight, group = cell_type, color = cell_type)) +
  geom_density(adjust = 1.5, alpha = .4) +
  ggtitle("Distribution of regNet edge weights > 0 in S858R Cerebral Cortex") +
  theme_bw() +
  scale_color_manual(values = c("#6968B4", "#DC71FA", "#0a2f6f", "#00BBDB", "#C216B7", "#027461", "#7997FF", "#00C1AA"))
dev.off()

# plotting only positive ctrl edge weights
png(
  filename = here("results/diff_targeting/cortex_regNet_edgeweight_filtered_dist_ctrl.png"),
  width = 1000,
  height = 500
)
ggplot(data = merged_control_pos, aes(x = edge_weight, group = cell_type, color = cell_type)) +
  geom_density(adjust = 1.5, alpha = .4) +
  ggtitle("Distribution of regNet edge weights >0 in control Cerebral Cortex") +
  theme_bw() +
  scale_color_manual(values = c("#6968B4", "#DC71FA", "#0a2f6f", "#00BBDB", "#C216B7", "#027461", "#7997FF", "#00C1AA"))
dev.off()
```

## wrangling data to proper format, filtering for only positive edge weights, and calculating targeting scores:
Gene targeting score, a numerical score representing the extent to which a gene is targeted by TFs in a given biological context. The gene targeting score is calculated by summing the weights of all inbound regulatory edges for a gene or outbound edges for a TF (https://doi.org/10.3389/fgene.2021.649942). 
* regNetmatrix: The PANDA regNet output matrix of TF and Genes and their edge weights from running pandaR 
* variable_name: Here this is the cell type, or variable you would like the targeting score matrix stored as in your global environment 
* edge_weight_name: This is the name of the edge_weight column in your matrix (options: het_edge_weight, ctrl_edge_weight)
* condition: The group the data is from, in this instance it is ctrl or het.

```{r calc-targeting}
for (i in networks) {
  if (grepl("_control", i) == TRUE) {
    targetingCalc(regNetmatrix = get(paste0(i, ".regNet")), variable_name = sub("_control.*", "", i), edge_weight_name = ctrl_edge_weight, condition = "ctrl")
  } else if (grepl("_heterozygous", i) == TRUE) {
    targetingCalc(regNetmatrix = get(paste0(i, ".regNet")), variable_name = sub("_heterozygous.*", "", i), edge_weight_name = het_edge_weight, condition = "het")
  }
}
```

Confirm targeting is being calculated as suspected by validating with alternate method. Distributions are the same when calculated by aggregate or sums:
```{r comparing-alternate-targeting-score-calc}
pdf(here("results/diff_targeting/cortex_alt_targeting_calc.pdf"))
# HETEROZYGOUS----------------------
# visualize current distribution
hist(opcs_gene_targeting_het$het_edge_weight_pos)

# checking alt targeting calc method distribution matches
dim(opcs_heterozygous.regNet) # 1080 22033
pos <- as.data.frame(pmax(opcs_heterozygous.regNet, 0))
TF_t_het <- as.data.frame(rowSums(pos))
TF_t_het$TF <- rownames(pos)
hist(TF_t_het$`rowSums(pos)`)
G_t_het <- as.data.frame(colSums(pos))
G_t_het$gene <- colnames(pos)
hist(G_t_het$`colSums(pos)`)

# CONTROL----------------------
# visualize current distribution
hist(opcs_gene_targeting_ctrl$ctrl_edge_weight_pos)

# checking alt targeting calc method distribution matches
dim(opcs_control.regNet) # 1080 22033
pos <- as.data.frame(pmax(opcs_control.regNet, 0))
TF_t_ctrl <- as.data.frame(rowSums(pos))
TF_t_ctrl$TF <- rownames(pos)
hist(TF_t_ctrl$`rowSums(pos)`)
G_t_ctrl <- as.data.frame(colSums(pos))
G_t_ctrl$gene <- colnames(pos)
hist(G_t_ctrl$`colSums(pos)`)
dev.off()
```

## combine ctrl and het to into single data frame 
```{r wrangling-data-to-combine}
cell_names <- networks[grep("_control", networks)] %>% sub("_control.*", "", .)
cell_names # 17 total for cortex, note: this will not work if your cell types are different across conditions

targeting_list <- list() # generating empty list

for (i in cell_names) {
  # TF targeting
  x <- get(paste0(i, "_TF_targeting_ctrl"))
  y <- get(paste0(i, "_TF_targeting_het"))
  TF_targeting <- merge(x, y, by = "TF") # combine dataframes
  targeting_list[[paste0("TF_targeting_", i)]] <- TF_targeting

  # gene targeting
  x <- get(paste0(i, "_gene_targeting_ctrl"))
  y <- get(paste0(i, "_gene_targeting_het"))
  gene_targeting <- merge(x, y, by = "gene") # combine dataframes
  targeting_list[[paste0("Gene_targeting_", i)]] <- gene_targeting
}
```
## setting heatmap data and setbp1 input needed for the rest of the script
```{r heatmap-color-annotation}
# setting annotation colors
annotation_colors <- list("cell_type" = c("excitatory_neurons" = "#DC71FA", "inhibitory_neurons" = "#00BBDB", "oligodendrocytes" = "#7997FF", "astrocytes" = "#6968B4", "microglia" = "#C216B7", "opcs" = "#00C1AA", "mural" = "#027461", "fibro_cortex" = "#0a2f6f"))
```

```{r loading-geneset}
setbp1_genes <- read.csv(here("results/seurat/setbp1_targets.csv"))
setbp1_genes <- setbp1_genes[, -1] %>% as.vector()
setbp1_genes <- append(setbp1_genes, "Setbp1")
```

```{r constructing-het-targeting-matrix}
# grabbing all het targeting scores for TFs
TF_targeting_all_het <- targeting_list[grep("TF_", names(targeting_list))] %>%
  purrr::reduce(full_join, by = "TF") %>%
  select(., contains(c("TF", "het_edge_weight_pos")))

# grabbing colnames and setting for new dataframe generated above
names <- names(targeting_list[grep("TF_", names(targeting_list))]) %>%
  sub(".*TF_targeting_", "", .) %>%
  append("TF", .)
colnames(TF_targeting_all_het) <- names

# grabbing all het targeting scores for genes
gene_targeting_all_het <- targeting_list[grep("Gene_", names(targeting_list))] %>%
  purrr::reduce(full_join, by = "gene") %>%
  select(., contains(c("gene", "het_edge_weight_pos")))

# grabbing colnames and setting for new dataframe generated above
names <- names(targeting_list[grep("Gene_", names(targeting_list))]) %>%
  sub(".*Gene_targeting_", "", .) %>%
  append("gene", .)
colnames(gene_targeting_all_het) <- names
```


```{r constructing-ctrl-targeting-matrix}
# grabbing all ctrl targeting scores for TFs
TF_targeting_all_ctrl <- targeting_list[grep("TF_", names(targeting_list))] %>%
  purrr::reduce(full_join, by = "TF") %>%
  select(., contains(c("TF", "ctrl_edge_weight_pos")))

# grabbing colnames and setting for new dataframe generated above
names <- names(targeting_list[grep("TF_", names(targeting_list))]) %>%
  sub(".*TF_targeting_", "", .) %>%
  append("TF", .)
colnames(TF_targeting_all_ctrl) <- names

# grabbing all ctrl targeting scores for genes
gene_targeting_all_ctrl <- targeting_list[grep("Gene_", names(targeting_list))] %>%
  purrr::reduce(full_join, by = "gene") %>%
  select(., contains(c("gene", "ctrl_edge_weight_pos")))

# grabbing colnames and setting for new dataframe generated above
names <- names(targeting_list[grep("Gene_", names(targeting_list))]) %>%
  sub(".*Gene_targeting_", "", .) %>%
  append("gene", .)
colnames(gene_targeting_all_ctrl) <- names
```


# Investigate targeting differences across Het cell types:
```{r find-top-targeted-genes-across-het-celltypes}
# identify the maximum for each row
max <- do.call(pmax, gene_targeting_all_het[, -1]) # finding max for each gene across cell types

# subtract each value from the maximum by gene; the gene with the most positive rowsum is the most variably targeted across cell types, and the cell type with a value = 0 represents the cell type with the highest targeting for that gene
gene_targeting_ranked <- max - gene_targeting_all_het[, -1]
gene_targeting_ranked$rank_sum <- rowSums(gene_targeting_ranked)

# adding gene names back
gene_targeting_ranked$gene <- gene_targeting_all_het$gene

# sorting and annotating
gene_targeting_ranked <- gene_targeting_ranked %>% arrange(desc(rank_sum))
gene_targeting_ranked$rank <- seq_len(nrow(gene_targeting_ranked))

# adding rank to targeting score dataframe
gene_targeting_all_het <- merge(select(gene_targeting_ranked, c("rank", "gene")), gene_targeting_all_het, by = "gene")
gene_targeting_all_het <- arrange(gene_targeting_all_het, rank)
```

```{r visualize-het-gene-targeting-distribution}
pivoted <- pivot_longer(gene_targeting_all_het, values_to = "targeting_score", names_to = "cell_type", cols = -c("gene", "rank"))
pivoted


png(
  filename = here("results/diff_targeting/cortex_gene_targeting_dist_het.png"),
  width = 1000,
  height = 500
)
ggplot(data = pivoted, aes(x = targeting_score, group = cell_type, color = cell_type)) +
  geom_density(adjust = 1.5, alpha = .4) +
  ggtitle("cortex Gene Targeting Score Distribution in S858R Mice") +
  theme_bw() +
  scale_color_manual(values = c("#6968B4", "#DC71FA", "#0a2f6f", "#00BBDB", "#C216B7", "#027461", "#7997FF", "#00C1AA"))
dev.off()
```

## plotting targeting heatmaps
```{r visualizing-het-gene-rank-with-heatmap}
# formatting input data
data <- gene_targeting_all_het
rownames(data) <- data$gene
data <- data[, -c(1:2)]
# setting meta colname
meta_colname <- "cell_type"
# setting plot path
plot_path <- here("results/diff_targeting/cortex_gene_targeting_all_het_heatmap_test.png")
# setting rowtitle
rowtitle <- "genes"
# setting columntitle
plot_title <- "Targeting of All Genes Across Cell-types in Setbp1 S858R Cerebral Cortex"

targeting_heatmap(annotation_colors, data, meta_colname, plot_path, rowtitle, plot_title)
```

```{r visualizing-het-gene-top100-rank-with-heatmap}
# formatting input data
data <- gene_targeting_all_het %>% slice(1:100)
rownames(data) <- data$gene
data <- data[, -c(1:3)]
# setting meta colname
meta_colname <- "cell_type"
# setting plot path
plot_path <- here("results/diff_targeting/cortex_targeting_het_top100_heatmap.png")
# setting rowtitle
rowtitle <- "genes"
# setting columntitle
plot_title <- "Targeting of top 100 Genes Across Cell-types in Setbp1 S858R Cerebral Cortex"

targeting_heatmap(annotation_colors, data, meta_colname, plot_path, rowtitle, plot_title, show_names = TRUE)
```

```{r visualizing-setbp1-targets-gene-targeting-across-het}
# formatting input data
data <- gene_targeting_all_het[gene_targeting_all_het$gene %in% setbp1_genes, ]

rownames(data) <- data$gene
data <- data[, -c(1:3)]
# setting meta colname
meta_colname <- "cell_type"
# setting plot path
plot_path <- here("results/diff_targeting/cortex_gene_targeting_het_setbp1targets_across_heatmap.png")
# setting rowtitle
rowtitle <- "genes"
# setting columntitle
plot_title <- "Gene targeting of Setbp1 and it's targets Across Cell-types in Setbp1 S858R Cerebral Cortex"

targeting_heatmap(annotation_colors, data, meta_colname, plot_path, rowtitle, plot_title, show_names = TRUE, width = 1000, height = 2000)
```


```{r find-top-targeted-TFs-across-het-celltypes}
# identify the maximum for each row
max <- do.call(pmax, TF_targeting_all_het[, -1]) # finding max for each TF across cell types

# subtract each value from the maximum by TF; the TF with the most positive rowsum is the most variably targeted across cell types, and the cell type with a value = 0 represents the cell type with the highest targeting for that TF
TF_targeting_ranked <- max - TF_targeting_all_het[, -1]
TF_targeting_ranked$rank_sum <- rowSums(TF_targeting_ranked)

# adding TF names back
TF_targeting_ranked$TF <- TF_targeting_all_het$TF

# sorting and annotating
TF_targeting_ranked <- TF_targeting_ranked %>% arrange(desc(rank_sum))
TF_targeting_ranked$rank <- seq_len(nrow(TF_targeting_ranked))

# adding rank to targeting score dataframe
TF_targeting_all_het <- merge(select(TF_targeting_ranked, c("rank", "TF")), TF_targeting_all_het, by = "TF")
TF_targeting_all_het <- arrange(TF_targeting_all_het, rank)
```

```{r visualize-het-TF-targeting-distribution}
pivoted <- pivot_longer(TF_targeting_all_het, values_to = "targeting_score", names_to = "cell_type", cols = -c("TF", "rank"))

png(
  filename = here("results/diff_targeting/cortex_TF_targeting_dist_het.png"),
  width = 1000,
  height = 500
)
ggplot(data = pivoted, aes(x = targeting_score, group = cell_type, color = cell_type)) +
  geom_density(adjust = 1.5, alpha = .4) +
  ggtitle("TF Targeting Score Distribution in S858R Mice") +
  theme_bw() +
  scale_color_manual(values = c("#6968B4", "#DC71FA", "#0a2f6f", "#00BBDB", "#C216B7", "#027461", "#7997FF", "#00C1AA"))
dev.off()
```

```{r visualizing-het-TF-rank-with-heatmap}
# formatting input data
data <- TF_targeting_all_het
rownames(data) <- data$TF
data <- data[, -c(1:2)]
# setting meta colname
meta_colname <- "cell_type"
# setting plot path
plot_path <- here("results/diff_targeting/cortex_TF_targeting_all_het_heatmap.png")
# setting rowtitle
rowtitle <- "TFs"
# setting columntitle
plot_title <- "Targeting of All TFs Across Cell-types in Setbp1 S858R Cerebral Cortex"

targeting_heatmap(annotation_colors, data, meta_colname, plot_path, rowtitle, plot_title)
```

```{r visualizing-het-top100-TF-rank-with-heatmap}
# formatting input data
data <- TF_targeting_all_het %>% slice(1:100)
rownames(data) <- data$TF
data <- data[, -c(1:2)]
# setting meta colname
meta_colname <- "cell_type"
# setting plot path
plot_path <- here("results/diff_targeting/cortex_TF_targeting_het_top100_heatmap.png")
# setting rowtitle
rowtitle <- "TFs"
# setting columntitle
plot_title <- "Targeting of TF Across Cell-types in Setbp1 S858R Cerebral Cortex"

targeting_heatmap(annotation_colors, data, meta_colname, plot_path, rowtitle, plot_title, show_names = TRUE)
```

```{r visualizing-setbp1-targets-TF-targeting-across-het}
# formatting input data
data <- TF_targeting_all_het[TF_targeting_all_het$TF %in% setbp1_genes, ]

rownames(data) <- data$TF
data <- data[, -c(1:2)]
# setting meta colname
meta_colname <- "cell_type"
# setting plot path
plot_path <- here("results/diff_targeting/cortex_TF_targeting_het_setbp1targets_across_heatmap.png")
# setting rowtitle
rowtitle <- "TFs"
# setting columntitle
plot_title <- "TF targeting of Setbp1 and it's targets Across Cell-types in Setbp1 S858R Cerebral Cortex"

targeting_heatmap(annotation_colors, data, meta_colname, plot_path, rowtitle, plot_title, show_names = TRUE)
```

# Investigate targeting differences across ctrl cell types:

```{r find-top-targeted-genes-across-ctrl-celltypes}
# identify the maximum for each row
max <- do.call(pmax, gene_targeting_all_ctrl[, -1]) # finding max for each gene across cell types

# subtract each value from the maximum by gene; the gene with the most positive rowsum is the most variably targeted across cell types, and the cell type with a value = 0 represents the cell type with the highest targeting for that gene
gene_targeting_ranked <- max - gene_targeting_all_ctrl[, -1]
gene_targeting_ranked$rank_sum <- rowSums(gene_targeting_ranked)

# adding gene names back
gene_targeting_ranked$gene <- gene_targeting_all_ctrl$gene

# sorting and annotating
gene_targeting_ranked <- gene_targeting_ranked %>% arrange(desc(rank_sum))
gene_targeting_ranked$rank <- seq_len(nrow(gene_targeting_ranked))

# adding rank to targeting score dataframe
gene_targeting_all_ctrl <- merge(select(gene_targeting_ranked, c("rank", "gene")), gene_targeting_all_ctrl, by = "gene")
gene_targeting_all_ctrl <- arrange(gene_targeting_all_ctrl, rank)
```

```{r visualize-ctrl-gene-targeting-distribution}
pivoted <- pivot_longer(gene_targeting_all_ctrl, values_to = "targeting_score", names_to = "cell_type", cols = -c("gene", "rank"))
pivoted

png(
  filename = here("results/diff_targeting/cortex_gene_targeting_dist_ctrl.png"),
  width = 1000,
  height = 500
)
ggplot(data = pivoted, aes(x = targeting_score, group = cell_type, color = cell_type)) +
  geom_density(adjust = 1.5, alpha = .4) +
  ggtitle("cortex Gene Targeting Score Distribution in Ctrl Mice") +
  theme_bw() +
  scale_color_manual(values = c("#6968B4", "#DC71FA", "#0a2f6f", "#00BBDB", "#C216B7", "#027461", "#7997FF", "#00C1AA"))
dev.off()

summary(gene_targeting_all_ctrl)
```

```{r visualizing-ctrl-gene-rank-with-heatmap}
# formatting input data
data <- gene_targeting_all_ctrl
rownames(data) <- data$gene
data <- data[, -c(1:2)]
# setting meta colname
meta_colname <- "cell_type"
# setting plot path
plot_path <- here("results/diff_targeting/cortex_gene_targeting_all_ctrl_heatmap.png")
# setting rowtitle
rowtitle <- "genes"
# setting columntitle
plot_title <- "Targeting of All Genes Across Cell-types in Setbp1 Ctrl Cerebral Cortex"

targeting_heatmap(annotation_colors, data, meta_colname, plot_path, rowtitle, plot_title)
```

```{r visualizing-ctrl-gene-top100-rank-with-heatmap}
# formatting input data
data <- gene_targeting_all_ctrl %>% slice(1:100)
rownames(data) <- data$gene
data <- data[, -c(1:2)]
# setting meta colname
meta_colname <- "cell_type"
# setting plot path
plot_path <- here("results/diff_targeting/cortex_targeting_ctrl_top100_heatmap.png")
# setting rowtitle
rowtitle <- "genes"
# setting columntitle
plot_title <- "Targeting of top 100 Genes Across Cell-types in Setbp1 Ctrl Cerebral Cortex"

targeting_heatmap(annotation_colors, data, meta_colname, plot_path, rowtitle, plot_title, show_names = TRUE)
```

```{r visualizing-setbp1-targets-genes-targeting-across-ctrl}
data <- gene_targeting_all_ctrl[gene_targeting_all_ctrl$gene %in% setbp1_genes, ]

rownames(data) <- data$gene
data <- data[, -c(1:2)]
# setting meta colname
meta_colname <- "cell_type"
# setting plot path
plot_path <- here("results/diff_targeting/cortex_gene_targeting_ctrl_setbp1targets_across_heatmap.png")
# setting rowtitle
rowtitle <- "genes"
# setting columntitle
plot_title <- "Gene targeting of Setbp1 and its targets Across Cell-types in Setbp1 Control Kidey Cortex"

targeting_heatmap(annotation_colors, data, meta_colname, plot_path, rowtitle, plot_title, show_names = TRUE)
```

```{r find-top-targeted-TFs-across-ctrl-celltypes}
# identify the maximum for each row
max <- do.call(pmax, TF_targeting_all_ctrl[, -1]) # finding max for each TF across cell types

# subtract each value from the maximum by TF; the TF with the most positive rowsum is the most variably targeted across cell types, and the cell type with a value = 0 represents the cell type with the highest targeting for that TF
TF_targeting_ranked <- max - TF_targeting_all_ctrl[, -1]
TF_targeting_ranked$rank_sum <- rowSums(TF_targeting_ranked)

# adding TF names back
TF_targeting_ranked$TF <- TF_targeting_all_ctrl$TF

# sorting and annotating
TF_targeting_ranked <- TF_targeting_ranked %>% arrange(desc(rank_sum))
TF_targeting_ranked$rank <- seq_len(nrow(TF_targeting_ranked))

# adding rank to targeting score dataframe
TF_targeting_all_ctrl <- merge(select(TF_targeting_ranked, c("rank", "TF")), TF_targeting_all_ctrl, by = "TF")
TF_targeting_all_ctrl <- arrange(TF_targeting_all_ctrl, rank)
```

```{r visualize-ctrl-TF-targeting-distribution}
pivoted <- pivot_longer(TF_targeting_all_ctrl, values_to = "targeting_score", names_to = "cell_type", cols = -c("TF", "rank"))
pivoted


png(
  filename = here("results/diff_targeting/cortex_TF_targeting_dist_ctrl.png"),
  width = 1000,
  height = 500
)
ggplot(data = pivoted, aes(x = targeting_score, group = cell_type, color = cell_type)) +
  geom_density(adjust = 1.5, alpha = .4) +
  ggtitle("cortex TF Targeting Score Distribution in Ctrl Mice") +
  theme_bw() +
  scale_color_manual(values = c("#6968B4", "#DC71FA", "#0a2f6f", "#00BBDB", "#C216B7", "#027461", "#7997FF", "#00C1AA"))
dev.off()

summary(TF_targeting_all_ctrl)
```

```{r visualizing-ctrl-TF-rank-with-heatmap}
## formatting input data
data <- TF_targeting_all_ctrl
rownames(data) <- data$TF
data <- data[, -c(1:2)]
# setting meta colname
meta_colname <- "cell_type"
# setting plot path
plot_path <- here("results/diff_targeting/cortex_TF_targeting_all_ctrl_heatmap.png")
# setting rowtitle
rowtitle <- "TFs"
# setting columntitle
plot_title <- "Targeting of All TFs Across Cell-types in Setbp1 Control Cerebral Cortex"

targeting_heatmap(annotation_colors, data, meta_colname, plot_path, rowtitle, plot_title)
```

```{r visualizing-ctrl-top100-TF-rank-with-heatmap}
# formatting input data
data <- TF_targeting_all_ctrl %>% slice(1:100)
rownames(data) <- data$TF
data <- data[, -c(1:2)]
# setting meta colname
meta_colname <- "cell_type"
# setting plot path
plot_path <- here("results/diff_targeting/cortex_TF_targeting_ctrl_top100_heatmap.png")
# setting rowtitle
rowtitle <- "TFs"
# setting columntitle
plot_title <- "Targeting of top 100 TF Across Cell-types in Setbp1 Control Cerebral Cortex"

targeting_heatmap(annotation_colors, data, meta_colname, plot_path, rowtitle, plot_title, show_names = TRUE)
```

```{r visualizing-setbp1-targets-TF-targeting-across-ctrl}
data <- TF_targeting_all_ctrl[TF_targeting_all_ctrl$TF %in% setbp1_genes, ]

rownames(data) <- data$TF
data <- data[, -c(1:2)]
# setting meta colname
meta_colname <- "cell_type"
# setting plot path
plot_path <- here("results/diff_targeting/cortex_TF_targeting_ctrl_setbp1targets_across_heatmap.png")
# setting rowtitle
rowtitle <- "TFs"
# setting columntitle
plot_title <- "TF targeting of Setbp1 and its targets Across Cell-types in Setbp1 Control Cerebral Cortex"

targeting_heatmap(annotation_colors, data, meta_colname, plot_path, rowtitle, plot_title, show_names = TRUE)
```

# Investigate targeting rank within each Het cell type

```{r find-top-targeted-TFs-within-het-celltypes}
TF_targeting_all_het <- subset(TF_targeting_all_het, select = -c(rank)) # removing rank from previous portion of analysis

cell_names <- colnames(TF_targeting_all_het) %>% .[which(. != "TF")] # note: need to change for gene, grabbing cell type names
cell_names

# rank TFs by targeting for each cell type , find top 20
TF_rank_het <- list() # creating empty list to store within

for (i in cell_names) {
  rank_het <- TF_targeting_all_het[, names(TF_targeting_all_het) %in% c("TF", i)] %>% # note: need to change TF to gene where applicable
    .[order(.[i], decreasing = TRUE), ] %>%
    mutate(rank = seq_len(nrow(.))) %>%
    mutate(cell_type = i) %>%
    slice(1:20) %>%
    rename(targeting_score = i)
  TF_rank_het[[i]] <- rank_het
}

het_TF_targeting_within <- do.call(bind_rows, TF_rank_het) # collapsing list into single df using row binding

# rank TFs by targeting for each cell type for all data
TF_rank_het <- list() # creating empty list to store within

for (i in cell_names) {
  rank_het <- TF_targeting_all_het[, names(TF_targeting_all_het) %in% c("TF", i)] %>% # note: need to change TF to gene where applicable
    .[order(.[i], decreasing = TRUE), ] %>%
    mutate(rank = seq_len(nrow(.))) %>%
    mutate(cell_type = i) %>%
    rename(targeting_score = i)
  TF_rank_het[[i]] <- rank_het
}

het_all_TF_targeting_within <- do.call(bind_rows, TF_rank_het) # collapsing list into single df using row binding
```

```{r visualizing-het-TF-rank-with-heatmap}
## formatting input data
data <- het_TF_targeting_within[, c("TF", "targeting_score", "cell_type"), drop = FALSE]
data <- pivot_wider(data, names_from = c("cell_type"), values_from = "targeting_score", names_repair = "minimal")
data[is.na(data)] <- 0
names <- data$TF # add gene names as rownames
data <- data[, -1, drop = FALSE] # drop gene column
rownames(data) <- names
# setting meta colname
meta_colname <- "cell_type"
# setting plot path
plot_path <- here("results/diff_targeting/cortex_TF_targeting_top20_het_within_celltype_heatmap.png")
# setting rowtitle
rowtitle <- "TFs"
# setting columntitle
plot_title <- "Targeting of Top 20 TFs within Cell-types in Setbp1 S858R Cerebral Cortex"

targeting_heatmap(annotation_colors, data, meta_colname, plot_path, rowtitle, plot_title, show_names = TRUE)
```

```{r visualizing-ALL-het-TF-rank-heatmap}
## formatting input data
data <- het_all_TF_targeting_within[, c("TF", "targeting_score", "cell_type"), drop = FALSE]
data <- pivot_wider(data, names_from = c("cell_type"), values_from = "targeting_score", names_repair = "minimal")
data[is.na(data)] <- 0
names <- data$TF # add gene names as rownames
data <- data[, -1, drop = FALSE] # drop gene column
rownames(data) <- names
# setting meta colname
meta_colname <- "cell_type"
# setting plot path
plot_path <- here("results/diff_targeting/cortex_TF_targeting_het_all_withincelltype_heatmap.png")
# setting rowtitle
rowtitle <- "TFs"
# setting columntitle
plot_title <- "Targeting of all TFs within Cell-types in Setbp1 S858R Cerebral Cortex"

targeting_heatmap(annotation_colors, data, meta_colname, plot_path, rowtitle, plot_title)
```

```{r het-setbp1-TF-targets-within}
data <- het_all_TF_targeting_within[het_all_TF_targeting_within$TF %in% setbp1_genes, ]

## formatting input data
data <- data[, c("TF", "targeting_score", "cell_type"), drop = FALSE]
data <- pivot_wider(data, names_from = c("cell_type"), values_from = "targeting_score", names_repair = "minimal")
data[is.na(data)] <- 0
names <- data$TF # add gene names as rownames
data <- data[, -1, drop = FALSE] # drop gene column
rownames(data) <- names

# setting meta colname
meta_colname <- "cell_type"
# setting plot path
plot_path <- here("results/diff_targeting/cortex_TF_targeting_het_Setbp1targets_withincelltype_heatmap.png")
# setting rowtitle
rowtitle <- "TFs"
# setting columntitle
plot_title <- "Targeting of Setbp1 and known target TFs within Cell-types in Setbp1 S858R Cerebral Cortex"

targeting_heatmap(annotation_colors, data, meta_colname, plot_path, rowtitle, plot_title, show_names = TRUE)
```


```{r find-top-targeted-genes-within-het-celltypes}
gene_targeting_all_het <- subset(gene_targeting_all_het, select = -c(rank)) # removing rank from previous portion of analysis

cell_names <- colnames(gene_targeting_all_het) %>% .[which(. != "gene")] # note: need to change for gene, grabbing cell type names
cell_names

# rank genes by targeting for each cell type , find top 20
gene_rank_het <- list() # creating empty list to store within

for (i in cell_names) {
  rank_het <- gene_targeting_all_het[, names(gene_targeting_all_het) %in% c("gene", i)] %>% # note: need to change TF to gene where applicable
    .[order(.[i], decreasing = TRUE), ] %>%
    mutate(rank = seq_len(nrow(.))) %>%
    mutate(cell_type = i) %>%
    slice(1:20) %>%
    rename(targeting_score = i)
  gene_rank_het[[i]] <- rank_het
}

het_gene_targeting_within <- do.call(bind_rows, gene_rank_het) # collapsing list into single df using row binding

# rank genes by targeting for each cell type for all data
gene_rank_het <- list() # creating empty list to store within

for (i in cell_names) {
  rank_het <- gene_targeting_all_het[, names(gene_targeting_all_het) %in% c("gene", i)] %>% # note: need to change TF to gene where applicable
    .[order(.[i], decreasing = TRUE), ] %>%
    mutate(rank = seq_len(nrow(.))) %>%
    mutate(cell_type = i) %>%
    rename(targeting_score = i)
  gene_rank_het[[i]] <- rank_het
}

het_all_gene_targeting_within <- do.call(bind_rows, gene_rank_het) # collapsing list into single df using row binding
```


```{r visualizing-het-gene-rank-with-heatmap}
## formatting input data
data <- het_gene_targeting_within[, c("gene", "targeting_score", "cell_type"), drop = FALSE]
data <- pivot_wider(data, names_from = c("cell_type"), values_from = "targeting_score", names_repair = "minimal")
data[is.na(data)] <- 0
names <- data$gene # add gene names as rownames
data <- data[, -1, drop = FALSE] # drop gene column
rownames(data) <- names
# setting meta colname
meta_colname <- "cell_type"
# setting plot path
plot_path <- here("results/diff_targeting/cortex_gene_targeting_top20_het_within_celltype_heatmap.png")
# setting rowtitle
rowtitle <- "genes"
# setting columntitle
plot_title <- "Targeting of Top 20 genes within Cell-types in Setbp1 S858R Cerebral Cortex"

targeting_heatmap(annotation_colors, data, meta_colname, plot_path, rowtitle, plot_title, show_names = TRUE)
```

```{r visualizing-ALL-het-gene-rank-heatmap}
## formatting input data
data <- het_all_gene_targeting_within[, c("gene", "targeting_score", "cell_type"), drop = FALSE]
data <- pivot_wider(data, names_from = c("cell_type"), values_from = "targeting_score", names_repair = "minimal")
data[is.na(data)] <- 0
names <- data$gene # add gene names as rownames
data <- data[, -1, drop = FALSE] # drop gene column
rownames(data) <- names
# setting meta colname
meta_colname <- "cell_type"
# setting plot path
plot_path <- here("results/diff_targeting/cortex_gene_targeting_het_all_withincelltype_heatmap.png")
# setting rowtitle
rowtitle <- "genes"
# setting columntitle
plot_title <- "Targeting of all genes within Cell-types in Setbp1 S858R Cerebral Cortex"

targeting_heatmap(annotation_colors, data, meta_colname, plot_path, rowtitle, plot_title)
```

```{r het-setbp1-gene-targets-within}
data <- het_all_gene_targeting_within[het_all_gene_targeting_within$gene %in% setbp1_genes, ]

## formatting input data
data <- data[, c("gene", "targeting_score", "cell_type"), drop = FALSE]
data <- pivot_wider(data, names_from = c("cell_type"), values_from = "targeting_score", names_repair = "minimal")
data[is.na(data)] <- 0
names <- data$gene # add gene names as rownames
data <- data[, -1, drop = FALSE] # drop gene column
rownames(data) <- names

# setting meta colname
meta_colname <- "cell_type"
# setting plot path
plot_path <- here("results/diff_targeting/cortex_gene_targeting_het_Setbp1targets_withincelltype_heatmap.png")
# setting rowtitle
rowtitle <- "genes"
# setting columntitle
plot_title <- "Targeting of Setbp1 and known target genes within Cell-types in Setbp1 S858R Cerebral Cortex"

targeting_heatmap(annotation_colors, data, meta_colname, plot_path, rowtitle, plot_title, show_names = TRUE)
```


# Investigate targeting rank within each Ctrl cell type
```{r find-top-targeted-genes-within-ctrl-celltypes}
gene_targeting_all_ctrl <- subset(gene_targeting_all_ctrl, select = -c(rank)) # removing rank from previous portion of analysis

cell_names <- colnames(gene_targeting_all_ctrl) %>% .[which(. != "gene")] # note: need to change for gene, grabbing cell type names
cell_names

# rank genes by targeting for each cell type , find top 20
gene_rank_ctrl <- list() # creating empty list to store within

for (i in cell_names) {
  rank_ctrl <- gene_targeting_all_ctrl[, names(gene_targeting_all_ctrl) %in% c("gene", i)] %>% # note: need to change TF to gene where applicable
    .[order(.[i], decreasing = TRUE), ] %>%
    mutate(rank = seq_len(nrow(.))) %>%
    mutate(cell_type = i) %>%
    slice(1:20) %>%
    rename(targeting_score = i)
  gene_rank_ctrl[[i]] <- rank_ctrl
}

ctrl_gene_targeting_within <- do.call(bind_rows, gene_rank_ctrl) # collapsing list into single df using row binding

# rank genes by targeting for each cell type for all data
gene_rank_ctrl <- list() # creating empty list to store within

for (i in cell_names) {
  rank_ctrl <- gene_targeting_all_ctrl[, names(gene_targeting_all_ctrl) %in% c("gene", i)] %>% # note: need to change TF to gene where applicable
    .[order(.[i], decreasing = TRUE), ] %>%
    mutate(rank = seq_len(nrow(.))) %>%
    mutate(cell_type = i) %>%
    rename(targeting_score = i)
  gene_rank_ctrl[[i]] <- rank_ctrl
}

ctrl_all_gene_targeting_within <- do.call(bind_rows, gene_rank_ctrl) # collapsing list into single df using row binding
```

```{r find-top-targeted-TFs-within-ctrl-celltypes}
TF_targeting_all_ctrl <- subset(TF_targeting_all_ctrl, select = -c(rank)) # removing rank from previous portion of analysis

cell_names <- colnames(TF_targeting_all_ctrl) %>% .[which(. != "TF")] # note: need to change for TF, grabbing cell type names
cell_names

# rank TFs by targeting for each cell type , find top 20
TF_rank_ctrl <- list() # creating empty list to store within

for (i in cell_names) {
  rank_ctrl <- TF_targeting_all_ctrl[, names(TF_targeting_all_ctrl) %in% c("TF", i)] %>% # note: need to change TF to gene where applicable
    .[order(.[i], decreasing = TRUE), ] %>%
    mutate(rank = seq_len(nrow(.))) %>%
    mutate(cell_type = i) %>%
    slice(1:20) %>%
    rename(targeting_score = i)
  TF_rank_ctrl[[i]] <- rank_ctrl
}

ctrl_TF_targeting_within <- do.call(bind_rows, TF_rank_ctrl) # collapsing list into single df using row binding

# rank TFs by targeting for each cell type for all data
TF_rank_ctrl <- list() # creating empty list to store within

for (i in cell_names) {
  rank_ctrl <- TF_targeting_all_ctrl[, names(TF_targeting_all_ctrl) %in% c("TF", i)] %>% # note: need to change TF to TF where applicable
    .[order(.[i], decreasing = TRUE), ] %>%
    mutate(rank = seq_len(nrow(.))) %>%
    mutate(cell_type = i) %>%
    rename(targeting_score = i)
  TF_rank_ctrl[[i]] <- rank_ctrl
}

ctrl_all_TF_targeting_within <- do.call(bind_rows, TF_rank_ctrl) # collapsing list into single df using row binding
```

```{r visualizing-ctrl-gene-rank-with-heatmap}
## formatting input data
data <- ctrl_gene_targeting_within[, c("gene", "targeting_score", "cell_type"), drop = FALSE]
data <- pivot_wider(data, names_from = c("cell_type"), values_from = "targeting_score", names_repair = "minimal")
data[is.na(data)] <- 0
names <- data$gene # add gene names as rownames
data <- data[, -1, drop = FALSE] # drop gene column
rownames(data) <- names
# setting meta colname
meta_colname <- "cell_type"
# setting plot path
plot_path <- here("results/diff_targeting/cortex_gene_targeting_top20_ctrl_within_celltype_heatmap.png")
# setting rowtitle
rowtitle <- "genes"
# setting columntitle
plot_title <- "Targeting of Top 20 genes within Cell-types in Setbp1 Ctrl Cerebral Cortex"

targeting_heatmap(annotation_colors, data, meta_colname, plot_path, rowtitle, plot_title, show_names = TRUE)
```

```{r visualizing-ALL-ctrl-gene-rank-heatmap}
## formatting input data
data <- ctrl_all_gene_targeting_within[, c("gene", "targeting_score", "cell_type"), drop = FALSE]
data <- pivot_wider(data, names_from = c("cell_type"), values_from = "targeting_score", names_repair = "minimal")
data[is.na(data)] <- 0
names <- data$gene # add gene names as rownames
data <- data[, -1, drop = FALSE] # drop gene column
rownames(data) <- names
# setting meta colname
meta_colname <- "cell_type"
# setting plot path
plot_path <- here("results/diff_targeting/cortex_gene_targeting_ctrl_all_withincelltype_heatmap.png")
# setting rowtitle
rowtitle <- "genes"
# setting columntitle
plot_title <- "Targeting of all genes within Cell-types in Setbp1 Ctrl Cerebral Cortex"

targeting_heatmap(annotation_colors, data, meta_colname, plot_path, rowtitle, plot_title)
```


```{r ctrl-setbp1-gene-targets-within}
# filtering based off of gene set
data <- ctrl_all_gene_targeting_within[ctrl_all_gene_targeting_within$gene %in% setbp1_genes, ]

## formatting input data
data <- data[, c("gene", "targeting_score", "cell_type"), drop = FALSE]
data <- pivot_wider(data, names_from = c("cell_type"), values_from = "targeting_score", names_repair = "minimal")
data[is.na(data)] <- 0
names <- data$gene # add gene names as rownames
data <- data[, -1, drop = FALSE] # drop gene column
rownames(data) <- names
# setting meta colname
meta_colname <- "cell_type"
# setting plot path
plot_path <- here("results/diff_targeting/cortex_gene_targeting_ctrl_Setbp1targets_withincelltype_heatmap.png")
# setting rowtitle
rowtitle <- "genes"
# setting columntitle
plot_title <- "Targeting of Setbp1 and known target genes within Cell-types in Setbp1 Control Cerebral Cortex"

targeting_heatmap(annotation_colors, data, meta_colname, plot_path, rowtitle, plot_title, show_names = TRUE)
```

```{r visualizing-ctrl-TF-rank-with-heatmap}
## formatting input data
data <- ctrl_TF_targeting_within[, c("TF", "targeting_score", "cell_type"), drop = FALSE]
data <- pivot_wider(data, names_from = c("cell_type"), values_from = "targeting_score", names_repair = "minimal")
data[is.na(data)] <- 0
names <- data$TF # add TF names as rownames
data <- data[, -1, drop = FALSE] # drop TF column
rownames(data) <- names
# setting meta colname
meta_colname <- "cell_type"
# setting plot path
plot_path <- here("results/diff_targeting/cortex_TF_targeting_top20_ctrl_within_celltype_heatmap.png")
# setting rowtitle
rowtitle <- "TFs"
# setting columntitle
plot_title <- "Targeting of Top 20 TFs within Cell-types in Setbp1 Ctrl Cerebral Cortex"

targeting_heatmap(annotation_colors, data, meta_colname, plot_path, rowtitle, plot_title, show_names = TRUE)
```

```{r visualizing-ALL-ctrl-TF-rank-heatmap}
## formatting input data
data <- ctrl_all_TF_targeting_within[, c("TF", "targeting_score", "cell_type"), drop = FALSE]
data <- pivot_wider(data, names_from = c("cell_type"), values_from = "targeting_score", names_repair = "minimal")
data[is.na(data)] <- 0
names <- data$TF # add TF names as rownames
data <- data[, -1, drop = FALSE] # drop TF column
rownames(data) <- names
# setting meta colname
meta_colname <- "cell_type"
# setting plot path
plot_path <- here("results/diff_targeting/cortex_TF_targeting_ctrl_all_withincelltype_heatmap.png")
# setting rowtitle
rowtitle <- "TFs"
# setting columntitle
plot_title <- "Targeting of all TFs within Cell-types in Setbp1 Ctrl Cerebral Cortex"

targeting_heatmap(annotation_colors, data, meta_colname, plot_path, rowtitle, plot_title)
```

```{r ctrl-setbp1-TF-targets-within}
# subsetting for gene set
data <- ctrl_all_TF_targeting_within[ctrl_all_TF_targeting_within$TF %in% setbp1_genes, ]

## formatting input data
data <- data[, c("TF", "targeting_score", "cell_type"), drop = FALSE]
data <- pivot_wider(data, names_from = c("cell_type"), values_from = "targeting_score", names_repair = "minimal")
data[is.na(data)] <- 0
names <- data$TF # add TF names as rownames
data <- data[, -1, drop = FALSE] # drop TF column
rownames(data) <- names

# setting meta colname
meta_colname <- "cell_type"
# setting plot path
plot_path <- here("results/diff_targeting/cortex_TF_targeting_ctrl_Setbp1targets_withincelltype_heatmap.png")
# setting rowtitle
rowtitle <- "TFs"
# setting columntitle
plot_title <- "Targeting of Setbp1 and known target TFs within Cell-types in Setbp1 Control Cerebral Cortex"

targeting_heatmap(annotation_colors, data, meta_colname, plot_path, rowtitle, plot_title, show_names = TRUE)
```

# calculating differential targeting score between conditions 
GOAL: ID genes/TFs with largest difference in targeting _between conditions within each cell type_:
* Using the combined het/ctrl dataframe for each cell type above, calculate the difference (het-ctrl) for the targeting weight where the more positive difference indicates higher gene targeting in het and more negative diff indicates higher gene targeting in ctrl. 
* The closer a value is to 0, the less difference in targeting between conditions for that cell type

```{r rearrange-data-calc-diff}
# ordering data the same
gene_targeting_all_het <- gene_targeting_all_het[order(gene_targeting_all_het["gene"]), ]
gene_targeting_all_ctrl <- gene_targeting_all_ctrl[order(gene_targeting_all_ctrl["gene"]), ]

# Add suffix to column names except the gene/TF column
new_colnames <- ifelse(names(gene_targeting_all_het) != "gene", paste0(names(gene_targeting_all_het), "_het"), names(gene_targeting_all_het))
names(gene_targeting_all_het) <- new_colnames
print(gene_targeting_all_het) # Print the updated dataframe

new_colnames <- ifelse(names(gene_targeting_all_ctrl) != "gene", paste0(names(gene_targeting_all_ctrl), "_ctrl"), names(gene_targeting_all_ctrl))
names(gene_targeting_all_ctrl) <- new_colnames
print(gene_targeting_all_ctrl) # Print the updated dataframe

# binding data together
combined_gene_targeting <- merge(gene_targeting_all_het, gene_targeting_all_ctrl, by = "gene")
```

```{r calculate-het-enriched-diff-targeting}
# creating empty data frame
diff_gene_targeting <- data.frame(matrix(NA, nrow = nrow(combined_gene_targeting), ncol = 1))
colnames(diff_gene_targeting) <- "gene" # this will need to be gene or TF respectively
diff_gene_targeting$gene <- combined_gene_targeting$gene # adding gene/TF

# calc diff targeting
cell_names <- colnames(gene_targeting_all_ctrl[, -1]) %>% sub("_ctrl.*", "", .) # all cell type prefixes except gene name, which should be first column in data
for (i in cell_names) {
  selected_columns <- names(combined_gene_targeting)[startsWith(names(combined_gene_targeting), i)]
  # Check there are two columns with the specified prefix
  if (length(selected_columns) == 2) {
    diff_col <- combined_gene_targeting[[selected_columns[1]]] - combined_gene_targeting[[selected_columns[2]]] # ensure difference is `het - control` here
    # Create a new column name
    diff_colname <- paste0(i)
    # Add the result to the result dataframe
    diff_gene_targeting[[diff_colname]] <- diff_col
  }
}

het_enriched <- replace(diff_gene_targeting, diff_gene_targeting[, ] < 0, 0) # creating dataframe where the differential is enriched in the het condition
save(het_enriched, file = here("results/diff_targeting/cortex_het_enriched_gene.Rdata"))
```

```{r calculate-ctrl-enriched-diff-targeting}
# creating empty data frame
diff_gene_targeting <- data.frame(matrix(NA, nrow = nrow(combined_gene_targeting), ncol = 1))
colnames(diff_gene_targeting) <- "gene" # this will need to be gene or TF respectively
diff_gene_targeting$gene <- combined_gene_targeting$gene # adding gene/TF

# calc diff targeting
cell_names <- colnames(gene_targeting_all_ctrl[, -1]) %>% sub("_ctrl.*", "", .) # all cell type prefixes except gene name, which should be first column in data
for (i in cell_names) {
  selected_columns <- names(combined_gene_targeting)[startsWith(names(combined_gene_targeting), i)]
  # Check there are two columns with the specified prefix
  if (length(selected_columns) == 2) {
    diff_col <- combined_gene_targeting[[selected_columns[2]]] - combined_gene_targeting[[selected_columns[1]]] # ensure difference is `control - het` here
    # Create a new column name
    diff_colname <- paste0(i)
    # Add the result to the result dataframe
    diff_gene_targeting[[diff_colname]] <- diff_col
  }
}

ctrl_enriched <- replace(diff_gene_targeting, diff_gene_targeting[, ] < 0, 0) # creating dataframe where the differential is enriched in the ctrl condition
save(ctrl_enriched, file = here("results/diff_targeting/cortex_ctrl_enriched_gene.Rdata"))
```

```{r visualizing-all-diff-gene-het-enriched}
## formatting input data
data <- het_enriched
names <- data$gene # add gene names as rownames
data <- data[, -1, drop = FALSE] # drop TF column
rownames(data) <- names
# setting meta colname
meta_colname <- "cell_type"
# setting plot path
plot_path <- here("results/diff_targeting/cortex_gene_DIFFtargeting_het_enriched_heatmap.png")
# setting rowtitle
rowtitle <- "genes"
# setting columntitle
plot_title <- "Differential Targeting of S858R enriched genes in Cerebral Cortex"

targeting_heatmap(annotation_colors, data, meta_colname, plot_path, rowtitle, plot_title)
```

```{r visualizing-all-diff-gene-het-enriched-setbp1genes}
# filtering for gene set
data <- het_enriched[het_enriched$gene %in% setbp1_genes, ]

names <- data$gene # add names as rownames
data <- data[, -1, drop = FALSE] # drop TF column
rownames(data) <- names
# setting meta colname
meta_colname <- "cell_type"
# setting plot path
plot_path <- here("results/diff_targeting/cortex_gene_DIFFtargeting_het_Setbp1targets_enriched_heatmap.png")
# setting rowtitle
rowtitle <- "genes"
# setting columntitle
plot_title <- "Differential Gene Targeting of known targets of Setbp1 in S858R enriched genes of Cerebral Cortex"

targeting_heatmap(annotation_colors, data, meta_colname, plot_path, rowtitle, plot_title, show_names = TRUE, height = 2200)
```

```{r GO-difftargeting-genes-Setbp1Target-het-enriched}
## pathway analysis and plotting
exclude_column <- "gene" # note: this needs to be changed for TF or gene respectively, indicates which columns to pivot longer below,
data <- het_enriched[het_enriched$gene %in% setbp1_genes, ] %>%
  pivot_longer(., cols = which(colnames(.) != exclude_column), names_to = "cell_type", values_to = "targeting_score") %>%
  .[!.$targeting_score == 0, ]

for (i in cell_names) {
  # grab genes/TFs from cell type i ----------
  top <- data[data$cell_type == i, ]
  genes <- as.character(top$gene)
  # submit genes to pathway analysis function ----------
  fea_result_filt <- fea(genes = genes, organism = "mmusculus", max_term = 1000, min_term = 1)
  png(paste0(here("results/diff_targeting/"), "DiffTargeting_", i, "genes_ Setbp1targets_het_enriched.png"), width = 1000, height = 1000)
  # plot dotplot and add title based on input variable ----------
  plot <- bubbleplot(fea_result_filt) +
    ggtitle(paste0("Pathways Associated with Differentially Targeted genes (Setbp1 or known targets) in Setbp1 S858R ", i)) +
    theme(plot.title = element_text(hjust = 0.5, size = 12)) +
    theme_minimal()
  print(plot)
  dev.off()
  print(paste0("pathways investigated and plotted in ", i))
}
``` 
[1] "pathways investigated and plotted in astrocytes"
[1] "pathways investigated and plotted in excitatory_neurons"
[1] "pathways investigated and plotted in fibro_cortex"
[1] "pathways investigated and plotted in inhibitory_neurons"
[1] "pathways investigated and plotted in microglia"
[1] "pathways investigated and plotted in mural"
[1] "pathways investigated and plotted in oligodendrocytes"
[1] "pathways investigated and plotted in opcs"
```{r visualizing-all-diff-gene-ctrl-enriched}
## formatting input data
data <- ctrl_enriched
names <- data$gene # add gene names as rownames
data <- abs(data[, -1])
rownames(data) <- names
# setting meta colname
meta_colname <- "cell_type"
# setting plot path
plot_path <- here("results/diff_targeting/cortex_gene_DIFFtargeting_ctrl_enriched_heatmap.png")
# setting rowtitle
rowtitle <- "genes"
# setting columntitle
plot_title <- "Differential Targeting of Control enriched genes in Cerebral Cortex"

targeting_heatmap(annotation_colors, data, meta_colname, plot_path, rowtitle, plot_title)
```

```{r rearrange-data-calc-diff-TF}
# ordering data the same
TF_targeting_all_het <- TF_targeting_all_het[order(TF_targeting_all_het["TF"]), ]
TF_targeting_all_ctrl <- TF_targeting_all_ctrl[order(TF_targeting_all_ctrl["TF"]), ]

# Add suffix to column names except the gene/TF column
new_colnames <- ifelse(names(TF_targeting_all_het) != "TF", paste0(names(TF_targeting_all_het), "_het"), names(TF_targeting_all_het))
names(TF_targeting_all_het) <- new_colnames
print(TF_targeting_all_het) # Print the updated dataframe

new_colnames <- ifelse(names(TF_targeting_all_ctrl) != "TF", paste0(names(TF_targeting_all_ctrl), "_ctrl"), names(TF_targeting_all_ctrl))
names(TF_targeting_all_ctrl) <- new_colnames
print(TF_targeting_all_ctrl) # Print the updated dataframe

# binding data together
combined_TF_targeting <- merge(TF_targeting_all_het, TF_targeting_all_ctrl, by = "TF")
```

```{r calculate-het-enriched-diff-targeting}
# creating empty data frame
diff_TF_targeting <- data.frame(matrix(NA, nrow = nrow(combined_TF_targeting), ncol = 1))
colnames(diff_TF_targeting) <- "TF" # this will need to be gene or TF respectively
diff_TF_targeting$TF <- combined_TF_targeting$TF # adding gene/TF

# calc diff targeting
cell_names <- colnames(TF_targeting_all_ctrl[, -1]) %>% sub("_ctrl.*", "", .) # all cell type prefixes except TF name, which should be first column in data
for (i in cell_names) {
  selected_columns <- names(combined_TF_targeting)[startsWith(names(combined_TF_targeting), i)]
  # Check there are two columns with the specified prefix
  if (length(selected_columns) == 2) {
    diff_col <- combined_TF_targeting[[selected_columns[1]]] - combined_TF_targeting[[selected_columns[2]]] # ensure difference is `het - control` here
    # Create a new column name
    diff_colname <- paste0(i)
    # Add the result to the result dataframe
    diff_TF_targeting[[diff_colname]] <- diff_col
  }
}

het_enriched_TF <- replace(diff_TF_targeting, diff_TF_targeting[, ] < 0, 0) # creating dataframe where the differential is enriched in the het condition
save(het_enriched_TF, file = here("results/diff_targeting/cortex_het_enriched_TF.Rdata"))
```

```{r calculate-ctrl-enriched-diff-targeting}
# creating empty data frame
diff_TF_targeting <- data.frame(matrix(NA, nrow = nrow(combined_TF_targeting), ncol = 1))
colnames(diff_TF_targeting) <- "TF" # this will need to be gene or TF respectively
diff_TF_targeting$TF <- combined_TF_targeting$TF # adding gene/TF

# calc diff targeting
cell_names <- colnames(TF_targeting_all_ctrl[, -1]) %>% sub("_ctrl.*", "", .) # all cell type prefixes except TF name, which should be first column in data
for (i in cell_names) {
  selected_columns <- names(combined_TF_targeting)[startsWith(names(combined_TF_targeting), i)]
  # Check there are two columns with the specified prefix
  if (length(selected_columns) == 2) {
    diff_col <- combined_TF_targeting[[selected_columns[2]]] - combined_TF_targeting[[selected_columns[1]]] # ensure difference is `control - het` here
    # Create a new column name
    diff_colname <- paste0(i)
    # Add the result to the result dataframe
    diff_TF_targeting[[diff_colname]] <- diff_col
  }
}

ctrl_enriched_TF <- replace(diff_TF_targeting, diff_TF_targeting[, ] < 0, 0) # creating dataframe where the differential is enriched in the ctrl condition
save(ctrl_enriched_TF, file = here("results/diff_targeting/cortex_ctrl_enriched_TF.Rdata"))
```

```{r visualizing-all-diff-TF-het-enriched}
## formatting input data
data <- het_enriched_TF
names <- data$TF # add TF names as rownames
data <- data[, -1, drop = FALSE] # drop TF column
rownames(data) <- names
# setting meta colname
meta_colname <- "cell_type"
# setting plot path
plot_path <- here("results/diff_targeting/cortex_TF_DIFFtargeting_het_enriched_heatmap.png")
# setting rowtitle
rowtitle <- "TFs"
# setting columntitle
plot_title <- "Differential Targeting of S858R enriched TFs in Cerebral Cortex"

targeting_heatmap(annotation_colors, data, meta_colname, plot_path, rowtitle, plot_title)
```

```{r visualizing-all-diff-TF-het-enriched-setbp1TFs}
## formatting input data
data <- het_enriched_TF[het_enriched_TF$TF %in% setbp1_genes, ]

names <- data$TF # add names as rownames
data <- data[, -1, drop = FALSE] # drop TF column
rownames(data) <- names
# setting meta colname
meta_colname <- "cell_type"
# setting plot path
plot_path <- here("results/diff_targeting/cortex_TF_DIFFtargeting_het_Setbp1targets_enriched_heatmap.png")
# setting rowtitle
rowtitle <- "TFs"
# setting columntitle
plot_title <- "Differential TF Targeting of known targets of Setbp1 in S858R enriched TFs of Cerebral Cortex"

targeting_heatmap(annotation_colors, data, meta_colname, plot_path, rowtitle, plot_title, show_names = TRUE)
```

```{r GO-difftargeting-TFs-Setbp1Target-het-enriched}
## pathway analysis and plotting
exclude_column <- "TF" # note: this needs to be changed for TF or gene respectively, indicates which columns to pivot longer below,
data <- het_enriched_TF[het_enriched_TF$TF %in% setbp1_genes, ] %>%
  pivot_longer(., cols = which(colnames(.) != exclude_column), names_to = "cell_type", values_to = "targeting_score") %>%
  .[!.$targeting_score == 0, ]

for (i in cell_names) {
  # grab genes/TFs from cell type i ----------
  top <- data[data$cell_type == i, ]
  genes <- as.character(top$TF)
  # submit genes to pathway analysis function ----------
  fea_result_filt <- fea(genes = genes, organism = "mmusculus", max_term = 1000, min_term = 1)
  png(paste0(here("results/diff_targeting/"), "DiffTargeting_", i, "TFs_ Setbp1targets_het_enriched.png"), width = 1000, height = 1000)
  # plot dotplot and add title based on input variable ----------
  plot <- bubbleplot(fea_result_filt) +
    ggtitle(paste0("Pathways Associated with Differentially Targeted TFs (Setbp1 or known targets) in Setbp1 S858R ", i)) +
    theme(plot.title = element_text(hjust = 0.5, size = 12)) +
    theme_minimal()
  print(plot)
  dev.off()
  print(paste0("pathways investigated and plotted in ", i))
}
``` 
[1] "pathways investigated and plotted in astrocytes"
[1] "pathways investigated and plotted in excitatory_neurons"
[1] "pathways investigated and plotted in fibro_cortex"
[1] "pathways investigated and plotted in inhibitory_neurons"
[1] "pathways investigated and plotted in microglia"
[1] "pathways investigated and plotted in mural"
[1] "pathways investigated and plotted in oligodendrocytes"
[1] "pathways investigated and plotted in opcs"
```{r visualizing-all-diff-TF-ctrl-enriched}
## formatting input data
data <- ctrl_enriched_TF
names <- data$TF # add TF names as rownames
data <- abs(data[, -1])
rownames(data) <- names
# setting meta colname
meta_colname <- "cell_type"
# setting plot path
plot_path <- here("results/diff_targeting/cortex_TF_DIFFtargeting_ctrl_enriched_heatmap.png")
# setting rowtitle
rowtitle <- "TFs"
# setting columntitle
plot_title <- "Differential Targeting of Control enriched TFs in Cerebral Cortex"

targeting_heatmap(annotation_colors, data, meta_colname, plot_path, rowtitle, plot_title)
```

```{r visualize-distribution-of-het-gene-difftargeting-scores}
# het gene distribution
exclude_column <- "gene"

pivoted <- pivot_longer(het_enriched, values_to = "targeting_score", names_to = "cell_type", cols = which(colnames(het_enriched) != exclude_column))
ggplot(data = pivoted, aes(x = targeting_score, group = cell_type, color = cell_type)) +
  geom_density(adjust = 1.5, alpha = .4) +
  ggtitle("Gene targeting score distribution in S8585R") +
  scale_color_manual(values = c("#6968B4", "#DC71FA", "#0a2f6f", "#00BBDB", "#C216B7", "#027461", "#7997FF", "#00C1AA"))
summary(het_enriched)
```


```{r het-diff-gene-targeting-distribution-nonzero}
# het gene distribution nonzero
nonzero <- pivoted[pivoted$targeting_score != 0, ]
png(
  filename = here("results/diff_targeting/cortex_difftargeting_het_gene_distribution.png"),
  width = 1000,
  height = 500
)
ggplot(data = nonzero, aes(x = targeting_score, group = cell_type, color = cell_type)) +
  geom_density(adjust = 1.5, alpha = .4) +
  ggtitle("Gene differential targeting score distribution in S8585R") +
  theme_bw() +
  scale_color_manual(values = c("#6968B4", "#DC71FA", "#0a2f6f", "#00BBDB", "#C216B7", "#027461", "#7997FF", "#00C1AA"))
dev.off()

nonzero_long <- pivot_wider(nonzero, names_from = cell_type, values_from = targeting_score)
summary(nonzero_long)

exclude_column <- "TF"
pivoted <- pivot_longer(het_enriched_TF, values_to = "targeting_score", names_to = "cell_type", cols = which(colnames(het_enriched_TF) != exclude_column))
summary(het_enriched)
```

```{r control-diff-gene-targeting-distribution}
# control gene distribution
exclude_column <- "gene"
pivoted <- pivot_longer(ctrl_enriched, values_to = "targeting_score", names_to = "cell_type", cols = which(colnames(ctrl_enriched) != exclude_column))
ggplot(data = pivoted, aes(x = abs(targeting_score), group = cell_type, color = cell_type)) +
  geom_density(adjust = 1.5, alpha = .4) +
  ggtitle("Gene targeting score distribution in Control") +
  scale_color_manual(values = c("#6968B4", "#DC71FA", "#0a2f6f", "#00BBDB", "#C216B7", "#027461", "#7997FF", "#00C1AA"))

summary(ctrl_enriched)


nonzero <- pivoted[pivoted$targeting_score != 0, ]
png(
  filename = here("results/diff_targeting/cortex_difftargeting_ctrl_gene_distribution.png"),
  width = 1000,
  height = 500
)
ggplot(data = nonzero, aes(x = abs(targeting_score), group = cell_type, color = cell_type)) +
  geom_density(adjust = 1.5, alpha = .4) +
  ggtitle("Gene differential targeting score distribution in Control") +
  theme_bw() +
  scale_color_manual(values = c("#6968B4", "#DC71FA", "#0a2f6f", "#00BBDB", "#C216B7", "#027461", "#7997FF", "#00C1AA"))
dev.off()

nonzero_long <- pivot_wider(nonzero, names_from = cell_type, values_from = targeting_score)
nonzero_long <- abs(nonzero_long[, which(colnames(nonzero_long) != exclude_column)])
summary(nonzero_long)
```

```{r het-diff-TF-targting-dist}
# het TF distribution
exclude_column <- "TF"

pivoted <- pivot_longer(het_enriched_TF, values_to = "targeting_score", names_to = "cell_type", cols = which(colnames(het_enriched_TF) != exclude_column))
ggplot(data = pivoted, aes(x = targeting_score, group = cell_type, color = cell_type)) +
  geom_density(adjust = 1.5, alpha = .4) +
  ggtitle("TF targeting score distribution in S8585R") +
  scale_color_manual(values = c("#6968B4", "#DC71FA", "#0a2f6f", "#00BBDB", "#C216B7", "#027461", "#7997FF", "#00C1AA"))
summary(het_enriched)

nonzero <- pivoted[pivoted$targeting_score != 0, ]
png(
  filename = here("results/diff_targeting/cortex_difftargeting_het_TF_distribution.png"),
  width = 1000,
  height = 500
)
ggplot(data = nonzero, aes(x = targeting_score, group = cell_type, color = cell_type)) +
  geom_density(adjust = 1.5, alpha = .4) +
  ggtitle("TF differential targeting score distribution in S8585R") +
  theme_bw() +
  scale_color_manual(values = c("#6968B4", "#DC71FA", "#0a2f6f", "#00BBDB", "#C216B7", "#027461", "#7997FF", "#00C1AA"))
dev.off()


nonzero_long <- pivot_wider(nonzero, names_from = cell_type, values_from = targeting_score)
summary(nonzero_long)
```

```{r ctrl-diff-TF-targting-dist}
exclude_column <- "TF"
pivoted <- pivot_longer(ctrl_enriched_TF, values_to = "targeting_score", names_to = "cell_type", cols = which(colnames(het_enriched_TF) != exclude_column))
ggplot(data = pivoted, aes(x = abs(targeting_score), group = cell_type, color = cell_type)) +
  geom_density(adjust = 1.5, alpha = .4) +
  ggtitle("TF targeting score distribution in Control") +
  scale_color_manual(values = c("#6968B4", "#DC71FA", "#0a2f6f", "#00BBDB", "#C216B7", "#027461", "#7997FF", "#00C1AA"))

summary(ctrl_enriched)

nonzero <- pivoted[pivoted$targeting_score != 0, ]
png(
  filename = here("results/diff_targeting/cortex_difftargeting_ctrl_TF_distribution.png"),
  width = 1000,
  height = 500
)
ggplot(data = nonzero, aes(x = abs(targeting_score), group = cell_type, color = cell_type)) +
  geom_density(adjust = 1.5, alpha = .4) +
  ggtitle("TF differential targeting score distribution in Control") +
  theme_bw() +
  scale_color_manual(values = c("#6968B4", "#DC71FA", "#0a2f6f", "#00BBDB", "#C216B7", "#027461", "#7997FF", "#00C1AA"))
dev.off()

nonzero_long <- pivot_wider(nonzero, names_from = cell_type, values_from = targeting_score)
nonzero_long <- abs(nonzero_long[, c(2:8)])
summary(nonzero_long)
```

```{r find-top-diff-targeted-genes-het-celltypes}
cell_names <- colnames(het_enriched) %>% .[which(. != "gene")] # note: need to change for gene, grabbing cell type names
cell_names

# rank genes by targeting for each cell type , find top 20
gene_diff_het <- list() # creating empty list to store within

for (i in cell_names) {
  rank_het <- het_enriched[, names(het_enriched) %in% c("gene", i)] %>% # note: need to change TF to gene where applicable
    .[order(.[i], decreasing = TRUE), ] %>%
    mutate(rank = seq_len(nrow(.))) %>%
    mutate(cell_type = i) %>%
    slice(1:20) %>%
    rename(targeting_score = i)
  gene_diff_het[[i]] <- rank_het
}

cortex_het_enriched_top_genes <- do.call(bind_rows, gene_diff_het) # collapsing list into single df using row binding

# rank genes by targeting for each cell type for all data
gene_diff_het <- list() # creating empty list to store within

for (i in cell_names) {
  rank_het <- het_enriched[, names(het_enriched) %in% c("gene", i)] %>% # note: need to change TF to gene where applicable
    .[order(.[i], decreasing = TRUE), ] %>%
    mutate(rank = seq_len(nrow(.))) %>%
    mutate(cell_type = i) %>%
    rename(targeting_score = i)
  gene_diff_het[[i]] <- rank_het
}

cortex_het_enriched_all_genes <- do.call(bind_rows, gene_diff_het) # collapsing list into single df using row binding
```

```{r visualize-het-enriched-gene-top20-heatmap}
## formatting input data
data <- cortex_het_enriched_top_genes[, c("gene", "targeting_score", "cell_type"), drop = FALSE]
data <- pivot_wider(data, names_from = c("cell_type"), values_from = "targeting_score", names_repair = "minimal")
data[is.na(data)] <- 0
names <- data$gene # add gene names as rownames
data <- data[, -1, drop = FALSE] # drop gene column
rownames(data) <- names
# setting meta colname
meta_colname <- "cell_type"
# setting plot path
plot_path <- here("results/diff_targeting/cortex_gene_DIFFtargeting_top20_het_heatmap.png")
# setting rowtitle
rowtitle <- "genes"
# setting columntitle
plot_title <- "Differential Targeting of Top 20 Genes within Cell-types in Setbp1 S858R Cerebral Cortex"

targeting_heatmap(annotation_colors, data, meta_colname, plot_path, rowtitle, plot_title, show_names = TRUE, height = 2000)
```


```{r find-top-diff-targeted-het-enriched-TF-top20}
cell_names <- colnames(het_enriched_TF) %>% .[which(. != "TF")] # note: need to change for gene, grabbing cell type names
cell_names

# rank TFs by targeting for each cell type , find top 20
TF_diff_het <- list() # creating empty list to store within

for (i in cell_names) {
  rank_het <- het_enriched_TF[, names(het_enriched_TF) %in% c("TF", i)] %>% # note: need to change TF to gene where applicable
    .[order(.[i], decreasing = TRUE), ] %>%
    mutate(rank = seq_len(nrow(.))) %>%
    mutate(cell_type = i) %>%
    slice(1:20) %>%
    rename(targeting_score = i)
  TF_diff_het[[i]] <- rank_het
}

cortex_het_enriched_top_TFs <- do.call(bind_rows, TF_diff_het) # collapsing list into single df using row binding

# rank TFs by targeting for each cell type for all data
TF_diff_het <- list() # creating empty list to store within

for (i in cell_names) {
  rank_het <- het_enriched_TF[, names(het_enriched_TF) %in% c("TF", i)] %>% # note: need to change TF to TF where applicable
    .[order(.[i], decreasing = TRUE), ] %>%
    mutate(rank = seq_len(nrow(.))) %>%
    mutate(cell_type = i) %>%
    rename(targeting_score = i)
  TF_diff_het[[i]] <- rank_het
}

cortex_het_enriched_all_TFs <- do.call(bind_rows, TF_diff_het) # collapsing list into single df using row binding
```

```{r visualize-het-enriched-TF-top20-heatmap}
## formatting input data
data <- cortex_het_enriched_top_TFs[, c("TF", "targeting_score", "cell_type"), drop = FALSE]
data <- pivot_wider(data, names_from = c("cell_type"), values_from = "targeting_score", names_repair = "minimal")
data[is.na(data)] <- 0
names <- data$TF # add TF names as rownames
data <- data[, -1, drop = FALSE] # drop TF column
rownames(data) <- names
# setting meta colname
meta_colname <- "cell_type"
# setting plot path
plot_path <- here("results/diff_targeting/cortex_TF_DIFFtargeting_top20_het_heatmap.png")
# setting rowtitle
rowtitle <- "TFs"
# setting columntitle
plot_title <- "Differential Targeting of Top 20 TFs within Cell-types in Setbp1 S858R Cerebral Cortex"

targeting_heatmap(annotation_colors, data, meta_colname, plot_path, rowtitle, plot_title, show_names = TRUE, height = 2000)
```

```{r GO-difftargeting-top20genes-het-enriched}
## pathway analysis and plotting
data <- cortex_het_enriched_top_genes

for (i in cell_names) {
  top <- data[data$cell_type == i, ]
  genes <- as.character(top$gene)
  # submit genes to pathway analysis function ----------
  fea_result_filt <- fea(genes = genes, organism = "mmusculus", max_term = 1000, min_term = 1)
  png(paste0(here("results/diff_targeting/"), "DiffTargeting_", i, "_top20_genes_het_enriched.png"), width = 1000, height = 1000)
  # plot dotplot and add title based on input variable ----------
  plot <- bubbleplot(fea_result_filt) +
    ggtitle(paste0("Pathways Associated with Top 20 Differentially Targeted Genes in Setbp1 S858R ", i)) +
    theme(plot.title = element_text(hjust = 0.5, size = 12)) +
    theme_minimal()
  print(plot)
  dev.off()
  print(paste0("pathways investigated and plotted in ", i))
}
```
[1] "pathways investigated and plotted in astrocytes"
[1] "pathways investigated and plotted in excitatory_neurons"
[1] "pathways investigated and plotted in fibro_cortex"
[1] "pathways investigated and plotted in inhibitory_neurons"
[1] "pathways investigated and plotted in microglia"
[1] "pathways investigated and plotted in mural"
[1] "pathways investigated and plotted in oligodendrocytes"
[1] "pathways investigated and plotted in opcs"

```{r}
## pathway analysis and plotting
data <- cortex_het_enriched_top_TFs

for (i in cell_names) {
  # grab genes/TFs from cell type i ----------
  top <- data[data$cell_type == i, ]
  genes <- as.character(top$TF)
  # submit genes to pathway analysis function ----------
  fea_result_filt <- fea(genes = genes, organism = "mmusculus", max_term = 1000, min_term = 1)
  png(paste0(here("results/diff_targeting/"), "DiffTargeting_", i, "_top20_TFs_het_enriched.png"), width = 1000, height = 1000)
  # plot dotplot and add title based on input variable ----------
  plot <- bubbleplot(fea_result_filt) +
    ggtitle(paste0("Pathways Associated with Top 20 Differentially Targeted TFs in Setbp1 S858R ", i)) +
    theme(plot.title = element_text(hjust = 0.5, size = 12)) +
    theme_minimal()
  print(plot)
  dev.off()
  print(paste0("pathways investigated and plotted in ", i))
}
```
[1] "pathways investigated and plotted in astrocytes"
[1] "pathways investigated and plotted in excitatory_neurons"
[1] "pathways investigated and plotted in fibro_cortex"
[1] "pathways investigated and plotted in inhibitory_neurons"
[1] "pathways investigated and plotted in microglia"
[1] "pathways investigated and plotted in mural"
[1] "pathways investigated and plotted in oligodendrocytes"
[1] "pathways investigated and plotted in opcs"

```{r find-top-diff-targeted-genes-ctrl-celltypes}
cell_names <- colnames(ctrl_enriched) %>% .[which(. != "gene")] # note: need to change for gene, grabbing cell type names
cell_names

# rank genes by targeting for each cell type , find top 20
gene_diff_ctrl <- list() # creating empty list to store within

for (i in cell_names) {
  rank_ctrl <- ctrl_enriched[, names(ctrl_enriched) %in% c("gene", i)] %>% # note: need to change TF to gene where applicable
    .[order(.[i], decreasing = TRUE), ] %>%
    mutate(rank = seq_len(nrow(.))) %>%
    mutate(cell_type = i) %>%
    slice(1:20) %>%
    rename(targeting_score = i)
  gene_diff_ctrl[[i]] <- rank_ctrl
}

cortex_ctrl_enriched_top_genes <- do.call(bind_rows, gene_diff_ctrl) # collapsing list into single df using row binding

# rank genes by targeting for each cell type for all data
gene_diff_ctrl <- list() # creating empty list to store within

for (i in cell_names) {
  rank_ctrl <- ctrl_enriched[, names(ctrl_enriched) %in% c("gene", i)] %>% # note: need to change TF to gene where applicable
    .[order(.[i], decreasing = TRUE), ] %>%
    mutate(rank = seq_len(nrow(.))) %>%
    mutate(cell_type = i) %>%
    rename(targeting_score = i)
  gene_diff_ctrl[[i]] <- rank_ctrl
}

cortex_ctrl_enriched_all_genes <- do.call(bind_rows, gene_diff_ctrl) # collapsing list into single df using row binding
```

```{r visualize-ctrl-enriched-gene-top20-heatmap}
## formatting input data
data <- cortex_ctrl_enriched_top_genes[, c("gene", "targeting_score", "cell_type"), drop = FALSE]
data <- pivot_wider(data, names_from = c("cell_type"), values_from = "targeting_score", names_repair = "minimal")
data[is.na(data)] <- 0
names <- data$gene # add gene names as rownames
data <- data[, -1, drop = FALSE] # drop gene column
rownames(data) <- names
# setting meta colname
meta_colname <- "cell_type"
# setting plot path
plot_path <- here("results/diff_targeting/cortex_gene_DIFFtargeting_top20_ctrl_heatmap.png")
# setting rowtitle
rowtitle <- "genes"
# setting columntitle
plot_title <- "Differential Targeting of Top 20 Genes within Cell-types in Setbp1 S858R Cerebral Cortex"

targeting_heatmap(annotation_colors, data, meta_colname, plot_path, rowtitle, plot_title, show_names = TRUE, height = 2000)
```


```{r find-top-diff-targeted-ctrl-enriched-TF-top20}
cell_names <- colnames(ctrl_enriched_TF) %>% .[which(. != "TF")] # note: need to change for gene, grabbing cell type names
cell_names

# rank TFs by targeting for each cell type , find top 20
TF_diff_ctrl <- list() # creating empty list to store within

for (i in cell_names) {
  rank_ctrl <- ctrl_enriched_TF[, names(ctrl_enriched_TF) %in% c("TF", i)] %>% # note: need to change TF to gene where applicable
    .[order(.[i], decreasing = TRUE), ] %>%
    mutate(rank = seq_len(nrow(.))) %>%
    mutate(cell_type = i) %>%
    slice(1:20) %>%
    rename(targeting_score = i)
  TF_diff_ctrl[[i]] <- rank_ctrl
}

cortex_ctrl_enriched_top_TFs <- do.call(bind_rows, TF_diff_ctrl) # collapsing list into single df using row binding

# rank TFs by targeting for each cell type for all data
TF_diff_ctrl <- list() # creating empty list to store within

for (i in cell_names) {
  rank_ctrl <- ctrl_enriched_TF[, names(ctrl_enriched_TF) %in% c("TF", i)] %>% # note: need to change TF to TF where applicable
    .[order(.[i], decreasing = TRUE), ] %>%
    mutate(rank = seq_len(nrow(.))) %>%
    mutate(cell_type = i) %>%
    rename(targeting_score = i)
  TF_diff_ctrl[[i]] <- rank_ctrl
}

cortex_ctrl_enriched_all_TFs <- do.call(bind_rows, TF_diff_ctrl) # collapsing list into single df using row binding
```

```{r visualize-ctrl-enriched-TF-top20-heatmap}
## formatting input data
data <- cortex_ctrl_enriched_top_TFs[, c("TF", "targeting_score", "cell_type"), drop = FALSE]
data <- pivot_wider(data, names_from = c("cell_type"), values_from = "targeting_score", names_repair = "minimal")
data[is.na(data)] <- 0
names <- data$TF # add TF names as rownames
data <- data[, -1, drop = FALSE] # drop TF column
rownames(data) <- names
# setting meta colname
meta_colname <- "cell_type"
# setting plot path
plot_path <- here("results/diff_targeting/cortex_TF_DIFFtargeting_top20_ctrl_heatmap.png")
# setting rowtitle
rowtitle <- "TFs"
# setting columntitle
plot_title <- "Differential Targeting of Top 20 TFs within Cell-types in Setbp1 S858R Cerebral Cortex"

targeting_heatmap(annotation_colors, data, meta_colname, plot_path, rowtitle, plot_title, show_names = TRUE, height = 2000)
```

```{r GO-difftargeting-top20genes-ctrl-enriched}
## pathway analysis and plotting
data <- cortex_ctrl_enriched_top_genes

for (i in cell_names) {
  # grab genes/TFs from cell type i ----------
  top <- data[data$cell_type == i, ]
  genes <- as.character(top$gene)
  # submit genes to pathway analysis function ----------
  fea_result_filt <- fea(genes = genes, organism = "mmusculus", max_term = 1000, min_term = 1)
  png(paste0(here("results/diff_targeting/"), "DiffTargeting_", i, "_top20_genes_ctrl_enriched.png"), width = 1000, height = 1000)
  # plot dotplot and add title based on input variable ----------
  plot <- bubbleplot(fea_result_filt) +
    ggtitle(paste0("Pathways Associated with Top 20 Differentially Targeted genes in Setbp1 S858R ", i)) +
    theme(plot.title = element_text(hjust = 0.5, size = 12)) +
    theme_minimal()
  print(plot)
  dev.off()
  print(paste0("pathways investigated and plotted in ", i))
}
```
[1] "pathways investigated and plotted in astrocytes"
[1] "pathways investigated and plotted in excitatory_neurons"
[1] "pathways investigated and plotted in fibro_cortex"
[1] "pathways investigated and plotted in inhibitory_neurons"
[1] "pathways investigated and plotted in microglia"
[1] "pathways investigated and plotted in mural"
[1] "pathways investigated and plotted in oligodendrocytes"
[1] "pathways investigated and plotted in opcs"

```{r}
## pathway analysis and plotting
data <- cortex_ctrl_enriched_top_TFs

for (i in cell_names) {
  # grab genes/TFs from cell type i ----------
  top <- data[data$cell_type == i, ]
  genes <- as.character(top$TF)
  # submit genes to pathway analysis function ----------
  fea_result_filt <- fea(genes = genes, organism = "mmusculus", max_term = 1000, min_term = 1)
  png(paste0(here("results/diff_targeting/"), "DiffTargeting_", i, "_top20_TFs_ctrl_enriched.png"), width = 1000, height = 1000)
  # plot dotplot and add title based on input variable ----------
  plot <- bubbleplot(fea_result_filt) +
    ggtitle(paste0("Pathways Associated with Top 20 Differentially Targeted TFs in Setbp1 S858R ", i)) +
    theme(plot.title = element_text(hjust = 0.5, size = 12)) +
    theme_minimal()
  print(plot)
  dev.off()
  print(paste0("pathways investigated and plotted in ", i))
}
```
[1] "pathways investigated and plotted in astrocytes"
[1] "pathways investigated and plotted in excitatory_neurons"
[1] "pathways investigated and plotted in fibro_cortex"
[1] "pathways investigated and plotted in inhibitory_neurons"
[1] "pathways investigated and plotted in microglia"
[1] "pathways investigated and plotted in mural"
[1] "pathways investigated and plotted in oligodendrocytes"
[1] "pathways investigated and plotted in opcs"

```{r}
proc.time()
```
     user    system   elapsed 
 5609.788  1391.155 82715.950
 
```{r}
# run style
#style_file(here("src/network_scripts/differential_targeting/Setbp1_DiffTargetingCortex_01.Rmd")) #commented out after being run once
# lintr was run as well
```

```{r}
sessionInfo()
```
R version 4.1.3 (2022-03-10)
Platform: x86_64-pc-linux-gnu (64-bit)
Running under: Ubuntu 20.04.5 LTS

Matrix products: default
BLAS/LAPACK: /usr/lib/x86_64-linux-gnu/openblas-pthread/libopenblasp-r0.3.8.so

locale:
[1] C

attached base packages:
[1] grid      stats     graphics  grDevices utils     datasets  methods   base     

other attached packages:
 [1] lintr_3.0.2           styler_1.9.0          here_1.0.1            gprofiler2_0.2.1      readr_2.1.4           circlize_0.4.15      
 [7] magrittr_2.0.3        RColorBrewer_1.1-3    ComplexHeatmap_2.10.0 ggpubr_0.6.0          ggplot2_3.4.1         tidyr_1.3.0          
[13] limma_3.50.3          dplyr_1.1.0           data.table_1.14.8     netZooR_1.2.1         matrixcalc_1.0-6      yarn_1.20.0          
[19] pandaR_1.26.0         Biobase_2.54.0        BiocGenerics_0.40.0   reticulate_1.28       igraph_1.4.1         

loaded via a namespace (and not attached):
  [1] rappdirs_0.3.3              rtracklayer_1.54.0          pbdZMQ_0.3-9                AnnotationForge_1.36.0      R.methodsS3_1.8.2          
  [6] bumphunter_1.36.0           minfi_1.40.0                bit64_4.0.5                 knitr_1.42                  R.utils_2.12.2             
 [11] DelayedArray_0.20.0         KEGGREST_1.34.0             RCurl_1.98-1.10             GEOquery_2.62.2             doParallel_1.0.17          
 [16] generics_0.1.3              GenomicFeatures_1.46.5      preprocessCore_1.56.0       callr_3.7.3                 RSQLite_2.3.0              
 [21] chron_2.3-59                bit_4.0.5                   tzdb_0.3.0                  base64url_1.4               xml2_1.3.3                 
 [26] SummarizedExperiment_1.24.0 assertthat_0.2.1            STRINGdb_2.6.5              xfun_0.37                   hms_1.1.2                  
 [31] evaluate_0.20               fansi_1.0.4                 restfulr_0.0.15             scrime_1.3.5                progress_1.2.2             
 [36] caTools_1.18.2              dbplyr_2.3.1                Rgraphviz_2.38.0            DBI_1.1.3                   htmlwidgets_1.6.1          
 [41] reshape_0.8.9               stats4_4.1.3                purrr_1.0.1                 hash_2.2.6.2                ellipsis_0.3.2             
 [46] backports_1.4.1             permute_0.9-7               annotate_1.72.0             biomaRt_2.50.3              sparseMatrixStats_1.6.0    
 [51] MatrixGenerics_1.6.0        vctrs_0.5.2                 remotes_2.4.2               penalized_0.9-52            abind_1.4-5                
 [56] cachem_1.0.7                withr_2.5.0                 grr_0.9.5                   vegan_2.6-4                 GenomicAlignments_1.30.0   
 [61] prettyunits_1.1.1           mclust_6.0.0                cyclocomp_1.1.0             cluster_2.1.2               IRdisplay_1.1              
 [66] lazyeval_0.2.2              crayon_1.5.2                uchardet_1.1.1              genefilter_1.76.0           labeling_0.4.2             
 [71] edgeR_3.36.0                pkgconfig_2.0.3             GenomeInfoDb_1.30.1         nlme_3.1-155                nnet_7.3-17                
 [76] rlang_1.0.6                 RJSONIO_1.3-1.8             lifecycle_1.0.3             downloader_0.4              filelock_1.0.2             
 [81] BiocFileCache_2.2.1         rex_1.2.1                   GOstats_2.60.0              quantro_1.28.0              rprojroot_2.0.3            
 [86] matrixStats_0.63.0          graph_1.72.0                rngtools_1.5.2              base64_2.0.1                Matrix_1.5-3               
 [91] IRkernel_1.3.2              carData_3.0-5               Rhdf5lib_1.16.0             base64enc_0.1-3             processx_3.8.0             
 [96] GlobalOptions_0.1.2         png_0.1-8                   viridisLite_0.4.1           rjson_0.2.21                bitops_1.0-7               
[101] R.oo_1.25.0                 KernSmooth_2.23-20          rhdf5filters_1.6.0          Biostrings_2.62.0           blob_1.2.3                 
[106] DelayedMatrixStats_1.16.0   doRNG_1.8.6                 shape_1.4.6                 stringr_1.5.0               nor1mix_1.3-0              
[111] R.cache_0.16.0              rstatix_0.7.2               S4Vectors_0.32.4            ggsignif_0.6.4              scales_1.2.1               
[116] memoise_2.0.1               GSEABase_1.56.0             plyr_1.8.8                  hexbin_1.28.2               gplots_3.1.3               
[121] zlibbioc_1.40.0             compiler_4.1.3              BiocIO_1.4.0                illuminaio_0.36.0           plotrix_3.8-2              
[126] clue_0.3-64                 Rsamtools_2.10.0            cli_3.6.0                   XVector_0.34.0              Category_2.60.0            
[131] ps_1.7.2                    MASS_7.3-55                 mgcv_1.8-39                 tidyselect_1.2.0            stringi_1.7.12             
[136] yaml_2.3.7                  askpass_1.1                 locfit_1.5-9.7              ggrepel_0.9.3               tools_4.1.3                
[141] parallel_4.1.3              rstudioapi_0.14             uuid_1.1-0                  foreach_1.5.2               farver_2.1.1               
[146] digest_0.6.31               proto_1.0.0                 quadprog_1.5-8              Rcpp_1.0.10                 GenomicRanges_1.46.1       
[151] car_3.1-1                   siggenes_1.68.0             broom_1.0.3                 org.Hs.eg.db_3.14.0         httr_1.4.5                 
[156] ggdendro_0.1.23             AnnotationDbi_1.56.2        colorspace_2.1-0            XML_3.99-0.13               fs_1.6.1                   
[161] IRanges_2.28.0              splines_4.1.3               RBGL_1.70.0                 multtest_2.50.0             plotly_4.10.1              
[166] xtable_1.8-4                jsonlite_1.8.4              R6_2.5.1                    RUnit_0.4.32                gsubfn_0.7                 
[171] pillar_1.8.1                htmltools_0.5.4             glue_1.6.2                  fastmap_1.1.1               BiocParallel_1.28.3        
[176] beanplot_1.3.1              codetools_0.2-18            utf8_1.2.3                  lattice_0.20-45             tibble_3.1.8               
[181] sqldf_0.4-11                curl_5.0.0                  gtools_3.9.4                GO.db_3.14.0                openssl_2.0.5              
[186] survival_3.3-1              rmarkdown_2.20              repr_1.1.6                  desc_1.4.2                  munsell_0.5.0              
[191] GetoptLong_1.0.5            rhdf5_2.38.1                GenomeInfoDbData_1.2.7      iterators_1.0.14            RCy3_2.14.2                
[196] HDF5Array_1.22.1            reshape2_1.4.4              gtable_0.3.1               
