---
title: "02_Setbp1_DiffTargeting_FEA"
author: "Jordan Whitlock"
date: '2023-05-24'
output: html_document
---
using docker 1.0.6

Goal: Perform Functional Enrichment Analysis (FEA) on heterozygous enriched, differentially targeted genes from the Setbp1 gene set using Gene Ontology and Human Phenotype from gprofiler. 

```{r libraries-processingtime-seed}
set.seed(2178)
library(ComplexHeatmap)
library(gprofiler2)
library(here)
library(styler)
library(lintr)
library(tidyverse)
library(patchwork)
library(cowplot)
library(org.Mm.eg.db)
source(here("src/functions/functions.R"))
ptm <- proc.time()
```

```{r loading-geneset}
# loading gene set
setbp1_genes <- read.csv(here("results/seurat/setbp1_targets.csv"))
setbp1_genes <- setbp1_genes[, -1] %>% as.vector()
setbp1_genes <- append(setbp1_genes, "Setbp1")
```

Identifying differentially targeted genes from the Setbp1 gene set
```{r loading-DT-genes}
# loading het enriched DT genes and filtering for gene set
load(here("results/diff_targeting/cortex_het_enriched_gene.Rdata"))
cortex_DT <- het_enriched

cortex_genes <- cortex_DT[cortex_DT$gene %in% setbp1_genes, ]
cortex_genes <- cortex_genes$gene # 195 genes

load(here("results/diff_targeting/kidney_het_enriched_gene.Rdata"))
kidney_DT <- het_enriched

kid_genes <- kidney_DT[kidney_DT$gene %in% setbp1_genes, ]
kid_genes <- kid_genes$gene # 194 genes

rm(het_enriched)
```

Investigating whether there are any DT genes from the gene set that are brain-specific, kidney-specific, or DT in both tissues:
```{r tissue-specific-DT-genes}
both <- intersect(kid_genes, cortex_genes)
both # 194 genes

kid_specific <- setdiff(kid_genes, cortex_genes)
kid_specific # no genes

cortex_specific <- setdiff(cortex_genes, kid_genes)
cortex_specific # Trex1
```

Investigating whether there are any DT genes from the gene set that are cerebral cortex cell-type specific 
```{r cortex-cell-specific-DT-geneset}
data <- cortex_DT[cortex_DT$gene %in% setbp1_genes, ]

# formatting data
data <- pivot_longer(data,
  cols = c(2:9),
  values_to = "DT Score",
  names_to = "Cell Type"
)
data <- subset(data, `DT Score` != 0)

# Count the occurrences of each value in the column
value_counts <- table(data$gene)

# Get the values that occur only once
unique_values <- names(value_counts[value_counts == 1])

# Subset the dataframe based on the unique values
DT_cerebral_cts <- data[data$gene %in% unique_values, ]
write.csv(DT_cerebral_cts,
  here("results/diff_targeting/cortex_cts_DT_genes.csv"),
  row.names = FALSE
)
```

Investigating whether there are any DT genes from the gene set that are cerebral kidney cell-type specific 
```{r kidney-cell-specific-DT-geneset}
data <- kidney_DT[kidney_DT$gene %in% setbp1_genes, ]

# formatting data
data <- pivot_longer(data,
  cols = c(2:18),
  values_to = "DT Score",
  names_to = "Cell Type"
)
data <- subset(data, `DT Score` != 0)

# Count the occurrences of each value in the column
value_counts <- table(data$gene)

# Get the values that occur only once
unique_values <- names(value_counts[value_counts == 1])

# Subset the dataframe based on the unique values
DT_kidney_cts <- data[data$gene %in% unique_values, ] # None
```

Using a custom background for gost (custom_bg), of the Setbp1 Target gene set. 
* Rationale: (from LW) The background (others call the 'universe') for a FEA is just saying all possible genes that could have been used as input (i.e. if input genes were filtered only for protein-coding genes, there's no chance a non-protein-coding gene would show up, so the stats need to be adjusted for what's possible.) 
* custom_bg: vector of gene names to use as a statistical background. If given, the domain_scope is by default set to "custom", if domain_scope is set to "custom_annotated", then this is used instead

```{r fea-GO-HP-cortex-setbp1-geneset}
data <- cortex_DT[cortex_DT$gene %in% setbp1_genes, ]
# formatting data
data <- pivot_longer(data,
  cols = c(2:9),
  values_to = "DT Score",
  names_to = "Cell Type"
)

data <- subset(data, `DT Score` != 0)

# indivudal cell types
cell_types <- unique(data$`Cell Type`)

fea_celltypes <- list()
fea_celltypes_nonsig <- list()

for (i in cell_types) {
  temp <- data[data$`Cell Type` == i, ]
  genes <- as.character(temp$gene)
  if (i == "pericytes") { #
    fea_result_filt <- fea_custom_no_sig(
      genes = genes,
      organism = "mmusculus",
      sources = c("GO", "HP")
    )
    fea_celltypes_nonsig[[i]] <- fea_result_filt
    print(paste0("nonsig fea analysis complete on ", i))
  } else {
    fea_result_filt <- fea_custom(
      genes = genes,
      organism = "mmusculus",
      sources = c("GO", "HP"),
      custom_bg = setbp1_genes
    )
    fea_celltypes[[i]] <- fea_result_filt
    print(paste0("fea analysis complete on ", i))
  }
}

# Unlist and annotate source
cortex_fea <- fea_celltypes %>%
  map_df(~ tibble::as_tibble(.x), .id = "source")

data <- cortex_fea[, c("term_name", "source", "recall")]

# saving fea results
# cortex_fea <- cortex_fea[, -which(names(cortex_fea) == "parents")] # removing column of list items in order to save
# write.csv(cortex_fea, here("results/diff_targeting/cortex_fea_DT_genes.csv"))

# wrangle for plotting
data <- pivot_wider(data,
  names_from = source,
  values_from = recall
) %>%
  mutate_all(~ replace_na(., 0))

# plot GO terms
names <- data$term_name # add names as rownames
cell_type <- colnames(data) %>% .[-1]
data <- data[, -which(names(data) == "term_name")] # drop term_name column
rownames(data) <- names

meta <- as.data.frame(cell_type)
colnames(meta) <- "Cell Type"

## set heatmap annotations
annotation_colors <- list(`Cell Type` = c(
  "astrocytes" = "#6968B4",
  "opcs" = "#00C1AA",
  # "pericytes" = "#027461",
  "fibro_cortex" = "#0a2f6f",
  "excitatory_neurons" = "#DC71FA",
  "inhibitory_neurons" = "#00BBDB",
  "microglia" = "#C216B7",
  "oligodendrocytes" = "#7997FF"
))

heat.anno <- HeatmapAnnotation(
  df = meta,
  show_annotation_name = TRUE,
  col = annotation_colors
)

## convert data to matrix
mat <- as.matrix(data)

## plot heatmap
cor_heatmap <- Heatmap(mat,
  col = colorRampPalette(c("black", "yellow", "orange", "red"))(4),
  heatmap_legend_param = list(title = "Recall"),
  cluster_rows = TRUE,
  cluster_columns = TRUE,
  column_order = NULL,
  show_row_dend = TRUE,
  show_column_dend = TRUE,
  show_row_names = TRUE,
  show_column_names = FALSE,
  use_raster = TRUE,
  raster_device = c("png"),
  bottom_annotation = NULL,
  top_annotation = heat.anno,
  row_title = "Pathways", row_title_side = "left",
  row_title_gp = gpar(fontface = "bold"),
  row_names_gp = gpar(fontface = "bold", fontfamily = "Helvetica")
)


plot <- plot_grid(grid.grabExpr(draw(cor_heatmap,
  annotation_legend_side = "right",
  legend_grouping = "original"
)))
png(here("results/diff_targeting/cortex_het_enriched_DT_from_geneset_GO_HP_pathways_recall.png"),
  res = 250,
  height = 5000,
  width = 3000
)
plot
dev.off()
```

```{r fea-GO-HP-kidney-setbp1-geneset}
data <- kidney_DT[kidney_DT$gene %in% setbp1_genes, ]

# formatting data
data <- pivot_longer(data,
  cols = c(2:18),
  values_to = "DT Score",
  names_to = "Cell Type"
)
data <- subset(data, `DT Score` != 0)

# indivudal cell types
cell_types <- unique(data$`Cell Type`)

fea_celltypes <- list()

for (i in cell_types) {
  temp <- data[data$`Cell Type` == i, ]
  genes <- as.character(temp$gene)
  fea_result_filt <- fea_custom(genes = genes, organism = "mmusculus", sources = c("GO", "HP"), custom_bg = setbp1_genes)
  fea_celltypes[[i]] <- fea_result_filt
  print(paste0("fea analysis complete on ", i))
}

# Unlist and annotate source
kidney_fea <- fea_celltypes %>%
  map_df(~ tibble::as_tibble(.x), .id = "source")

data <- kidney_fea[, c("term_name", "source", "recall")]

# saving fea results
# kidney_fea <- kidney_fea[, -which(names(kidney_fea) == "parents")] # removing column of list items in order to save
# write.csv(kidney_fea, here("results/diff_targeting/kidney_fea_DT_genes.csv"))

# wrangle for plotting
data <- pivot_wider(data,
  names_from = source,
  values_from = recall
) %>%
  mutate_all(~ replace_na(., 0))

# plot GO terms
names <- data$term_name # add names as rownames
cell_type <- colnames(data) %>% .[-1]
data <- data[, -which(names(data) == "term_name")] # drop term_name column
rownames(data) <- names

meta <- as.data.frame(cell_type)
colnames(meta) <- "Cell Type"
rownames(meta) <- colnames(mat)
## convert data to matrix
mat <- as.matrix(data)

## set heatmap annotations
annotation_colors <- list(`Cell Type` = c(
    "Bcell" = "#FFBE1D",
    "CDIC_typeA" = "#4E3801",
    "CDIC_typeB" = "#EEDF37",
    "CDPC" = "#F03F00",
    "DCT" = "#6D0404",
    "DLH" = "#AA6320",
    "endothelial" = "#CF9400",
    "fibroblasts" = "#FFC8C4",
    "LOH" = "#AA937E",
    "macrophages" = "#C70F0F",
    "PCTS1" = "#A85A5A",
    "podocytes" = "#BD085E",
    "PST" = "#948802",
    "pericytes" = "#FFE196",
    "PCTS2" = "#FF7600",
    "PT" = "#F8766D",
    "smcs" = "#FFE196"
))


heat.anno <- HeatmapAnnotation(df = meta, show_annotation_name = TRUE, col = annotation_colors)


## plot heatmap
kid_heatmap <- Heatmap(mat,
  col = colorRampPalette(c("black", "yellow", "orange", "red"))(4),
  heatmap_legend_param = list(title = "Recall"),
  cluster_rows = TRUE,
  cluster_columns = TRUE,
  column_order = NULL,
  show_row_dend = TRUE,
  show_column_dend = TRUE,
  show_row_names = TRUE,
  show_column_names = FALSE,
  use_raster = TRUE,
  raster_device = c("png"),
  bottom_annotation = NULL,
  top_annotation = heat.anno,
  row_title = "Pathways", row_title_side = "left",
  row_title_gp = gpar(fontface = "bold"),
  row_names_gp = gpar(fontface = "bold", fontfamily = "Helvetica")
)


plot <- plot_grid(grid.grabExpr(draw(kid_heatmap,
  annotation_legend_side = "right",
  legend_grouping = "original"
)))
png(here("results/diff_targeting/kidney_het_enriched_DT_geneset_GO_HP_pathways_recall.png"),
  res = 250,
  height = 5000,
  width = 3000
)
plot
dev.off()
```

```{r processing-time}
fptm <- proc.time() - ptm
fptm[3] / 60
```
  elapsed 
0.4909333

```{r style-file}
# run style
style_file(here("src/network_scripts/differential_targeting/02_Setbp1_DiffTargeting_FEA.Rmd")) # commented out after being run once
# lintr was run as well
```

```{r session-info}
sessionInfo()
```
R version 4.1.3 (2022-03-10)
Platform: x86_64-pc-linux-gnu (64-bit)
Running under: Ubuntu 20.04.6 LTS

Matrix products: default
BLAS/LAPACK: /usr/lib/x86_64-linux-gnu/openblas-pthread/libopenblasp-r0.3.8.so

locale:
[1] C

attached base packages:
[1] stats4    grid      stats     graphics  grDevices utils     datasets  methods   base     

other attached packages:
 [1] org.Mm.eg.db_3.14.0   AnnotationDbi_1.56.2  IRanges_2.28.0        S4Vectors_0.32.4      Biobase_2.54.0        BiocGenerics_0.40.0  
 [7] cowplot_1.1.1         patchwork_1.1.2       lubridate_1.9.2       forcats_1.0.0         stringr_1.5.0         dplyr_1.1.0          
[13] purrr_1.0.1           readr_2.1.4           tidyr_1.3.0           tibble_3.2.1          ggplot2_3.4.2         tidyverse_2.0.0      
[19] lintr_3.0.2           styler_1.9.0          here_1.0.1            gprofiler2_0.2.1      ComplexHeatmap_2.10.0

loaded via a namespace (and not attached):
 [1] bitops_1.0-7           matrixStats_0.63.0     bit64_4.0.5            doParallel_1.0.17      RColorBrewer_1.1-3     httr_1.4.5            
 [7] GenomeInfoDb_1.30.1    rprojroot_2.0.3        R.cache_0.16.0         tools_4.1.3            utf8_1.2.3             R6_2.5.1              
[13] DBI_1.1.3              lazyeval_0.2.2         colorspace_2.1-0       GetoptLong_1.0.5       withr_2.5.0            tidyselect_1.2.0      
[19] processx_3.8.0         bit_4.0.5              compiler_4.1.3         cli_3.6.1              xml2_1.3.3             desc_1.4.2            
[25] plotly_4.10.1          scales_1.2.1           callr_3.7.3            digest_0.6.31          R.utils_2.12.2         XVector_0.34.0        
[31] pkgconfig_2.0.3        htmltools_0.5.4        fastmap_1.1.1          htmlwidgets_1.6.1      rlang_1.1.1            GlobalOptions_0.1.2   
[37] rstudioapi_0.14        RSQLite_2.3.0          shape_1.4.6            generics_0.1.3         jsonlite_1.8.4         R.oo_1.25.0           
[43] RCurl_1.98-1.10        magrittr_2.0.3         GenomeInfoDbData_1.2.7 munsell_0.5.0          fansi_1.0.4            lifecycle_1.0.3       
[49] R.methodsS3_1.8.2      stringi_1.7.12         zlibbioc_1.40.0        blob_1.2.3             parallel_4.1.3         crayon_1.5.2          
[55] Biostrings_2.62.0      KEGGREST_1.34.0        circlize_0.4.15        hms_1.1.2              knitr_1.42             ps_1.7.2              
[61] pillar_1.9.0           rjson_0.2.21           codetools_0.2-18       glue_1.6.2             rex_1.2.1              data.table_1.14.8     
[67] remotes_2.4.2          png_0.1-8              vctrs_0.6.2            tzdb_0.3.0             foreach_1.5.2          gtable_0.3.3          
[73] clue_0.3-64            cachem_1.0.8           xfun_0.37              cyclocomp_1.1.0        viridisLite_0.4.2      iterators_1.0.14
